---
title: "phix_DMS_fitness_data_normalization"
author: "Huijin Wei"
date: "2026-07-30"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Set working working directory
```{r}
here::here()

```



## Load Functions
```{r}
# Define a function to merge and filter datasets
merge_and_filter <- function(mut_info, fitness_info, merged_name) {
  if (exists(mut_info, envir = .GlobalEnv) & exists(fitness_info, envir = .GlobalEnv)) {
    merged_df <- merge(get(mut_info), get(fitness_info), by = "oligo_mut_region")
    
    # Remove rows where fitness_dimsum is NA
    if ("fitness_dimsum" %in% colnames(merged_df)) {
      merged_df <- merged_df %>% filter(!is.na(fitness_dimsum))
    } else {
      cat("Warning: fitness_dimsum column not found in", merged_name, "\n")
    }

    # Assign the cleaned data frame
    assign(merged_name, merged_df, envir = .GlobalEnv)
    
    # Print number of rows after filtering
    cat("Merged:", merged_name, "| Rows after filtering:", nrow(merged_df), "\n")
  } else {
    cat("Skipping:", mut_info, "or", fitness_info, "- Not found\n")
  }
}


# Function to find the modes using kernel density estimation
find_modes_from_df <- function(data_frame, column_name) {
  # Ensure the specified column exists
  if (!column_name %in% names(data_frame)) {
    stop("Column not found in the data frame.")
  }
  
  # Extract the specified column
  data_values <- data_frame[[column_name]]
  
  # Remove NA values
  data_values <- na.omit(data_values)
  
  # Ensure there are enough data points to compute density
  if (length(data_values) < 2) {
    return(NA)
  }
  
  # Kernel Density Estimation
  density_est <- density(data_values)
  
  # Find local maxima (modes)
  local_max <- which(diff(sign(diff(density_est$y))) == -2) + 1
  modes <- density_est$x[local_max]
  
  # Return modes as a comma-separated string
  return(ifelse(length(modes) > 0, paste(modes, collapse = ","), NA))
}


# Function to extract oligo_seq_pool and oligo_seq_lib properly
# 1) Only Substitutions Data
extract_oligo_info <- function(df_name) {
  # Case: *_libr1_dms_sub or similar
  if (grepl("_libr\\d+_dms_sub$", df_name)) {
    pool_part <- sub("_libr\\d+_dms_sub$", "", df_name)
    lib_part <- sub(".*_libr(\\d+)_dms_sub$", "r\\1", df_name)

  # Case: *_lib1_dms_HS_sub
  } else if (grepl("_lib\\d+_dms_HS_sub$", df_name)) {
    pool_part <- sub("_lib\\d+_dms_HS_sub$", "", df_name)
    lib_part <- sub(".*_lib(\\d+)_dms_HS_sub$", "\\1_HS", df_name)

  # Case: *_lib1_dms_nHS_sub
  } else if (grepl("_lib\\d+_dms_nHS_sub$", df_name)) {
    pool_part <- sub("_lib\\d+_dms_nHS_sub$", "", df_name)
    lib_part <- sub(".*_lib(\\d+)_dms_nHS_sub$", "\\1_nHS", df_name)

  # Case: *_lib1_dms_sub
  } else if (grepl("_lib\\d+_dms_sub$", df_name)) {
    pool_part <- sub("_lib\\d+_dms_sub$", "", df_name)
    lib_part <- sub(".*_lib(\\d+)_dms_sub$", "\\1", df_name)

  # Fallback (optional)
  } else {
    warning("Pattern not recognized:", df_name)
    pool_part <- NA
    lib_part <- NA
  }

  return(list(pool = pool_part, lib = lib_part))
}


# 2) All data
extract_oligo_info2 <- function(df_name) {
  # Case: *_libr1_dms
  if (grepl("_libr\\d+_dms$", df_name)) {
    pool_part <- sub("_libr\\d+_dms$", "", df_name)
    lib_part <- sub(".*_libr(\\d+)_dms$", "r\\1", df_name)

  # Case: *_lib1_dms_HS
  } else if (grepl("_lib\\d+_dms_HS$", df_name)) {
    pool_part <- sub("_lib\\d+_dms_HS$", "", df_name)
    lib_part <- sub(".*_lib(\\d+)_dms_HS$", "\\1_HS", df_name)

  # Case: *_lib1_dms_nHS
  } else if (grepl("_lib\\d+_dms_nHS$", df_name)) {
    pool_part <- sub("_lib\\d+_dms_nHS$", "", df_name)
    lib_part <- sub(".*_lib(\\d+)_dms_nHS$", "\\1_nHS", df_name)

  # Standard case: *_lib1_dms
  } else if (grepl("_lib\\d+_dms$", df_name)) {
    pool_part <- sub("_lib\\d+_dms$", "", df_name)
    lib_part <- sub(".*_lib(\\d+)_dms$", "\\1", df_name)

  # Fallback for unmatched pattern
  } else {
    warning("Pattern not recognized:", df_name)
    pool_part <- NA
    lib_part <- NA
  }

  return(list(pool = pool_part, lib = lib_part))
}


# Function to compute fitness modes for Synonymous and Nonsense mutations
compute_fitness_modes <- function(library_list) {
  # Initialize an empty list to store results
  results_list <- list()
  
  for (df_name in library_list) {
    df <- get(df_name)
    
    # Ensure required columns exist
    if (!("phix_aa_mut_class" %in% names(df)) || !("fitness_dimsum" %in% names(df))) {
      next
    }
    
    # Filter Synonymous and Nonsense mutations
    synonymous_df <- df[df$phix_aa_mut_class == "Synonymous", , drop = FALSE]
    nonsense_df <- df[df$phix_aa_mut_class == "Nonsense", , drop = FALSE]

    # Compute modes using Kernel Density Estimation
    synonymous_mode_str <- if (nrow(synonymous_df) > 0) find_modes_from_df(synonymous_df, "fitness_dimsum") else NA
    nonsense_mode_str <- if (nrow(nonsense_df) > 0) find_modes_from_df(nonsense_df, "fitness_dimsum") else NA
    
    # Count occurrences
    synonymous_count <- nrow(synonymous_df)
    nonsense_count <- nrow(nonsense_df)

    # Extract oligo_seq_pool and oligo_seq_lib using the helper function
    oligo_info <- extract_oligo_info(df_name)
    
    # Store results
    results_list[[df_name]] <- data.frame(
      library = df_name,
      oligo_seq_pool = oligo_info$pool,
      oligo_seq_lib = oligo_info$lib,
      synonymous_mode = synonymous_mode_str,
      nonsense_mode = nonsense_mode_str,
      syn_mode_final = NA,  # Placeholder for manual selection
      non_mode_final = NA,  # Placeholder for manual selection
      synonymous_count = synonymous_count,
      nonsense_count = nonsense_count,
      stringsAsFactors = FALSE
    )
  }
  
  # Combine all results into a single data frame
  results_df <- do.call(rbind, results_list)
  
  return(results_df)
}


## Normalization and Rescaling
# Function to normalize both fitness and propagate sigma for each data frame
normalize_fitness_and_sigma <- function(df_name, mode_summary_df) {
  df <- get(df_name)  # Retrieve the data frame

  # Extract corresponding mode and normalization factor
  mode_row <- mode_summary_df[mode_summary_df$library == df_name, ]

  syn_mode <- as.numeric(mode_row$syn_mode_final)
  non_mode <- as.numeric(mode_row$non_mode_final)
  norm_factor <- as.numeric(mode_row$normalization_factor)

  # Ensure modes are valid before proceeding
  if (is.na(syn_mode) || is.na(non_mode) || syn_mode == non_mode) {
    warning(paste("Skipping normalization for", df_name, "due to missing or identical modes"))
    return(df)  # Return unchanged df if modes are invalid
  }

  # Normalize fitness values
  df$fitness_normalized <- (df$fitness_dimsum - non_mode) / (syn_mode - non_mode)

  # Normalize sigma using the normalization factor (error propagation)
  if (!is.na(norm_factor)) {
    df$sigma_rescaled <- norm_factor * df$sigma_dimsum
  } else {
    warning(paste("Skipping sigma rescaling for", df_name, "due to missing normalization factor"))
  }

  # Assign the updated data frame back to the global environment
  assign(df_name, df, envir = .GlobalEnv)
  cat("Normalization complete for:", df_name, "\n")
}


# Function to normalize both fitness and propagate sigma for each data frame
normalize_fitness_and_sigma_all <- function(df_name, mode_summary_df) {
  df <- get(df_name)  # Retrieve the data frame
  
  # Ensure df contains oligo_seq_lib before proceeding
  if (!"oligo_seq_lib" %in% colnames(df)) {
    warning(paste("Skipping", df_name, "- missing oligo_seq_lib column"))
    return(df)
  }
  
  # Extract the oligo_seq_lib for this df
  oligo_lib <- unique(df$oligo_seq_lib)
  
  # Check if multiple unique values exist (unexpected case)
  if (length(oligo_lib) != 1) {
    warning(paste("Skipping", df_name, "- multiple unique oligo_seq_lib values found"))
    return(df)
  }
  
  # Extract mode values based on oligo_seq_lib instead of df_name
  mode_row <- mode_summary_df[mode_summary_df$oligo_seq_lib == oligo_lib, ]
  
  if (nrow(mode_row) == 0) {
    warning(paste("Skipping normalization for", df_name, "- no matching mode summary found"))
    return(df)
  }
  
  # Compute the normalization factors
  syn_mode <- as.numeric(mode_row$syn_mode_final)
  non_mode <- as.numeric(mode_row$non_mode_final)
  norm_factor <- as.numeric(mode_row$normalization_factor)
  
  # Ensure valid mode values before proceeding
  if (is.na(syn_mode) || is.na(non_mode) || syn_mode == non_mode) {
    warning(paste("Skipping normalization for", df_name, "- missing or identical modes"))
    return(df)
  }
  
  # Normalize fitness values
  df$fitness_normalized <- (df$fitness_dimsum - non_mode) / (syn_mode - non_mode)
  
  # Normalize sigma using the normalization factor (error propagation)
  if (!is.na(norm_factor)) {
    df$sigma_rescaled <- norm_factor * df$sigma_dimsum
  } else {
    warning(paste("Skipping sigma rescaling for", df_name, "- missing normalization factor"))
  }
  
  # Assign the updated data frame back to the global environment
  assign(df_name, df, envir = .GlobalEnv)
  cat("Normalization complete for:", df_name, "\n")
}


```



## Prepare Fitness Data and Mutant Design (DiMSum >- all variants)
### 1) Fitness data from DiMSum results - fitness replicates
Load raw fitness data from DiMSum results - phix_dimsum_fitness_libraries

```{r}
# load Rdata
phix_dimsum_fitness_libraries_loaded <- c(
  "pool1_twist5_lib6_dimsum_fitness",
  "pool1_twist6_lib7_dimsum_fitness",
  "pool1_twist7_lib8_dimsum_fitness",
  "pool1_twist8_lib9_dimsum_fitness",
  "pool1_twist9_lib10_dimsum_fitness",
  "pool1_twist10_lib11_dimsum_fitness",
  "pool1_twist11_lib12_dimsum_fitness",
  
  "pool2_twist3_lib3_dimsum_fitness",
  "pool2_twist4_lib4_dimsum_fitness",
  "pool2_twist5_lib5_dimsum_fitness",
  "pool2_twist13_lib13_dimsum_fitness",
  "pool2_twist15_lib15_dimsum_fitness",
  "pool2_twist20_lib21_dimsum_fitness",
  "pool2_twist21_lib22_dimsum_fitness",
  
  "pool3_twist14_lib14_dimsum_fitness",
  "pool3_twist16_lib16_dimsum_fitness",
  "pool3_twist17_lib17_dimsum_fitness",
  "pool3_twist18_lib18_dimsum_fitness",
  "pool3_twist19_lib19_dimsum_fitness",
  "pool3_twist20_1_lib20_dimsum_fitness",
  "pool3_twist29_lib30_dimsum_fitness",
  
  "pool4_twist1_lib1_dimsum_fitness",
  "pool4_twist2_lib2_dimsum_fitness",
  "pool4_twist22_lib23_dimsum_fitness",
  "pool4_twist23_lib24_dimsum_fitness",
  "pool4_twist24_lib25_dimsum_fitness",
  "pool4_twist25_lib26_dimsum_fitness",
  "pool4_twist26_lib27_dimsum_fitness",
  "pool4_twist27_lib28_dimsum_fitness",
  "pool4_twist28_lib29_dimsum_fitness",
  
  "pool4_twist30_libr1_dimsum_fitness"
)

for (lib in phix_dimsum_fitness_libraries_loaded) {

  file_path <- file.path(
    "data",
    "raw_data",
    "raw_fitness_data",
    paste0(lib, ".RData")
  )

  if (file.exists(file_path)) {

    # Load the data as `all_variants`
    load(file_path)

    # Filter variants by mean count
    if ("mean_count" %in% colnames(all_variants)) {
      all_variants <- all_variants %>%
        filter(mean_count > 100)
    } else {
      warning(
        "Column `mean_count` was not found in: ",
        lib
      )
    }

    # Assign the filtered data using the existing standardized name
    assign(
      lib,
      all_variants,
      envir = .GlobalEnv
    )

    cat(
      "Loaded:", file_path,
      "| Object:", lib,
      "| Rows after filtering:", nrow(all_variants),
      "\n"
    )

  } else {
    warning("File not found: ", file_path)
  }
}

# remove other files
rm(all_variants)
rm(doubles)
rm(singles)
rm(wildtype)


# save filtered raw fitness data
phix_dimsum_fitness_libraries <- mget(
  phix_dimsum_fitness_libraries_loaded,
  envir = .GlobalEnv
)

saveRDS(
  phix_dimsum_fitness_libraries,
  file = "./data/raw_data/phix_dimsum_fitness_libraries.rds",
  compress = "xz"
)


 # load filtered data
sub_libs <- readRDS("./data/raw_data/phix_dimsum_fitness_libraries.rds")
list2env(sub_libs, envir = .GlobalEnv)


```


### 2) Substitutions mutation annotations
All mutation design in the list - phix_mut_library_sub_designs

```{r}
## Load Substitution mut library design - all annotated
load("./data/mutant_library_data/phix_dms_substitutions_annotated.RData")
load("./data/mutant_library_data/pool4_twist30_libr1_mut_sub_info.RData")

# List of mutation library design data frames
phix_mut_library_sub_designs <- c(
  "pool1_twist5_lib6_mut_sub_info",
  "pool1_twist6_lib7_mut_sub_info",
  "pool1_twist7_lib8_mut_sub_info",
  "pool1_twist8_lib9_mut_sub_info",
  "pool1_twist9_lib10_mut_sub_info",
  "pool1_twist10_lib11_mut_sub_info",
  "pool1_twist11_lib12_mut_sub_info",
  
  "pool2_twist3_lib3_mut_sub_info",
  "pool2_twist4_lib4_mut_sub_info",
  "pool2_twist5_lib5_mut_sub_info",
  "pool2_twist13_lib13_mut_sub_info",
  "pool2_twist15_lib15_mut_sub_info",
  "pool2_twist20_lib21_mut_sub_info",
  "pool2_twist21_lib22_mut_sub_info",
  
  "pool3_twist14_lib14_mut_sub_info",
  "pool3_twist16_lib16_mut_sub_info",
  "pool3_twist17_lib17_mut_sub_info",
  "pool3_twist18_lib18_mut_sub_info",
  "pool3_twist19_lib19_mut_sub_info",
  "pool3_twist20_1_lib20_mut_sub_info",
  "pool3_twist29_lib30_mut_sub_info",
  
  "pool4_twist1_lib1_mut_sub_info",
  "pool4_twist2_lib2_mut_sub_info",
  "pool4_twist22_lib23_mut_sub_info",
  "pool4_twist23_lib24_mut_sub_info",
  "pool4_twist24_lib25_mut_sub_info",
  "pool4_twist25_lib26_mut_sub_info",
  "pool4_twist26_lib27_mut_sub_info",
  "pool4_twist27_lib28_mut_sub_info",
  "pool4_twist28_lib29_mut_sub_info",
  
  "pool4_twist30_libr1_mut_sub_info"
)


# filter substitution data
for (df_name in phix_dimsum_fitness_libraries_loaded) {
  
  df_subs_name <- paste0(df_name, "_subs")  # New name for filtered substitutions
  
  if (exists(df_name)) {
    df <- get(df_name)  # Retrieve fitness data frame
    
    # Step 1: Extract only substitution data
    df_subs <- df %>%
      filter(is.na(WT), indel == FALSE) %>%
      select(nt_seq, Nham_nt, mean_count, fitness1_uncorr, fitness2_uncorr, fitness3_uncorr, sigma1_uncorr, sigma2_uncorr, sigma3_uncorr, fitness, sigma)
    
    # Step 2: Rename columns for consistency
    colnames(df_subs) <- c("oligo_mut_region", "oligo_nt_mut_ham", "oligo_input_mean_count", "fitness_rep1_uncorr", "fitness_rep2_uncorr", "fitness_rep3_uncorr", "sigma_rep1_uncorr", "sigma_rep2_uncorr", "sigma_rep3_uncorr", "fitness_dimsum", "sigma_dimsum")
    
    # Step 3: Convert sequences to uppercase
    df_subs$oligo_mut_region <- toupper(df_subs$oligo_mut_region)
    
    # Assign the cleaned subset data frame
    assign(df_subs_name, df_subs)
    
    # Print the number of rows after filtering
    cat("Processed:", df_subs_name, "| Rows:", nrow(df_subs), "\n")
  } else {
    cat("Skipping:", df_name, "- Not found\n")
  }
}


# merge the fitness data with mutation information
for (mut_lib in phix_mut_library_sub_designs) {  
  # Generate the corresponding fitness substitution data frame name  
  fitness_df_name <- gsub("_mut_sub_info", "_dimsum_fitness_subs", mut_lib)  
  merged_df_name <- gsub("_mut_sub_info", "_dms_sub", mut_lib)  # New merged data frame name  

  # Check if both mutation design and fitness data exist before merging  
  if (exists(mut_lib) & exists(fitness_df_name)) {  
    merged_df <- merge(get(mut_lib), get(fitness_df_name), by = "oligo_mut_region")  

    # Remove rows where fitness_dimsum is NA  
    if ("fitness_dimsum" %in% colnames(merged_df)) {
      merged_df <- merged_df %>% filter(!is.na(fitness_dimsum))
    } else {
      cat("Warning: fitness_dimsum column not found in", merged_df_name, "\n")
    }

    # Assign merged data frame  
    assign(merged_df_name, merged_df, envir = .GlobalEnv)  

    # Print the number of merged rows after filtering  
    cat("Merged:", merged_df_name, "| Rows after filtering:", nrow(merged_df), "\n")  
  } else {  
    cat("Skipping:", mut_lib, "or", fitness_df_name, "- Not found\n")  
  }  
}  


## List all Substitutions data (including HS pilots)
merged_dms_sub_libraries <- c(
  "pool1_twist5_lib6_dms_sub",
  "pool1_twist6_lib7_dms_sub",
  "pool1_twist7_lib8_dms_sub",
  "pool1_twist8_lib9_dms_sub",
  "pool1_twist9_lib10_dms_sub",
  "pool1_twist10_lib11_dms_sub",
  "pool1_twist11_lib12_dms_sub",
  
  "pool2_twist3_lib3_dms_sub",
  "pool2_twist4_lib4_dms_sub",
  "pool2_twist5_lib5_dms_sub",
  "pool2_twist13_lib13_dms_sub",
  "pool2_twist15_lib15_dms_sub",
  "pool2_twist20_lib21_dms_sub",
  "pool2_twist21_lib22_dms_sub",
  
  "pool3_twist14_lib14_dms_sub",
  "pool3_twist16_lib16_dms_sub",
  "pool3_twist17_lib17_dms_sub",
  "pool3_twist18_lib18_dms_sub",
  "pool3_twist19_lib19_dms_sub",
  "pool3_twist20_1_lib20_dms_sub",
  "pool3_twist29_lib30_dms_sub",
  
  "pool4_twist1_lib1_dms_sub",
  "pool4_twist2_lib2_dms_sub",
  "pool4_twist22_lib23_dms_sub",
  "pool4_twist23_lib24_dms_sub",
  "pool4_twist24_lib25_dms_sub",
  "pool4_twist25_lib26_dms_sub",
  "pool4_twist26_lib27_dms_sub",
  "pool4_twist27_lib28_dms_sub",
  "pool4_twist28_lib29_dms_sub",
  
  "pool4_twist30_libr1_dms_sub"
)


```



## Fitness Data Normalization
### 1) get Synonymous and Nonsense Modes
get Synonymous and Nonsense modes within each library by kernel density estimation;

```{r}
# Run the function on the merged list to extract the modes of Synonymous and Nonsense
phix_dms_sub_libraries_mode_summary <- compute_fitness_modes(merged_dms_sub_libraries)
head(phix_dms_sub_libraries_mode_summary, 2)

# Check the distribution and manually pick one value from the all found as the final one for Normalization;
## Pool1
#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 6]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 6]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 6] <- 0.256133449073762
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 6] <- -2.20360978128535

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 7]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 7]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 7] <- 0.574511080258657
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 7] <- -2.19962194245897

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 8]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 8]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 8] <- 0.323329633178167
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 8] <- -2.0236375378613

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 9]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 9]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 9] <- 0.481198040170701
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 9] <- -1.76425285993907

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 10]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 10]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 10] <- 0.372471855150933
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 10] <- -2.222327385966

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 11]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 11]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 11] <- 0.416686180893889
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 11] <- -1.69552899007643

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 12]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 12]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 12] <- 0.315667447361166
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 12] <- -1.63594273293534

## Pool1 HS 
#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "7_HS"]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "7_HS"]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "7_HS"] <- 0.517318147097636
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "7_HS"] <- -1.60300103310104

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "8_HS"]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "8_HS"]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "8_HS"] <- 0.424323360694605
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "8_HS"] <- -1.91226637349082

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "8_nHS"]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "8_nHS"]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "8_nHS"] <- 0.325526926004084
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "8_nHS"] <- -2.34249909739244

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "9_HS"]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "9_HS"]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "9_HS"] <- 0.49854882764141
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "9_HS"] <- -1.52861403163384

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "10_HS"]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "10_HS"]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "10_HS"] <- 0.384464377736902
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "10_HS"] <- -1.47192637364592

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "10_nHS"]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "10_nHS"]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "10_nHS"] <- 0.268090409008748
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "10_nHS"] <- -1.65250752973503

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "11_HS"]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "11_HS"]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "11_HS"] <- 0.443040676402206
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "11_HS"] <- -1.08301219755084

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "12_HS"]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "12_HS"]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "12_HS"] <- 0.621197743777606
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "12_HS"] <- -1.57565162793579

## Pool2
#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 3]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 3]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 3] <- 0.8323308747399
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 3] <- -1.7976605122722

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 4]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 4]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 4] <- 0.78720107197068
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 4] <- -1.80969795146536

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 5]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 5]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 5] <- 0.544334986599149
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 5] <- -2.44029250693867

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 13]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 13]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 13] <- 0.0448513420823224
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 13] <- -2.11579007203709

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 15]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 15]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 15] <- 0.450573473596564
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 15] <- -1.75256431098751

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 21]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 21]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 21] <- 0.209098707986782
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 21] <- -1.59996704867478

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 22]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 22]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 22] <- 0.0475822795299057
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 22] <- -1.01585824982994 # Pick Structural protein H

## Pool3
#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 14]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 14]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 14] <- 0.429189493024795
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 14] <- -2.69683734770226

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 16]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 16]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 16] <- 0.500506508522238
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 16] <- -2.38944021296657

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 17]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 17]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 17] <- 0.336910551135972
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 17] <- -1.823841305908

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 18]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 18]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 18] <- 0.253625126480594
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 18] <- -1.59637910303852

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 19]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 19]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 19] <- 0.303361366555724
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 19] <- -1.87552032715709

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 20]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 20]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 20] <- 0.348227404410338
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 20] <- -2.10211293851513

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 30]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 30]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 30] <- 0.312778364 # Here only 1 synonymous due to highly overlapping ORFs
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 30] <- -2.55330828879562

## Pool4
#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 1]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 1]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 1] <- 0.378418515511031
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 1] <- -1.60033089699229

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 2]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 2]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 2] <- 0.255742222926761
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 2] <- -1.31021821710037

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 23]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 23]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 23] <- 0.342956691207138
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 23] <- -1.6873930304212

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 24]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 24]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 24] <- 0.281095820419728
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 24] <- -2.84133958763077

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 25]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 25]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 25] <- 0.297589307356069
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 25] <- -3.71186625487644

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 26]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 26]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 26] <- 0.140124152704035
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 26] <- -3.33885525456184

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 27]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 27]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 27] <- 0.367562300490852
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 27] <- -3.35626008472777

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 28]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 28]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 28] <- 0.260060882048925
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 28] <- -1.89441331622316 # Here pick structural protein B

#phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 29]
#phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 29]
phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 29] <- 0.521170330873223
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == 29] <- -1.73452146233666

phix_dms_sub_libraries_mode_summary$synonymous_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "r1"]
phix_dms_sub_libraries_mode_summary$nonsense_mode[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "r1"]
#phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "r1"] <- 0.13254965118588
#phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "r1"] <- -2.41380940559281

phix_dms_sub_libraries_mode_summary$syn_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "r1"] <- -0.0829502911731281
phix_dms_sub_libraries_mode_summary$non_mode_final[phix_dms_sub_libraries_mode_summary$oligo_seq_lib == "r1"] <- -2.46944304455617


# Add a new column to calculate normalization factor for error propagation of sigma
phix_dms_sub_libraries_mode_summary$normalization_factor <- with(
  phix_dms_sub_libraries_mode_summary, 
  ifelse(!is.na(syn_mode_final) & !is.na(non_mode_final) & (syn_mode_final != non_mode_final), 
         abs(1 / (syn_mode_final - non_mode_final)), 
         NA)  # Avoid division by zero
)



```


### 2) Get normalized references for gene H and A 
[First run next stage (3) step (1) - normalize by Syn & Non modes, then come back step (2)]

1) got a unique 1st normalized mode of nonsense mutations for gene H in the library containing both gene H and G - "pool3_twist16_lib16";
  - use the mode of nonsense mutations for gene H to build the 1st rescaling function - 'rescale_geneH_normalization', and rescaling factor - 'rescale_geneH_normalization_factor'
2) rescaled the libraries including protein H by the 1st normalized mode of nonsense mutations and get fitness normalized2;  
  - encountered a library containing both gene H and gene A - "pool2_twist21_lib22", whose 1st normalization is done by the mode of gene H's nonsense mutations;
3) got a different mode of nonsense mutations for gene A across the libraries containing gene A;
  - for "pool2_twist21_lib22", use rescaled normalized fitness - "fitness_normalized2"
  - for the other libraries containing both A and another structural protein, use 1st normalized fitness - "fitness_normalized"
  - combine these fitness data for nonsense mutations of gene A, and get a new mode of nonsense mutations for gene A to build the 2nd rescaling function - 'rescale_geneA_normalization';
4) reload the rescaling functions and normalize in parallel below;

```{r}
# 1) after 1st normalization, and get mode of nonsense mutations for gene H in 'pool3_twist16_lib16'
pool3_twist16_normalized1_nonsense_H <- pool3_twist16_lib16_dms_sub %>%
  filter(gene_sub_aa_mut_gene == "H", gene_sub_aa_mut_class == "Nonsense")

pool3_twist16_normalized1_mode_nonsense_H <- find_modes_from_df(pool3_twist16_normalized1_nonsense_H, column_name = "fitness_normalized")
# [1] "0.214763339263732,0.239109175698378,0.309712101358854"

# Linear regression function based on points (0, 0.2147633) and (1, 1)
rescale_geneH_normalization <- function(x) {
  slope <- (1 - 0.2147633) / (1 - 0)
  intercept <- 0.2147633
  y = slope * x + intercept
  return(y)
}
rescale_geneH_normalization_factor <- abs(slope <- (1 - 0.2147633) / (1 - 0))
rescale_geneH_normalization_factor


# 2) 2nd normalization of rescaling libraries containing only gene H: Twist 21, 22, 17, 18, 19, 20
# ---- List of data.frame names ----
geneH_dfs <- c(
  "pool3_twist17_lib17_dms_sub",
  "pool3_twist18_lib18_dms_sub",
  "pool3_twist19_lib19_dms_sub",
  "pool3_twist20_1_lib20_dms_sub",
  "pool2_twist20_lib21_dms_sub",
  "pool2_twist21_lib22_dms_sub"
)

# ---- Apply normalization ----
for (df_name in geneH_dfs) {
  
  df <- get(df_name)
  
  df <- df %>%
    dplyr::mutate(
      fitness_normalized2 =
        rescale_geneH_normalization(fitness_normalized),
      
      sigma_rescaled2 =
        sigma_rescaled *
        rescale_geneH_normalization_factor
    )
  
  assign(df_name, df, envir = .GlobalEnv)
  
  cat("Normalized:", df_name, "\n")
}


# 3) 3rd normalization of rescaling libraries containing gene A: Twist 28, 29, 30, 1, r1
# ---- Data.frames to use with fitness_normalized ----
other_dfs <- c(
  "pool3_twist29_lib30_dms_sub",
  "pool4_twist1_lib1_dms_sub",
  "pool4_twist27_lib28_dms_sub",
  "pool4_twist28_lib29_dms_sub",
  "pool4_twist30_libr1_dms_sub"
)

# ---- Extract lib22 nonsense mutations from gene A ----
pool2_twist21_lib22_nonsense_A <- pool2_twist21_lib22_dms_sub %>%
  dplyr::filter(
    gene_sub_aa_mut3_gene == "A",
    #gene_sub_aa_mut3_class == "Nonsense",
    stringr::str_detect(
      gene_sub_aa_mut3_class,
      "Nonsense"
    )
  ) %>%
  dplyr::mutate(
    fitness_combined = fitness_normalized2
  )

# ---- Extract from other data.frames ----
other_nonsense_A_list <- lapply(other_dfs, function(df_name) {
  
  df <- get(df_name)
  
  df %>%
    dplyr::filter(
      gene_sub_aa_mut3_gene == "A",
      #gene_sub_aa_mut3_class == "Nonsense",
      stringr::str_detect(
        gene_sub_aa_mut3_class,
        "Nonsense"
      )
    ) %>%
    dplyr::mutate(
      fitness_combined = fitness_normalized
    )
})

# ---- Combine all data.frames ----
combined_nonsense_A <- dplyr::bind_rows(
  c(
    list(pool2_twist21_lib22_nonsense_A),
    other_nonsense_A_list
  )
)

# ---- Remove NA values ----
#combined_nonsense_A <- combined_nonsense_A %>%
  #dplyr::filter(!is.na(fitness_combined))

# ---- Find mode using your function ----
combined_nonsense_A_mode <- find_modes_from_df(
  combined_nonsense_A,
  column_name = "fitness_combined"
)

combined_nonsense_A_mode
# [1] "-0.641869598618608,-0.267754460820804,-0.0170973184962746"

# print(combined_nonsense_A_mode_new) # [1] -0.641869598618608


# Linear regression function based on points (0, -0.638541479) and (1, 1)
rescale_geneA_normalization <- function(x) {
  slope <- (1 - (-0.6418696)) / (1 - 0)
  intercept <- -0.6418696
  y = slope * x + intercept
  return(y)
}
rescale_geneA_normalization_factor <- abs((1 - (-0.6418696)) / (1 - 0))
rescale_geneA_normalization_factor


# Function to apply fitness normalization conditions
apply_fitness_and_sigma_normalization2 <- function(df_name) {
  df <- get(df_name)  # Retrieve the data frame by name
  
  # Ensure the required columns exist
  if (!all(c("fitness_normalized", "sigma_rescaled", "oligo_seq_lib") %in% names(df))) {
    warning(paste("Skipping", df_name, "- Missing required columns"))
    return(NULL)
  }
  
  # Assign fitness_normalized2 as default (same as fitness_normalized)
  df$fitness_normalized2 <- df$fitness_normalized
  df$sigma_rescaled2 <- df$sigma_rescaled  # Default to sigma_rescaled

  # Determine normalization based on oligo_seq_lib values
  if (df$oligo_seq_lib[1] %in% c(6, 7, 8, 9, 10, 11, 12, 3, 4, 5, 13, 15, 14, 16, 1, 2, 28, 29, "r1")) {
    df$fitness_normalized2 <- df$fitness_normalized  # Keep same values
    df$sigma_rescaled2 <- df$sigma_rescaled  # No further scaling
  } else if (df$oligo_seq_lib[1] %in% c(21, 22, 17, 18, 19, 20)) {
    df$fitness_normalized2 <- rescale_geneH_normalization(df$fitness_normalized)
    df$sigma_rescaled2 <- rescale_geneH_normalization_factor * df$sigma_rescaled
  } else if (df$oligo_seq_lib[1] %in% c(23, 24, 25, 26, 27)) {
    df$fitness_normalized2 <- rescale_geneA_normalization(df$fitness_normalized)
    df$sigma_rescaled2 <- rescale_geneA_normalization_factor * df$sigma_rescaled
  }

  # Assign updated data frame back to the global environment
  assign(df_name, df, envir = .GlobalEnv)
  cat("Normalization (Step 2) complete for:", df_name, "\n")
}



```


### 3) Normalize Substitutions data
Normalize all substitutions data; 
1) Syn&Non normalization; 
2) Rescaling Gene H's and Gene A's;

```{r}
# Step 1) Normalize each library by it's own Syn&Non mode; get column fitness_normalized and sigma_rescaled
# Apply normalization for all merged data frames
for (df_name in merged_dms_sub_libraries) {
  normalize_fitness_and_sigma(df_name, phix_dms_sub_libraries_mode_summary)
}

### after step (1), go back to stage (2) and prepare data and functions for downstream normalization ### 


# Step 2) Rescale Gene H's data; get column fitness_normalized2 and sigma_rescaled2;
# Apply the function to all merged data frames
for (df_name in merged_dms_sub_libraries) {
  apply_fitness_and_sigma_normalization2(df_name)
}


# Step 3) Save the data.frames with normalized fitness
for (df_name in merged_dms_sub_libraries) {
  new_df_name <- paste0(df_name, "_normalized")  # Append "_normalized" to the new name
  assign(new_df_name, get(df_name), envir = .GlobalEnv)  # Copy the data frame
  cat("Copied", df_name, "to", new_df_name, "\n")
}


## List all the Substitutions data normalized (Normal Selection)
phix_merged_dms_sub_libraries_normalized <- c(
  "pool1_twist5_lib6_dms_sub_normalized",
  "pool1_twist6_lib7_dms_sub_normalized",
  "pool1_twist7_lib8_dms_sub_normalized",
  "pool1_twist8_lib9_dms_sub_normalized",
  "pool1_twist9_lib10_dms_sub_normalized",
  "pool1_twist10_lib11_dms_sub_normalized",
  "pool1_twist11_lib12_dms_sub_normalized",
  
  "pool2_twist3_lib3_dms_sub_normalized",
  "pool2_twist4_lib4_dms_sub_normalized",
  "pool2_twist5_lib5_dms_sub_normalized",
  "pool2_twist13_lib13_dms_sub_normalized",
  "pool2_twist15_lib15_dms_sub_normalized",
  "pool2_twist20_lib21_dms_sub_normalized",
  "pool2_twist21_lib22_dms_sub_normalized",
  
  "pool3_twist14_lib14_dms_sub_normalized",
  "pool3_twist16_lib16_dms_sub_normalized",
  "pool3_twist17_lib17_dms_sub_normalized",
  "pool3_twist18_lib18_dms_sub_normalized",
  "pool3_twist19_lib19_dms_sub_normalized",
  "pool3_twist20_1_lib20_dms_sub_normalized",
  "pool3_twist29_lib30_dms_sub_normalized",
  
  "pool4_twist1_lib1_dms_sub_normalized",
  "pool4_twist2_lib2_dms_sub_normalized",
  "pool4_twist22_lib23_dms_sub_normalized",
  "pool4_twist23_lib24_dms_sub_normalized",
  "pool4_twist24_lib25_dms_sub_normalized",
  "pool4_twist25_lib26_dms_sub_normalized",
  "pool4_twist26_lib27_dms_sub_normalized",
  "pool4_twist27_lib28_dms_sub_normalized",
  "pool4_twist28_lib29_dms_sub_normalized",
  
  "pool4_twist30_libr1_dms_sub_normalized"
  )


# save filtered raw fitness data
phix_dms_sub_libraries_normalized <- mget(
  phix_merged_dms_sub_libraries_normalized,
  envir = .GlobalEnv
)

saveRDS(
  phix_dms_sub_libraries_normalized,
  file = "./data/raw_data/phix_dms_sub_libraries_normalized.rds",
  compress = "xz"
)


# load normalized data
sub_libs <- readRDS("./data/raw_data/phix_dms_sub_libraries_normalized.rds")
list2env(sub_libs, envir = .GlobalEnv)




```


## Save script checkpoint
save a snapshot of all objects produced by the script

```{r}
# checkpoint - 01 within-library normalization
# save.image(file = "data/checkpoints/01_fitness_data_normalization.RData")


```


