---
title: "phix_DMS_all_variants_data_analysis"
author: "Huijin Wei"
date: "2026-07-30"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Set working working directory
```{r}
here::here()

```



## Load functions 
### 1) substitutions data analysis
```{r}
# Function to calculate weighted mean and propagated error
# 1) For library-overlapping data
calculate_weighted_mean_and_error <- function(df, col1, column_fitness, column_sigma,
                                              weighted_mean_col, propagated_error_col) {
  df %>%
    group_by_at(col1) %>%
    summarise(
      across(everything(), ~ first(.), .names = "{.col}"),  # Keep metadata from first row
      !!weighted_mean_col := sum((!!sym(column_fitness)) / (!!sym(column_sigma))^2, na.rm = TRUE) /
                              sum(1 / (!!sym(column_sigma))^2, na.rm = TRUE),
      !!propagated_error_col := sqrt(1 / sum(1 / (!!sym(column_sigma))^2, na.rm = TRUE)),
      .groups = "drop"
    )
}

# Function to calculate Weighted.mean and pass error 
# 2) For getting fitness data of unique AA mut
calculate_unique_weighted_mean_and_error <- function(df, col1, column_fitness, column_sigma, 
                                              weighted_mean_col, propagated_error_col) {
  df %>%
    group_by_at(col1) %>%
    mutate(
      row_count = n(),  # Count the number of rows per group
      
      # Weighted mean for fitness
      !!weighted_mean_col := ifelse(
        row_count == 1, !!sym(column_fitness),  # If only one row, keep original value
        sum((!!sym(column_fitness)) / (!!sym(column_sigma))^2, na.rm = TRUE) / 
        sum(1 / (!!sym(column_sigma))^2, na.rm = TRUE)
      ),
      
      # Propagated error
      !!propagated_error_col := ifelse(
        row_count == 1, !!sym(column_sigma),  # If only one row, keep original sigma
        sqrt(1 / sum(1 / (!!sym(column_sigma))^2, na.rm = TRUE))
      )
    ) %>%
    select(-row_count) %>%  # Remove temporary column
    ungroup()
}

# Function to supplement annotations at genome nt level based on the position 1 of mutations
update_genome_nt_positions <- function(df, pos_col = "genome_nt_mut_pos1") {
  
  # Initialize columns with 0 (No)
  df <- df %>%
    mutate(
      genome_nt_pos_noncoding = NA,
      genome_nt_pos_coding = NA,
      genome_nt_pos_overlap_orf = NA,
      genome_nt_pos_structural = NA
    )
  
  # Define position ranges for each classification
  noncoding_ranges <- list(
    c(965, 1000), c(2285, 2394), c(2923, 2930),
    c(3918, 3980)
  )
  
  coding_ranges <- list(
    c(1, 964), c(1001, 2284), c(2395, 2922),
    c(2931, 3917), c(3981, 5386)
  )
  
  overlap_orf_ranges <- list(
    c(1, 221), c(390, 393), c(568, 843),
    c(848, 848), c(5075, 5386)
  )
  
  structural_ranges <- list(
    c(1, 51), c(390, 964), c(1001, 2284), 
    c(2395, 2922), c(5075, 5386)
  )
  
  # Function to check if a value falls within any of the specified ranges
  is_in_ranges <- function(value, ranges) {
    any(sapply(ranges, function(r) value >= r[1] & value <= r[2]))
  }
  
  # Apply the range checks to update the binary columns
  df <- df %>%
    mutate(
      genome_nt_pos_noncoding = ifelse(sapply(.data[[pos_col]], is_in_ranges, noncoding_ranges), "Noncoding", NA),
      genome_nt_pos_coding = ifelse(sapply(.data[[pos_col]], is_in_ranges, coding_ranges), "Coding", NA),
      genome_nt_pos_overlap_orf = ifelse(sapply(.data[[pos_col]], is_in_ranges, overlap_orf_ranges), "Overlapping Coding", NA),
      genome_nt_pos_structural = ifelse(sapply(.data[[pos_col]], is_in_ranges, structural_ranges), T, F)
    )
  
  return(df)
}


```


### 2) statistics and mutation category
```{r}
## Function to compute p-value for Statitic tests
# Two-samples t-test
two_samples_ttest <- function(mu1, sem1, n1, mu2, sem2, n2) {
  # Compute t-statistic using SEM directly
  t_stat <- (mu1 - mu2) / sqrt(sem1^2 + sem2^2)

  # Compute degrees of freedom (Welch’s approximation)
  df <- (sem1^2 + sem2^2)^2 /
        ((sem1^4 / (n1 - 1)) + (sem2^4 / (n2 - 1)))

  # Ensure df is valid
  if (is.na(df) || df <= 0) {
    df <- 1  # Prevent invalid df values
  }

  # Compute two-tailed p-value
  p_val <- 2 * (1 - pt(abs(t_stat), df))

  return(p_val)
}



# Function for mutation category and direction, using Fitness-increasing rather Beneficial 
assign_mutation_categories_fdr <- function(
  df,
  pval_col,
  effect_raw_col,
  alpha = 0.05,
  effect_breaks = c(0.10, 0.15, 0.50),

  effect_bin_col,
  category_fdr_col,
  category_fdr_lethal_col,
  category_cutoff_col,
  category_combined_col,

  direction_col,
  direction_lethal_col
) {

  # ---- Checks ----
  required_cols <- c(pval_col, effect_raw_col)
  missing_cols <- setdiff(required_cols, names(df))

  if (length(missing_cols) > 0) {
    stop("Missing required columns: ", paste(missing_cols, collapse = ", "))
  }

  if (!is.numeric(effect_breaks)) {
    stop("effect_breaks must be numeric.")
  }

  if (length(effect_breaks) != 3) {
    stop("effect_breaks should contain three values, e.g. c(0.10, 0.15, 0.50).")
  }

  effect_breaks <- sort(effect_breaks)

  # ---- Extract clean numeric vectors ----
  p <- as.numeric(df[[pval_col]])
  e <- as.numeric(df[[effect_raw_col]])

  # ---- Significance ----
  sig_flag <- !is.na(p) & (p < alpha)

  # ---- Base classification ----
  # Original category labels are kept:
  # Beneficial / Neutral / Detrimental
  category_fdr <- dplyr::case_when(
    is.na(p) | is.na(e) ~ NA_character_,
    !sig_flag ~ "Neutral",
    e > 0 ~ "Beneficial",
    e < 0 ~ "Detrimental",
    TRUE ~ "Neutral"
  )

  # ---- Effect bins ----
  effect_bin_num <- dplyr::case_when(
    is.na(e) ~ NA_real_,
    !sig_flag ~ 0,
    abs(e) >= effect_breaks[3] ~ effect_breaks[3],
    abs(e) >= effect_breaks[2] ~ effect_breaks[2],
    abs(e) >= effect_breaks[1] ~ effect_breaks[1],
    TRUE ~ 0
  )

  # ---- Lethal override ----
  category_fdr_lethal <- dplyr::case_when(
    is.na(category_fdr) ~ NA_character_,
    category_fdr == "Detrimental" &
      effect_bin_num >= effect_breaks[3] ~ "Lethal",
    TRUE ~ category_fdr
  )

  # ---- Cutoff refinement ----
  # Original 0.15 cutoff refinement is kept.
  category_cutoff <- dplyr::case_when(
    is.na(category_fdr) ~ NA_character_,
    category_fdr %in% c("Beneficial", "Detrimental") &
      effect_bin_num < effect_breaks[2] ~ "Neutral",
    category_fdr == "Detrimental" &
      effect_bin_num >= effect_breaks[3] ~ "Lethal",
    TRUE ~ category_fdr
  )

  # ---- Combined ----
  category_combined <- dplyr::case_when(
    category_cutoff %in% c("Beneficial", "Neutral") ~ "Tolerated",
    category_cutoff %in% c("Detrimental", "Lethal") ~ "Deleterious",
    TRUE ~ NA_character_
  )

  # ---- Direction column without lethal ----
  # Same logic as category_fdr, but with less interpretive labels.
  direction <- dplyr::case_when(
    is.na(category_fdr) ~ NA_character_,
    category_fdr == "Beneficial" ~ "Fitness-increasing",
    category_fdr == "Neutral" ~ "Neutral",
    category_fdr == "Detrimental" ~ "Fitness-decreasing",
    TRUE ~ NA_character_
  )

  # ---- Direction column with lethal ----
  # Same logic as category_fdr_lethal, but with less interpretive labels.
  direction_lethal <- dplyr::case_when(
    is.na(category_fdr_lethal) ~ NA_character_,
    category_fdr_lethal == "Beneficial" ~ "Fitness-increasing",
    category_fdr_lethal == "Neutral" ~ "Neutral",
    category_fdr_lethal == "Detrimental" ~ "Fitness-decreasing",
    category_fdr_lethal == "Lethal" ~ "Lethal",
    TRUE ~ NA_character_
  )

  # ---- Safe overwrite ----
  output_cols <- c(
    effect_bin_col,
    category_fdr_col,
    category_fdr_lethal_col,
    category_cutoff_col,
    category_combined_col,
    direction_col,
    direction_lethal_col
  )

  for (col in output_cols) {
    df[[col]] <- NULL
  }

  # ---- Assign outputs explicitly ----
  df[[effect_bin_col]] <- factor(
    effect_bin_num,
    levels = c(0, effect_breaks)
  )

  df[[category_fdr_col]] <- factor(
    category_fdr,
    levels = c("Beneficial", "Neutral", "Detrimental")
  )

  df[[category_fdr_lethal_col]] <- factor(
    category_fdr_lethal,
    levels = c("Beneficial", "Neutral", "Detrimental", "Lethal")
  )

  df[[category_cutoff_col]] <- factor(
    category_cutoff,
    levels = c("Beneficial", "Neutral", "Detrimental", "Lethal")
  )

  df[[category_combined_col]] <- factor(
    category_combined,
    levels = c("Tolerated", "Deleterious")
  )

  df[[direction_col]] <- factor(
    direction,
    levels = c("Fitness-increasing", "Neutral", "Fitness-decreasing")
  )

  df[[direction_lethal_col]] <- factor(
    direction_lethal,
    levels = c(
      "Fitness-increasing",
      "Neutral",
      "Fitness-decreasing",
      "Lethal"
    )
  )

  return(df)
}



# Function to generate summary tables
summarize_counts_df <- function(df, group_cols, count_col = NULL) {
  group_syms <- syms(group_cols)
  count_sym  <- if (!is.null(count_col)) sym(count_col) else NULL

  summary_df <- if (is.null(count_col)) {
    df %>%
      group_by(!!!group_syms) %>%
      summarise(n = n(), .groups = "drop")
  } else {
    df %>%
      group_by(!!!group_syms, !!count_sym) %>%
      summarise(n = n(), .groups = "drop")
  }

  return(as.data.frame(summary_df))
}




```


### 3) summary and plots 
```{r}
# Function to generate summary table
generate_mutation_category_summary <- function(
    df, 
    class_col,
    category_col,
    total_count = NULL
) {

  class_syms    <- rlang::syms(class_col)
  category_sym  <- rlang::ensym(category_col)

  df %>%
    dplyr::filter(!is.na(!!category_sym)) %>%
    dplyr::group_by(!!!class_syms, !!category_sym) %>%
    dplyr::summarise(count = dplyr::n(), .groups = "drop") %>%
    dplyr::group_by(!!!class_syms) %>%
    dplyr::mutate(
      proportion = if (is.null(total_count)) {
        count / sum(count)
      } else {
        count / total_count
      },
      percent = round(proportion * 100, 2)
    ) %>%
    dplyr::ungroup()
}


# Function to plot the correlation across bio-replicates
plot_replicate_reproducibility <- function(df,
                                           replicate_cols,
                                           cor_method = "pearson",
                                           point_alpha = 0.5,
                                           point_size = 1.2,
                                           title = "Reproducibility Across Replicates") {
  # ---- Check inputs ----
  stopifnot(all(replicate_cols %in% names(df)))
  
  # ---- Load required library ----
  if (!requireNamespace("GGally", quietly = TRUE)) {
    stop("Please install 'GGally' package: install.packages('GGally')")
  }

  # ---- Subset and clean data ----
  df_sub <- df[, replicate_cols]
  df_sub <- df_sub[stats::complete.cases(df_sub), ]

  # ---- Build ggpairs plot ----
  p <- ggpairs(
    df_sub,
    upper = list(continuous = wrap("cor",
                                           size = 4,
                                           method = cor_method,
                                           use = "pairwise.complete.obs",
                                           color = "black")),
    lower = list(continuous = wrap("points",
                                           alpha = point_alpha,
                                           size = point_size,
                                           color = "black")),
                                           #color = "gray88")),
    diag = list(continuous = wrap("densityDiag",
                                          alpha = 0.6,
                                          fill = "gray88"))
  ) +
    theme_classic() +
    theme(
      strip.text = element_text(size = 11, face = "bold"),
      plot.title = element_text(size = 14, face = "bold", hjust = 0.5)
    ) +
    ggtitle(title)

  return(p)
}

# Function - Pir charts
plot_pie_summary <- function(data, 
                              x_col = "Var1", 
                              y_col = "Freq", 
                              plot_title = NULL,
                              outline_color = "black",
                              outline_linewidth = 0.6,
                              palette = "Set3",
                              remove_legend = FALSE,
                              show_text = TRUE) {    
  
  x_sym <- rlang::sym(x_col)
  y_sym <- rlang::sym(y_col)

  # Prepare and order the data by descending y_col
  data <- data %>%
    mutate(
      percentage = round(100 * !!y_sym / sum(!!y_sym), 1),
      legend_label = paste0(!!x_sym, " (", percentage, "%)")
    ) %>%
    arrange(desc(!!y_sym)) %>%
    mutate(
      legend_label = factor(legend_label, levels = unique(legend_label)),  # Order legend
      fraction = !!y_sym / sum(!!y_sym),
      ymax = cumsum(fraction),
      ymin = c(0, head(ymax, -1))
    )

  p <- ggplot(data) +
    geom_rect(
      aes(ymin = ymin, ymax = ymax, xmin = 0.5, xmax = 1.5, fill = legend_label),
      color = outline_color,
      linewidth = outline_linewidth
    ) +
    coord_polar(theta = "y") +
    xlim(0.5, 2.2) +
    theme_void() +
    labs(title = plot_title, fill = NULL) +
    scale_fill_brewer(palette = palette) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold"),
      legend.title = element_blank(),
      legend.text = element_text(size = 10)
    )
  
  # Add text if requested
  if (show_text) {
    p <- p + geom_text(
      aes(x = 1, y = (ymin + ymax) / 2, label = paste0(percentage, "%")),
      size = 5,
      color = "black"
    )
  }

  # Optionally remove legend
  if (remove_legend) {
    p <- p + theme(legend.position = "none")
  }
  
  return(p)
}


# Function to plot ridgeline plot
plot_ridgeline_mut_class <- function(data,
                                     score_col = fitness_final,
                                     class_col = phix_aa_mut_class,
                                     plot_title = "Ridgeline Density Plot of Fitness Final",
                                     x_label = "Fitness") {
  
  p <- ggplot(data, aes(x = {{ score_col }}, y = {{ class_col }})) +
    geom_density_ridges(
      scale = 1.2,
      fill = "grey88",
      color = "black",
      alpha = 0.8
    ) +
    xlim(-1, 1.5) +
    labs(
      title = plot_title,
      x = x_label,
      y = "Mutation Class"
    ) +
    theme_classic() +
    theme(
      strip.text = element_text(face = "bold"),
      axis.text.y = element_text(face = "bold")
    )
  
  return(p)
}



```



## Load data
Load normalized fitness data in 31 libraries (data.frames)
```{r}
# load normalized data
sub_libs <- readRDS("./data/raw_data/phix_dms_sub_libraries_normalized.rds")
list2env(sub_libs, envir = .GlobalEnv)


```



## Deduplicate genotype (Unique Genotypes)
1) combine all 31 libraries with normalized fitness data;
2) for duplicate variants, get weighted mean for the same genotype;
```{r}
# list all libraries (Only Substitutions)
phix_dms_sub_libraries_normalized <- c(
  "pool1_twist5_lib6_dms_sub_normalized",
  "pool1_twist6_lib7_dms_sub_normalized",
  "pool1_twist7_lib8_dms_sub_normalized",
  "pool1_twist8_lib9_dms_sub_normalized",
  "pool1_twist9_lib10_dms_sub_normalized",
  "pool1_twist10_lib11_dms_sub_normalized",
  "pool1_twist11_lib12_dms_sub_normalized",
  
  "pool2_twist3_lib3_dms_sub_normalized",
  "pool2_twist4_lib4_dms_sub_normalized",
  "pool2_twist5_lib5_dms_sub_normalized",
  "pool2_twist13_lib13_dms_sub_normalized",
  "pool2_twist15_lib15_dms_sub_normalized",
  "pool2_twist20_lib21_dms_sub_normalized",
  "pool2_twist21_lib22_dms_sub_normalized",
  
  "pool3_twist14_lib14_dms_sub_normalized",
  "pool3_twist16_lib16_dms_sub_normalized",
  "pool3_twist17_lib17_dms_sub_normalized",
  "pool3_twist18_lib18_dms_sub_normalized",
  "pool3_twist19_lib19_dms_sub_normalized",
  "pool3_twist20_1_lib20_dms_sub_normalized",
  "pool3_twist29_lib30_dms_sub_normalized",
  
  "pool4_twist1_lib1_dms_sub_normalized",
  "pool4_twist2_lib2_dms_sub_normalized",
  "pool4_twist22_lib23_dms_sub_normalized",
  "pool4_twist23_lib24_dms_sub_normalized",
  "pool4_twist24_lib25_dms_sub_normalized",
  "pool4_twist25_lib26_dms_sub_normalized",
  "pool4_twist26_lib27_dms_sub_normalized",
  "pool4_twist27_lib28_dms_sub_normalized",
  "pool4_twist28_lib29_dms_sub_normalized",
  
  "pool4_twist30_libr1_dms_sub_normalized"
)

# 1) combine 
phix_all_dms_sub_libraries_normalized <- do.call(rbind, mget(phix_dms_sub_libraries_normalized))
phix_all_dms_sub_libraries_normalized <- phix_all_dms_sub_libraries_normalized[order(phix_all_dms_sub_libraries_normalized$genome_nt_mut_pos1),]
nrow(phix_all_dms_sub_libraries_normalized)
# [1] 50906x

table(duplicated(phix_all_dms_sub_libraries_normalized$genome_nt_mut_seq)) 
# FALSE  TRUE 
# 44195  6711 


# 2) deduplicate by phix genotypes
phix_all_dms_sub_libraries_normalized_dedup <- calculate_weighted_mean_and_error(
  phix_all_dms_sub_libraries_normalized,
  "genome_mut_ID_nt",
  "fitness_normalized2",
  "sigma_rescaled2",
  "fitness_final",
  "sigma_final")

table(duplicated(phix_all_dms_sub_libraries_normalized_dedup$genome_nt_mut_seq))
# FALSE 
# 44195 

## Unique genotypes: 44,195


# 3) add genomic annotations
# Remove oligo information 
phix_all_dms_sub_libraries_normalized_unique <- phix_all_dms_sub_libraries_normalized_dedup %>%
  select(-starts_with("oligo_"))

# add annotations by nucleotide residues
phix_all_dms_sub_libraries_normalized_unique$phix_aa_mut_class[is.na(phix_all_dms_sub_libraries_normalized_unique$phix_aa_mut_class)] <- "Noncoding"
phix_all_dms_sub_libraries_normalized_unique <- phix_all_dms_sub_libraries_normalized_unique %>%
  mutate(genome_nt_mut_pos1 = as.numeric(genome_nt_mut_pos1)) %>%
  mutate(
    phix_aa_mut_genes = case_when(
      genome_nt_mut_pos1 %in% 965:1000  ~ "Noncoding#1 J-F",
      genome_nt_mut_pos1 %in% 2285:2394 ~ "Noncoding#2 F-G",
      genome_nt_mut_pos1 %in% 2923:2930 ~ "Noncoding#3 G-H",
      genome_nt_mut_pos1 %in% 3918:3980 ~ "Noncoding#4 H-A",
      TRUE ~ phix_aa_mut_genes
    )
  )

# supplement nt annotations: Coding, protein-Coding, Overlapping Coding, and Structural proteins
phix_all_dms_sub_libraries_normalized_unique <- 
  update_genome_nt_positions(phix_all_dms_sub_libraries_normalized_unique, "genome_nt_mut_pos1")

# Correct 'ReadThrough' to 'Stop-loss'
phix_all_dms_sub_libraries_normalized_unique <- 
  phix_all_dms_sub_libraries_normalized_unique %>%
  dplyr::mutate(
    dplyr::across(
      where(is.character),
      ~ stringr::str_replace_all(.x, "ReadThrough", "Stop-loss")
    )
  )



```



## Statistics
t-test and FDR: two-sample t-test, compare to Synonymous distribution as the reference (for 1nt variants, separately);

1) 1nt Data
  Syn final fitness: phix_1nt_synonymous_combined_fitness_final
  Syn final sigma: phix_1nt_synonymous_combined_sigma_final
2) AA Data
  Syn final fitness: phix_all_synonymous_combined_fitness_final
  Syn final sigma: phix_all_synonymous_combined_sigma_final

### 1) Get Reference data (Synonymous)
Filter the synonymous fitness (normalized final) data, and calculate weighted.mean and propagate the error of SEM and make sure data is filtered correctly and remove NA values

```{r}
## All variants
filtered_data_all_synonymous <- phix_all_dms_sub_libraries_normalized_unique %>%
  filter(phix_aa_mut_class == "Synonymous", !is.na(fitness_final))

# get weighted.mean of all synonymous mutations;
phix_all_synonymous_combined_weights <- 1 / (filtered_data_all_synonymous$sigma_final^2)  # Weights are the inverse of variance
phix_all_synonymous_combined_fitness_final <- sum(filtered_data_all_synonymous$fitness_final * phix_all_synonymous_combined_weights) / sum(phix_all_synonymous_combined_weights) 
# [1] 0.9807827
phix_all_synonymous_combined_sigma_final <- sqrt(1/sum(phix_all_synonymous_combined_weights)) 
# [1] 0.0002573693



## 1nt variants
filtered_data_1nt_synonymous <- phix_all_dms_sub_libraries_normalized_unique %>%
  filter(genome_nt_mut_ham == 1, phix_aa_mut_class == "Synonymous", !is.na(fitness_final))

# get weighted.mean of all 1nt synonymous mutations;
phix_1nt_synonymous_combined_weights <- 1 / (filtered_data_1nt_synonymous$sigma_final^2)  # Weights are the inverse of variance
phix_1nt_synonymous_combined_fitness_final <- sum(filtered_data_1nt_synonymous$fitness_final * phix_1nt_synonymous_combined_weights) / sum(phix_1nt_synonymous_combined_weights) 
# [1] 0.978221
phix_1nt_synonymous_combined_sigma_final <- sqrt(1/sum(phix_1nt_synonymous_combined_weights)) # Sigma final = 0.0002678072, 0.0002678328, 0.0002812053
# [1] 0.0002813241



```


### 2) two-sample t-tests

```{r}
# Make a copy for statistics
phix_all_dms_sub_libraries_normalized_unique_stat <- phix_all_dms_sub_libraries_normalized_unique


# Apply the function row-wise, the sample size is related to how many nt muts are merged to one AA mut 
phix_all_dms_sub_libraries_normalized_unique_stat$p_value_two_sample_ttest_syn <- mapply(
  two_samples_ttest,
  phix_all_dms_sub_libraries_normalized_unique_stat$fitness_final,  # mu1 (row-wise)
  phix_all_dms_sub_libraries_normalized_unique_stat$sigma_final,    # sem1 (row-wise)
  MoreArgs = list(
    n1 = 3,  # Fixed (n1),
    mu2 = phix_all_synonymous_combined_fitness_final,  # Fixed (mu2)
    sem2 = phix_all_synonymous_combined_sigma_final,   # Fixed (sem2)
    n2 = 3  # Fixed (n2)
  )
)


# Apply FDR correction using the function p.adjust() from basic R;
phix_all_dms_sub_libraries_normalized_unique_stat <- phix_all_dms_sub_libraries_normalized_unique_stat %>%
  mutate(p_value_two_sample_ttest_syn_FDR = p.adjust(p_value_two_sample_ttest_syn, method = "BH"))


# Add raw effect size (difference from the synonymous baseline)
phix_all_dms_sub_libraries_normalized_unique_stat$effect_size_vs_syn <-
  phix_all_dms_sub_libraries_normalized_unique_stat$fitness_final -
  phix_all_synonymous_combined_fitness_final


# Add standardized effect (signal-to-noise; same scaling as t-stat numerator)
phix_all_dms_sub_libraries_normalized_unique_stat$effect_size_std_vs_syn <-
  (phix_all_dms_sub_libraries_normalized_unique_stat$fitness_final -
   phix_all_synonymous_combined_fitness_final) /
  sqrt(phix_all_dms_sub_libraries_normalized_unique_stat$sigma_final^2 +
       phix_all_synonymous_combined_sigma_final^2)



```


### 3) Mutation Category
FUNCTION: assign_mutation_categories_fdr
  
```{r}
# Running Mutation Category
phix_all_dms_sub_libraries_normalized_unique_stat <- assign_mutation_categories_fdr(
  phix_all_dms_sub_libraries_normalized_unique_stat,
  pval_col = "p_value_two_sample_ttest_syn_FDR",
  effect_raw_col = "effect_size_vs_syn",
  
  alpha = 0.05,
  effect_breaks = c(0.10, 0.15, 0.50),

  # Output column names
  effect_bin_col = "effect_size_bin",
  category_fdr_col = "mutation_category_FDR",
  category_fdr_lethal_col = "mutation_category_FDR_lethal",
  category_cutoff_col = "mutation_category_FDR_cutoff",
  category_combined_col = "mutation_category_FDR_cutoff_combined",
  
  direction_col = "mutation_direction",
  direction_lethal_col = "mutation_direction_lethal"
)

table(phix_all_dms_sub_libraries_normalized_unique_stat$mutation_direction) 
# Fitness-increasing            Neutral Fitness-decreasing 
#                533              14491              29171 

table(phix_all_dms_sub_libraries_normalized_unique_stat$mutation_direction_lethal)
# Fitness-increasing            Neutral Fitness-decreasing             Lethal 
#                533              14491               9118              20053 

table(phix_all_dms_sub_libraries_normalized_unique_stat$mutation_category_FDR) 
# Beneficial     Neutral Detrimental
#        533       14491       29171 

table(phix_all_dms_sub_libraries_normalized_unique_stat$mutation_category_FDR_lethal)
# Beneficial     Neutral Detrimental      Lethal 
#        533       14491        9118       20053 


```


### 4) Complete nucleotides annotations
separate noncoding annotations to the all sub data data.frame
- RBS, Promoters, Terminators, Ori+ ... 

```{r}
# Supplement annotations
phix_all_dms_sub_libraries_normalized_unique_stat <-
  phix_all_dms_sub_libraries_normalized_unique_stat %>%
  
  dplyr::mutate(
    
    # ---- 1) nucleotide class priority ----
    genome_nt_pos_nucleotide_class = purrr::pmap_chr(
      list(genome_nt_pos_overlap_orf, genome_nt_pos_coding, genome_nt_pos_noncoding),
      ~ {
        if (!is.na(..1)) {
          as.character(..1)
        } else if (!is.na(..2)) {
          as.character(..2)
        } else if (!is.na(..3)) {
          as.character(..3)
        } else {
          NA_character_
        }
      }
    )
  ) %>%

  # ---- RBS ----
  dplyr::left_join(
    tibble::tibble(
      genome_nt_mut_pos1 = c(
        5063:5068, 39:45, 124:130, 378:381, 555:558,
        836:840, 991:994, 2383:2387, 2923:2928, 3969:3973
      ),
      genome_nt_pos_RBS = "RBS",
      genome_nt_pos_RBS_gene = c(
        rep("RBS (B)", length(5063:5068)),
        rep("RBS (K)", length(39:45)),
        rep("RBS (C)", length(124:130)),
        rep("RBS (D)", length(378:381)),
        rep("RBS (E)", length(555:558)),
        rep("RBS (J)", length(836:840)),
        rep("RBS (F)", length(991:994)),
        rep("RBS (G)", length(2383:2387)),
        rep("RBS (H)", length(2923:2928)),
        rep("RBS (A)", length(3969:3973))
      )
    ),
    by = "genome_nt_mut_pos1"
  ) %>%

  # ---- promoter ----
  dplyr::left_join(
    tibble::tibble(
      genome_nt_mut_pos1 = c(3927:3955, 129:155, 4863:4892, 313:357),
      genome_nt_pos_promoter = "Promoter",
      genome_nt_pos_promoter_gene = c(
        rep("pA", length(3927:3955)),
        rep("pD2 (btss49)", length(129:155)),
        rep("pB", length(4863:4892)),
        rep("pD", length(313:357))
      )
    ),
    by = "genome_nt_mut_pos1"
  ) %>%

  # ---- anti-sense promoter ----
  dplyr::left_join(
    tibble::tibble(
      genome_nt_mut_pos1 = c(2753:2779, 4021:4052),
      genome_nt_pos_promoter_anti_sense = c(
        rep("pR3", length(2753:2779)),
        rep("pR12", length(4021:4052))
      )
    ),
    by = "genome_nt_mut_pos1"
  ) %>%

  # ---- terminator ----
  dplyr::left_join(
    tibble::tibble(
      genome_nt_mut_pos1 = c(3956:3974, 2909:2942, 5160:5238, 2308:2351, 969:985),
      genome_nt_pos_terminator = "Terminator",
      genome_nt_pos_terminator_gene = c(
        rep("TH", length(3956:3974)),
        rep("TG", length(2909:2942)),
        rep("TB (RUT-3)", length(5160:5238)),
        rep("TF", length(2308:2351)),
        rep("TJ", length(969:985))
      )
    ),
    by = "genome_nt_mut_pos1"
  ) %>%

  dplyr::mutate(
    
    # ---- ORI ----
    genome_nt_pos_ori = dplyr::case_when(
      genome_nt_mut_ham == 1 &
        genome_nt_mut_pos1 %in% c(4299:4328) ~ "Ori (+)",
        genome_nt_mut_pos1 %in% c(2308:2351) ~ "Ori (-)",
      TRUE ~ NA_character_
    ),
    
    # ---- combine annotations ----
    genome_nt_pos_annotation = purrr::pmap_chr(
      list(
        genome_nt_pos_RBS,
        genome_nt_pos_promoter,
        genome_nt_pos_terminator,
        genome_nt_pos_ori,
        genome_nt_mut_pos1
      ),
      ~ {
        values <- na.omit(c(..1, ..2, ..3, ..4))
        
        if (length(values) == 0 &&
            ..5 %in% c(965:1000, 2285:2394, 2923:2930, 3918:3980)) {
          "Unknown Noncoding"
        } else if (length(values) == 0) {
          NA_character_
        } else {
          paste(unique(values), collapse = ",")
        }
      }
    )
  )

table(phix_all_dms_sub_libraries_normalized_unique_stat$genome_nt_pos_annotation)
#           Ori (+)           Promoter                RBS       RBS,Promoter     RBS,Terminator         Terminator Terminator,Ori (-)  Unknown Noncoding 
#                90               1005                331                 20                 33               1270                130                269 

table(phix_all_dms_sub_libraries_normalized_unique_stat$phix_aa_mut_class)
#   Missense   Noncoding    Nonsense ReadThrough  Synonymous 
#      36996         636        2653         352        3558 

#save(phix_all_dms_sub_libraries_normalized_unique_stat, file = "./data/fitness_data/phix_all_dms100_sub_libraries_normalized_unique_stat.RData")


```




## Prepare 1nt and AA variants data
From the data.frame with statistics results for all variants - 'phix_all_dms_sub_libraries_normalized_unique_stat'

```{r}
# check all variants
head(as.data.frame(phix_all_dms_sub_libraries_normalized_unique_stat), 0)
table(duplicated(phix_all_dms_sub_libraries_normalized_unique_stat$genome_mut_ID_nt))
# FALSE 
# 44195 

table(phix_all_dms_sub_libraries_normalized_unique_stat$genome_nt_pos_coding) 
# Coding 
#  43559 

table(phix_all_dms_sub_libraries_normalized_unique_stat$genome_nt_pos_overlap_orf) 
# Overlapping Coding 
#              10133 


```


### 1) 1nt-level data.frame 
- phix_all_dms_sub_libraries_normalized_unique_stat_1nt;

```{r}
# Subset 1nt variants
phix_all_dms_sub_libraries_normalized_unique_stat_1nt <- phix_all_dms_sub_libraries_normalized_unique_stat %>%
  filter(genome_nt_mut_ham == 1)

table(phix_all_dms_sub_libraries_normalized_unique_stat_1nt$phix_aa_mut_class)
#   Missense   Noncoding    Nonsense Stop-loss  Synonymous 
#      11740         636         716          75        2931 



```


### 2) AA-level data.frame
Subset variants encoding amino-acid changes, then combine those causing same changes by taking weighted.mean and propagating the error;
- phix_all_dms_sub_libraries_normalized_unique_stat_AA

```{r}
# Subset Amino-acid variants
phix_all_dms_sub_libraries_normalized_unique_stat_aa <- phix_all_dms_sub_libraries_normalized_unique_stat %>%
  filter(genome_nt_pos_coding == "Coding") 


# Check the frequency of each mut ID, and supplement aa mut frequency and Sample size to the data.frame
count_table_phix_aa_mut_ID <- as.data.frame(table(phix_all_dms_sub_libraries_normalized_unique_stat_aa$phix_aa_mut_ID))
names(count_table_phix_aa_mut_ID) = c("phix_aa_mut_ID", "phix_aa_mut_ID_freq")
table(count_table_phix_aa_mut_ID$phix_aa_mut_ID_freq)
#     1     2     3     4     5 
# 38518   869   587     6   285

phix_all_dms_sub_libraries_normalized_unique_stat_aa <- 
  merge(phix_all_dms_sub_libraries_normalized_unique_stat_aa, count_table_phix_aa_mut_ID, by = "phix_aa_mut_ID")

phix_all_dms_sub_libraries_normalized_unique_stat_aa$phix_aa_mut_ID_sample_size <- 
  phix_all_dms_sub_libraries_normalized_unique_stat_aa$phix_aa_mut_ID_freq*3

table(phix_all_dms_sub_libraries_normalized_unique_stat_aa$phix_aa_mut_ID_freq)
#     1     2     3     4     5 
# 38518  1738  1761    24  1425 


# Calculate fitness for unique genotypes by weighte.mean for the same phix_aa_mut_ID
phix_all_dms_sub_libraries_normalized_unique_stat_aa_dedup <- calculate_unique_weighted_mean_and_error(
  phix_all_dms_sub_libraries_normalized_unique_stat_aa, 
  "phix_aa_mut_ID", "fitness_final", "sigma_final", 
  "fitness_final_unique", "sigma_final_unique"
)
head(as.data.frame(phix_all_dms_sub_libraries_normalized_unique_stat_aa_dedup), 0)


# Remove genome nt annotation columns and unique($phix_aa_mut_ID) and get AA-level data.frame
phix_all_dms_sub_libraries_normalized_unique_stat_AA <- phix_all_dms_sub_libraries_normalized_unique_stat_aa_dedup %>%
  select(
    -starts_with("genome_nt_"),
    -starts_with("gene_sub_nt_"),
    -starts_with("fitness_rep"),
    -starts_with("sigma_rep"),
    -starts_with("fitness_normalized"),
    -starts_with("sigma_rescaled"),
    -genome_mut_ID_nt,
    -fitness_dimsum,
    -sigma_dimsum
  ) %>%
  distinct(phix_aa_mut_ID, .keep_all = TRUE)

dim(phix_all_dms_sub_libraries_normalized_unique_stat_AA)
# [1] 40358    50



```


### 3) Synonymous variants 
Get variant genome sequences for synonymous variants.

```{r}
# mut
phix_restored_mut_genome_seq_D_syn <- phix_all_dms_sub_libraries_normalized_unique_stat %>% 
  filter(gene_sub_aa_mut_gene == "D", gene_sub_aa_mut_ham == 0) %>%
  select(genome_nt_mut_seq)
nrow(phix_restored_mut_genome_seq_D_syn)
# [1] 388x 

phix_restored_mut_genome_seq_F_syn <- phix_all_dms_sub_libraries_normalized_unique_stat %>% 
  filter(gene_sub_aa_mut_gene == "F", gene_sub_aa_mut_ham == 0) %>%
  select(genome_nt_mut_seq)
nrow(phix_restored_mut_genome_seq_F_syn)
# [1] 1058

phix_restored_mut_genome_seq_H_syn <- phix_all_dms_sub_libraries_normalized_unique_stat %>% 
  filter(gene_sub_aa_mut_gene == "H", gene_sub_aa_mut_ham == 0) %>%
  select(genome_nt_mut_seq)
nrow(phix_restored_mut_genome_seq_H_syn)
# [1] 781

phix_restored_mut_genome_seq_K_syn <- phix_all_dms_sub_libraries_normalized_unique_stat %>% 
  filter(gene_sub_aa_mut_gene == "K", gene_sub_aa_mut_ham == 0) %>%
  select(genome_nt_mut_seq)
nrow(phix_restored_mut_genome_seq_K_syn)
# [1] 176


# mut2
phix_restored_mut_genome_seq_B_syn <- phix_all_dms_sub_libraries_normalized_unique_stat %>% 
  filter(gene_sub_aa_mut2_gene == "B", gene_sub_aa_mut2_ham == 0) %>%
  select(genome_nt_mut_seq)
nrow(phix_restored_mut_genome_seq_B_syn)
# [1] 270x 

phix_restored_mut_genome_seq_C_syn <- phix_all_dms_sub_libraries_normalized_unique_stat %>% 
  filter(gene_sub_aa_mut2_gene == "C", gene_sub_aa_mut2_ham == 0) %>%
  select(genome_nt_mut_seq) 
nrow(phix_restored_mut_genome_seq_C_syn)
# [1] 246x 

phix_restored_mut_genome_seq_E_syn <- phix_all_dms_sub_libraries_normalized_unique_stat %>% 
  filter(gene_sub_aa_mut2_gene == "E", gene_sub_aa_mut2_ham == 0) %>%
  select(genome_nt_mut_seq)
nrow(phix_restored_mut_genome_seq_E_syn)
# [1] 283x 

 
# mut3
phix_restored_mut_genome_seq_A_syn <- phix_all_dms_sub_libraries_normalized_unique_stat %>% 
  filter(gene_sub_aa_mut3_gene == "A", gene_sub_aa_mut3_ham == 0) %>%
  select(genome_nt_mut_seq)
nrow(phix_restored_mut_genome_seq_A_syn)
# [1] 1262x 

#phix_restored_mut_genome_seq_Astar_syn <- phix_all_dms_sub_libraries_normalized_unique_stat %>%
phix_restored_mut_genome_seq_G_syn <- phix_all_dms_sub_libraries_normalized_unique_stat %>% 
  filter(gene_sub_aa_mut3_gene == "G", gene_sub_aa_mut3_ham == 0) %>%
  select(genome_nt_mut_seq)
nrow(phix_restored_mut_genome_seq_G_syn)
# [1] 447x 

phix_restored_mut_genome_seq_J_syn <- phix_all_dms_sub_libraries_normalized_unique_stat %>% 
  filter(gene_sub_aa_mut3_gene == "J", gene_sub_aa_mut3_ham == 0) %>%
  select(genome_nt_mut_seq)
nrow(phix_restored_mut_genome_seq_J_syn)
# [1] 108x 


# save 
save(
  phix_restored_mut_genome_seq_A_syn,
  phix_restored_mut_genome_seq_B_syn,
  phix_restored_mut_genome_seq_C_syn,
  phix_restored_mut_genome_seq_D_syn,
  phix_restored_mut_genome_seq_E_syn,
  phix_restored_mut_genome_seq_F_syn,
  phix_restored_mut_genome_seq_G_syn,
  phix_restored_mut_genome_seq_H_syn,
  phix_restored_mut_genome_seq_J_syn,
  phix_restored_mut_genome_seq_K_syn,
  file = "./data/mutant_library_data/Restored/restored_all_synonymous.RData"
)


```



## Update Statistics, separately within the Nt- and AA-level data
1) within 1Nt-level variants
  - use 1nt subs synonymous as reference, the Weighted.mean of the fitness from all 1nt Synonymous data;

2) within AA-level variants
 - for AA unique data, because we merged data of repetitive genotypes, so needs to update separately;
 - get the references for AA variants;
 
3) the fitness data columns are different for 1nt level data and AA level data;

### 1.1) 1nt data, Two-sample t-test and Mutation Category;
```{r}
# Check data
head(as.data.frame(phix_all_dms_sub_libraries_normalized_unique_stat_1nt), 0)


# Running Statistic Tests
phix_all_dms_sub_libraries_normalized_unique_stat_1nt$p_value_1nt_two_sample_ttest_syn <- mapply(
  two_samples_ttest,
  phix_all_dms_sub_libraries_normalized_unique_stat_1nt$fitness_final, 
  phix_all_dms_sub_libraries_normalized_unique_stat_1nt$sigma_final,
  MoreArgs = list(
    mu2 = phix_1nt_synonymous_combined_fitness_final,
    sem2 = phix_1nt_synonymous_combined_sigma_final,
    n1 = 3,  # Mutant replicates, 3x biological replicates per variant
    n2 = 3  # Synonymous population, sum of all at phix organism level from 30x libraries
  )
)


# Apply FDR correction using the function p.adjust() from basic R;
phix_all_dms_sub_libraries_normalized_unique_stat_1nt <- phix_all_dms_sub_libraries_normalized_unique_stat_1nt %>%
  mutate(p_value_1nt_two_sample_ttest_syn_FDR = p.adjust(p_value_1nt_two_sample_ttest_syn, method = "BH"))


# Add raw effect size (difference from the synonymous baseline)
phix_all_dms_sub_libraries_normalized_unique_stat_1nt$effect_size_vs_1nt_syn <-
  phix_all_dms_sub_libraries_normalized_unique_stat_1nt$fitness_final -
  phix_1nt_synonymous_combined_fitness_final


# Add standardized effect (signal-to-noise; same scaling as t-stat numerator)
phix_all_dms_sub_libraries_normalized_unique_stat_1nt$effect_size_std_vs_1nt_syn <-
  (phix_all_dms_sub_libraries_normalized_unique_stat_1nt$fitness_final -
   phix_1nt_synonymous_combined_fitness_final) /
  sqrt(phix_all_dms_sub_libraries_normalized_unique_stat_1nt$sigma_final^2 +
       phix_1nt_synonymous_combined_sigma_final^2)



```

- 1nt Mutation Category 
```{r}
# Check data
head(as.data.frame(phix_all_dms_sub_libraries_normalized_unique_stat_1nt), 0)


# Running Mutation Category
phix_all_dms_sub_libraries_normalized_unique_stat_1nt <- assign_mutation_categories_fdr(
  phix_all_dms_sub_libraries_normalized_unique_stat_1nt,
  pval_col = "p_value_1nt_two_sample_ttest_syn_FDR",
  effect_raw_col = "effect_size_vs_1nt_syn",
  
  alpha = 0.05,
  effect_breaks = c(0.10, 0.15, 0.50),

  # Output column names
  effect_bin_col = "effect_size_bin_1nt",
  category_fdr_col = "mutation_category_FDR_1nt",
  category_fdr_lethal_col = "mutation_category_FDR_lethal_1nt",
  category_cutoff_col = "mutation_category_FDR_cutoff_1nt",
  category_combined_col = "mutation_category_FDR_cutoff_1nt_combined",
  
  direction_col = "mutation_direction_1nt",
  direction_lethal_col = "mutation_direction_lethal_1nt"
)


table(phix_all_dms_sub_libraries_normalized_unique_stat_1nt$mutation_category_FDR_1nt) 
# Beneficial     Neutral Detrimental 
#        176        7693        8229 

table(phix_all_dms_sub_libraries_normalized_unique_stat_1nt$mutation_category_FDR_lethal_1nt)
# Beneficial     Neutral Detrimental      Lethal 
#        176        7693        3164        5065 

table(phix_all_dms_sub_libraries_normalized_unique_stat_1nt$mutation_direction_1nt) 
# Fitness-increasing            Neutral Fitness-decreasing 
#                176               7693               8229 

table(phix_all_dms_sub_libraries_normalized_unique_stat_1nt$mutation_direction_lethal_1nt) 
# Fitness-increasing            Neutral Fitness-decreasing             Lethal 
#                176               7693               3164               5065 



```


### 1.2) AA-level data, Two-sample t-test and Mutation Category;
```{r}
# Check AA data
head(as.data.frame(phix_all_dms_sub_libraries_normalized_unique_stat_AA), 0)
table(phix_all_dms_sub_libraries_normalized_unique_stat_AA$phix_aa_mut_class)
#   Missense    Nonsense   Stop-loss  Synonymous 
#      36081        2542         339        1396 


# Get AA-level reference
filtered_data_AA_synonymous <- phix_all_dms_sub_libraries_normalized_unique_stat_AA %>%
  filter(phix_aa_mut_class == "Synonymous", !is.na(fitness_final_unique))

# get weighted.mean of all 1nt synonymous mutations;
phix_AA_synonymous_combined_weights <- 1 / (filtered_data_AA_synonymous$sigma_final^2)  # Weights are the inverse of variance
phix_AA_synonymous_combined_fitness_final <- sum(filtered_data_AA_synonymous$fitness_final * phix_AA_synonymous_combined_weights) / sum(phix_AA_synonymous_combined_weights) 
# [1] 0.9779537
phix_AA_synonymous_combined_sigma_final <- sqrt(1/sum(phix_AA_synonymous_combined_weights)) 
# [1] 0.0004107135


# Apply the function row-wise, the sample size is related to how many nt muts are merged to one AA mut 
phix_all_dms_sub_libraries_normalized_unique_stat_AA$p_value_AA_two_sample_ttest_syn <- mapply(
  two_samples_ttest,
  phix_all_dms_sub_libraries_normalized_unique_stat_AA$fitness_final_unique,  # mu1 (row-wise)
  phix_all_dms_sub_libraries_normalized_unique_stat_AA$sigma_final_unique,    # sem1 (row-wise)
  MoreArgs = list(
    n1 = 3,  # Fixed (n1),
    mu2 = phix_AA_synonymous_combined_fitness_final,  # Fixed (mu2)
    sem2 = phix_AA_synonymous_combined_sigma_final,   # Fixed (sem2)
    n2 = 3  # Fixed (n2)
  )
)


# Apply FDR correction using the function p.adjust() from basic R;
phix_all_dms_sub_libraries_normalized_unique_stat_AA <- phix_all_dms_sub_libraries_normalized_unique_stat_AA %>%
  mutate(p_value_AA_two_sample_ttest_syn_FDR = p.adjust(p_value_AA_two_sample_ttest_syn, method = "BH"))


# Add raw effect size (difference from the synonymous baseline)
phix_all_dms_sub_libraries_normalized_unique_stat_AA$effect_size_vs_AA_syn <-
  phix_all_dms_sub_libraries_normalized_unique_stat_AA$fitness_final -
  phix_AA_synonymous_combined_fitness_final


# Add standardized effect (signal-to-noise; same scaling as t-stat numerator)
phix_all_dms_sub_libraries_normalized_unique_stat_AA$effect_size_std_vs_AA_syn <-
  (phix_all_dms_sub_libraries_normalized_unique_stat_AA$fitness_final -
   phix_AA_synonymous_combined_fitness_final) /
  sqrt(phix_all_dms_sub_libraries_normalized_unique_stat_AA$sigma_final^2 +
       phix_AA_synonymous_combined_sigma_final^2)



```

- AA Mutation Category
```{r}
# Check data
head(as.data.frame(phix_all_dms_sub_libraries_normalized_unique_stat_AA), 0)


# Running Mutation Category
phix_all_dms_sub_libraries_normalized_unique_stat_AA <- assign_mutation_categories_fdr(
  phix_all_dms_sub_libraries_normalized_unique_stat_AA,
  pval_col = "p_value_AA_two_sample_ttest_syn_FDR",
  effect_raw_col = "effect_size_vs_AA_syn",
  
  #alpha = 1,
  alpha = 0.05,
  effect_breaks = c(0.10, 0.15, 0.50),

  # Output column names
  effect_bin_col = "effect_size_bin_AA",
  category_fdr_col = "mutation_category_FDR_AA",
  category_fdr_lethal_col = "mutation_category_FDR_lethal_AA",
  category_cutoff_col = "mutation_category_FDR_cutoff_AA",
  category_combined_col = "mutation_category_FDR_cutoff_AA_combined",
  
  direction_col = "mutation_direction_AA",
  direction_lethal_col = "mutation_direction_lethal_AA"
)


table(phix_all_dms_sub_libraries_normalized_unique_stat_AA$mutation_category_FDR_AA) 
# Beneficial     Neutral Detrimental 
#        623       11394       28341 

table(phix_all_dms_sub_libraries_normalized_unique_stat_AA$mutation_category_FDR_lethal_AA)
# Beneficial     Neutral Detrimental      Lethal 
#        623       11394        8732       19609 



```



## Save RData
### All Substitutions 
```{r}
## All Data
save(phix_all_dms_sub_libraries_normalized_unique_stat, file = "./data/fitness_data/phix_all_dms100_sub_libraries_normalized_unique_stat.RData")
write_tsv(phix_all_dms_sub_libraries_normalized_unique_stat, file = "./data/fitness_data/phix_all_dms100_sub_libraries_normalized_unique_stat.tsv")
# 'Supplementary_file_phix_variants_fitness.tsv'

## 1nt Substitution Data
save(phix_all_dms_sub_libraries_normalized_unique_stat_1nt, file = "./data/fitness_data/phix_all_dms100_sub_libraries_normalized_unique_stat_1nt.RData")
write_tsv(phix_all_dms_sub_libraries_normalized_unique_stat_1nt, file = "./data/fitness_data/phix_all_dms100_sub_libraries_normalized_unique_stat_1nt.tsv")
# 'Supplementary_file_phix_1nt_variants_fitness.tsv'

## AA Substitution Data
save(phix_all_dms_sub_libraries_normalized_unique_stat_AA, file = "./data/fitness_data/phix_all_dms100_sub_libraries_normalized_unique_stat_AA.RData")
write_tsv(phix_all_dms_sub_libraries_normalized_unique_stat_AA, file = "./data/fitness_data/phix_all_dms100_sub_libraries_normalized_unique_stat_AA.tsv")



```



## Save script checkpoint
save a snapshot of all objects produced by the script

```{r}
# checkpoint - 02 combine, deduplicate and processs all variants, including statistics and mutation category
# save.image(file = "data/checkpoints/02_all_variants_analysis.RData")



```





## Process all variants
All variants and AA variants; (Single-nucleotide variants in script 03)

### 1) make plotting data.frames
- 1) subset three types data: all, nonoverlapping, AA; (1nt mutations are analyzed above)
- 2) add factor level to plotting columns;
  - all variants: phix_all_dms_sub_unique_plotting # [1] 44195
  - single-aa-changed variants: phix_nonover_dms_sub_unique_plotting (phix_aa_mut_ham < 2) # [1] 35235
  - AA variants: phix_all_dms_sub_unique_AA_plotting # [1] 40358

```{r}
# check data
head(as.data.frame(phix_all_dms_sub_libraries_normalized_unique_stat), 0)
table(phix_all_dms_sub_libraries_normalized_unique_stat$phix_aa_mut_class)
nrow(phix_all_dms_sub_libraries_normalized_unique_stat)
# [1] 44195

# prepare plotting data
# 1) all variants (n=44195)
phix_all_dms_sub_unique_plotting <- phix_all_dms_sub_libraries_normalized_unique_stat %>%
  mutate(
    genome_nt_mut_ham = factor(
      genome_nt_mut_ham,
      levels = c("1", "2", "3")
    ),
    phix_aa_mut_class = factor(
      phix_aa_mut_class,
      levels = c("Stop-loss", "Noncoding", "Synonymous", "Nonsense", "Missense")
    )
  ) 
nrow(phix_all_dms_sub_unique_plotting)
# [1] 44195


# 2) nonoverlapping variants (n=35235)
phix_nonover_dms_sub_unique_plotting <- phix_all_dms_sub_unique_plotting %>%
  filter(phix_aa_mut_ham < 2)
nrow(phix_nonover_dms_sub_unique_plotting)
# [1] 35235


# 3) AA substitution Variants (n=40358)
head(as.data.frame(phix_all_dms_sub_libraries_normalized_unique_stat_AA), 0)
phix_all_dms_sub_unique_AA_plotting <- phix_all_dms_sub_libraries_normalized_unique_stat_AA %>%
  mutate(
    phix_aa_mut_class = factor(
      phix_aa_mut_class,
      levels = c("Stop-loss", "Noncoding", "Synonymous", "Nonsense", "Missense")
    )
  ) 
nrow(phix_all_dms_sub_libraries_normalized_unique_stat_AA)
# [1] 40358



```


### 2) summary tables
- FUNCTION: generate_mutation_category_summary

```{r}
# check data
head(as.data.frame(phix_all_dms_sub_unique_plotting), 0)
table(phix_all_dms_sub_unique_plotting$mutation_category_FDR_lethal)
# Beneficial     Neutral Detrimental      Lethal 
#        533       14491        9118       20053 


# run the Summary Function 
summary_phix_sub_mut <- generate_mutation_category_summary(
  phix_all_dms_sub_unique_plotting,
  
  class_col = c("phix_aa_mut_genes"),
  #class_col = c("phix_aa_mut_class"),
  #class_col = c("genome_nt_pos_coding"),
  #class_col = c("genome_nt_pos_nucleotide_class", "phix_aa_mut_class", "phix_aa_mut_genes", "phix_aa_mut_ham"),
  
  #category_col = "mutation_category_FDR"
  category_col = "mutation_category_FDR_lethal"
  #category_col = "mutation_category_FDR_cutoff"
  #category_col = "mutation_category_FDR_cutoff_combined"
)

summary_phix_sub_mut_nonover <- generate_mutation_category_summary(
  phix_nonover_dms_sub_unique_plotting,
  
  class_col = c("genome_nt_pos_nucleotide_class", "phix_aa_mut_class", "phix_aa_mut_genes", "phix_aa_mut_ham"),
  #class_col = c("genome_nt_pos_coding"),
  
  #category_col = "mutation_category_FDR"
  category_col = "mutation_category_FDR_lethal"
  #category_col = "mutation_category_FDR_cutoff"
  #category_col = "mutation_category_FDR_cutoff_combined"
)


# AA
summary_phix_sub_mut_AA <- generate_mutation_category_summary(
  phix_all_dms_sub_unique_AA_plotting,
  phix_all_dms_sub_unique_AA_plotting[phix_all_dms_sub_unique_AA_plotting$phix_aa_mut_ham < 2,],
  
  class_col = c("phix_aa_mut_class", "phix_aa_mut_genes", "phix_aa_mut_ham"),
  category_col = "mutation_category_FDR_cutoff"
)





```



## Plots
### 1) pairwise scatterplot matrix
Replicate correlations (Figure S1)
```{r}
# plotting pearson correlation of 3 replicates
pair_rep_plot <- plot_replicate_reproducibility(
  #df = phix_all_dms_sub_libraries_normalized,
  #df = phix_all_dms_sub_libraries_normalized_unique,
  df = phix_all_dms_sub_libraries_normalized_unique_stat,
  
  replicate_cols = c("fitness_rep1_uncorr", "fitness_rep2_uncorr", "fitness_rep3_uncorr"),
  
  cor_method = "pearson",
  title = "Reproducibility of Fitness Replicates"
)
print(pair_rep_plot)

# saving
ggsave(pair_rep_plot, filename = "./plots/pair_rep_plot.pdf", width = 4, height = 4)




```


### 2) Bar plots + Pie Charts
Mutant Library Composition (Figure S1).
```{r}
## Prepare the pie chart showing 1nt percentage
mut_library_nt_ham <- as.data.frame(table(phix_all_dms_sub_unique_plotting$genome_nt_mut_ham))
mut_library_nt_ham$percent <- round(mut_library_nt_ham$Freq / 44195 * 100, 1)
mut_library_nt_ham$Nt_ham <- paste0(mut_library_nt_ham$Var1, "nt")
mut_library_nt_ham$Nt_ham <- factor(mut_library_nt_ham$Nt_ham, levels = c("1nt", "2nt", "3nt"))
pie_mut_library_nt_ham <- plot_pie_summary(
  mut_library_nt_ham, 
  x_col = "Nt_ham", 
  y_col = "percent",
  outline_linewidth = 0.5,
  plot_title = "Nucleotide Hamming Distance \n(n=43427)", 
  palette = "Set3"
)
print(pie_mut_library_nt_ham)


#ggsave(pie_mut_library_nt_ham, filename = "./plots/pie_mut_library_nt_ham.pdf", width = 5, height = 4)




```


### 3) Density plots (Ridgeline plots)
Mutational effects distribution (Figure S1).
```{r}
## Table Check
table(phix_all_dms_sub_unique_plotting$genome_nt_pos_annotation)


## Plot Ridgeline 
density_mut_class <- plot_ridgeline_mut_class(
  phix_all_dms_sub_unique_plotting, 
  
  class_col = phix_aa_mut_class,
  
  #plot_title = "Fitness Distribution of All Mutants (n=44102))"
  #plot_title = "Fitness Distribution of Nonoverlapping Mutants (n=35233))"
  plot_title = "Fitness Distribution of Single Nucleotide Variants (n=16097))"
  )  + 
  geom_vline(xintercept = 1, linetype = "dashed", color = "black") +
  geom_vline(xintercept = 0, linetype = "dashed", color = "black")
print(density_mut_class)


## Save Plots
ggsave(density_mut_class, filename = "./plots/ridgeline_phix_nt_mut_mut_class.pdf", width = 3.5, height = 3)



```


### 4) Volcano plots (Statistics)
Fitness effects and statistical significance
```{r}
## check data
head(as.data.frame(phix_all_dms_sub_unique_plotting), 0)


## Volcano plot - All Nucleotide Substitution Vatiants
volcano_sub <- plot_volcano(
  #phix_all_dms_sub_unique_plotting, 
  #phix_all_dms_sub_unique_no_plotting,
  #phix_all_dms_sub_unique_AA_plotting,
  
  x_col = "fitness_final", 
  
  #y_col = "p_value_two_sample_ttest_syn_FDR", 
  #y_col = "p_value_AA_two_sample_ttest_syn_FDR", 
  
  #color_col = "mutation_category_FDR",
  #color_col = "mutation_category_FDR_lethal",
  
  #color_col = "mutation_category_FDR_AA",
  #color_col = "mutation_category_FDR_lethal_AA",
  
  x_lim = c(-1, 1.5), 
  y_lim = c(0, 2.7), 
  
  plot_title = "All Substitutions (n=44195)"
  #plot_title = "Nonoverlapping AA Substitutions (AA ham < 2) (n=35227)"
  #plot_title = "Amino Acid Variants (n=40358)"
  ) + 
  #geom_vline(xintercept = phix_all_synonymous_combined_fitness_final, linetype = "dashed", color = "black") + # Lethal - Detrimental
  geom_vline(xintercept = phix_AA_synonymous_combined_fitness_final, linetype = "dashed", color = "black") + # Lethal - Detrimental
  
  #geom_vline(xintercept = 0.5*phix_all_synonymous_combined_fitness_final, linetype = "dashed", color = "black") + # Lethal - Detrimental
  geom_vline(xintercept = 0.5*phix_AA_synonymous_combined_fitness_final, linetype = "dashed", color = "black") + # Lethal - Detrimental
  
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "black")  # FDR threshold line

print(volcano_sub)


# Saving
#ggsave(volcano_sub, filename = "./plots/volcano_all.pdf", width = 3, height = 3)
#ggsave(volcano_sub, filename = "./plots/volcano_nonover.pdf", width = 3, height = 3)
ggsave(volcano_sub, filename = "./plots/volcano_AA.pdf", width = 3, height = 3)


```




## Save script checkpoint
save a snapshot of all objects produced by the script

```{r}
# checkpoint - 02 combine, deduplicate and processs all variants, including statistics and mutation category
# save.image(file = "data/checkpoints/02_all_variants_analysis.RData")



```





