---
title: "phix_DMS_single_nucleotide_variants"
author: "Huijin Wei"
date: "2026-07-30"
output: html_document
editor_options: 
  chunk_output_type: console
---

## Set working working directory
```{r}
here::here()

```



## Load functions
### 1) Data process
```{r}
# Function to annotate 1nt mutations 
annotate_1nt_mutations <- function(df,
                                  wt_col = "genome_nt_mut_wt",
                                  sub_col = "genome_nt_mut_sub") {
  
  stopifnot(wt_col %in% names(df),
            sub_col %in% names(df))
  
  purines <- c("A", "G")
  pyrimidines <- c("C", "T")
  
  df %>%
    dplyr::mutate(
      
      # Mutation string (e.g., A>C)
      genome_nt_mut_base = paste0(.data[[wt_col]], ">", .data[[sub_col]]),
      
      # WT nucleotide type
      genome_nt_mut_wt_type = dplyr::case_when(
        .data[[wt_col]] %in% purines ~ "Purine",
        .data[[wt_col]] %in% pyrimidines ~ "Pyrimidine",
        TRUE ~ NA_character_
      ),
      
      # Sub nucleotide type
      genome_nt_mut_sub_type = dplyr::case_when(
        .data[[sub_col]] %in% purines ~ "Purine",
        .data[[sub_col]] %in% pyrimidines ~ "Pyrimidine",
        TRUE ~ NA_character_
      ),
      
      # Mutation class: Transition / Transversion
      genome_nt_mut_base_type = dplyr::case_when(
        genome_nt_mut_wt_type == genome_nt_mut_sub_type ~ "Transition",
        genome_nt_mut_wt_type != genome_nt_mut_sub_type ~ "Transversion",
        TRUE ~ NA_character_
      ),
      
      # Simplified noncoding annotation
      genome_nt_pos_annotation_combined = case_when(
        grepl(",", genome_nt_pos_annotation) ~ "Multi-Annotation",
        TRUE ~ genome_nt_pos_annotation
      ),
      genome_nt_pos_noncoding_regions = case_when(
        phix_aa_mut_genes == "Noncoding#1 J-F" ~ "J-F",
        phix_aa_mut_genes == "Noncoding#2 F-G" ~ "F-G",
        phix_aa_mut_genes == "Noncoding#3 G-H" ~ "G-H",
        phix_aa_mut_genes == "Noncoding#4 H-A" ~ "H-A",
        TRUE ~ NA_character_
      )
    )
}


# Genome nt columns to keep during the collasping
nt_position_cols <- c(
  # ---- mutation position ----
  "genome_nt_mut_pos", "genome_nt_mut_pos1", "genome_nt_mut_wt", "genome_nt_mut_sub",
  
  # ---- AA mapping (multi-ORF aware) ----
  "gene_sub_nt_mut_pos1", "gene_sub_aa_mut_gene", "gene_sub_aa_mut_pos", "gene_sub_aa_mut_wt",
  "gene_sub_nt_mut2_pos1", "gene_sub_aa_mut2_gene", "gene_sub_aa_mut2_pos", "gene_sub_aa_mut2_wt",
  "gene_sub_nt_mut3_pos1", "gene_sub_aa_mut3_gene", "gene_sub_aa_mut3_pos", "gene_sub_aa_mut3_wt",
  
  # ---- structural annotation ----
  "genome_nt_pos_structural",
  "genome_nt_pos_overlap_orf",
  "genome_nt_pos_coding",
  "genome_nt_pos_protein_monomer",
  "genome_nt_pos_protein_binding",
  "genome_nt_pos_protein_residue",
  "genome_nt_pos_protein_binding_hostspot",
  "phix_aa_mut_genes",
  
  # ---- fitness ----
  "fitness_final", "sigma_final",
  
  # ---- nucleotide / noncoding ----
  "genome_nt_pos_noncoding",
  "genome_nt_pos_nucleotide_class",
  
  # ---- regulatory (broad) ----
  "genome_nt_pos_ori",
  "genome_nt_pos_terminator",
  "genome_nt_pos_promoter",
  "genome_nt_pos_RBS",
  
  # ---- regulatory (gene-level, NEW) ----
  "genome_nt_pos_terminator_gene",
  "genome_nt_pos_promoter_gene",
  "genome_nt_pos_promoter_anti_sense",
  "genome_nt_pos_RBS_gene",
  
  # ---- combined annotations ----
  "genome_nt_pos_annotation",
  "genome_nt_pos_annotation_combined",
  
  # ---- mutation type ----
  "genome_nt_mut_wt_type",
  
  # ---- global statistics ----
  "p_value_two_sample_ttest_syn",
  "p_value_two_sample_ttest_syn_FDR",
  "effect_size_vs_syn",
  "effect_size_std_vs_syn",
  "effect_size_bin",
  
  "mutation_category_FDR",
  "mutation_category_FDR_lethal",
  "mutation_category_FDR_cutoff",
  "mutation_category_FDR_cutoff_combined",
  
  # ---- 1nt subset statistics ----
  "p_value_1nt_two_sample_ttest_syn",
  "p_value_1nt_two_sample_ttest_syn_FDR",
  "effect_size_vs_1nt_syn",
  "effect_size_std_vs_1nt_syn",
  "effect_size_bin_1nt",
  
  "mutation_category_FDR_1nt",
  "mutation_category_FDR_lethal_1nt",
  "mutation_category_FDR_cutoff_1nt",
  "mutation_category_FDR_cutoff_1nt_combined"
)


# Function to summarise 1nt mut data to positions
collapse_nt_to_position <- function(df,
                                    pos_col = "genome_nt_mut_pos",
                                    fitness_col = "fitness_final",
                                    sigma_col = "sigma_final",
                                    wt_reference) {
  
  # ---- Safety check ----
  if (missing(wt_reference) || !is.numeric(wt_reference)) {
    stop("You must supply a numeric wt_reference value.")
  }
  
  df %>%
    dplyr::group_by(.data[[pos_col]]) %>%
    dplyr::summarise(
      
      # ------------------------
      # Sampling
      # ------------------------
      
      n_mutations_nt_pos = dplyr::n(),
      
      n_unique_substitutions =
        dplyr::n_distinct(genome_nt_mut_sub),
      
      n_valid_substitutions =
        sum(!is.na(mutation_category_FDR_1nt)),
      
      n_possible_substitutions = 3,
      
      is_fully_sampled =
        n_unique_substitutions == 3 &
        n_valid_substitutions == 3,
      
      # ------------------------
      # Fitness summaries
      # ------------------------
      
      min_fitness_nt_pos = min(.data[[fitness_col]], na.rm = TRUE),
      max_fitness_nt_pos = max(.data[[fitness_col]], na.rm = TRUE),
      median_fitness_nt_pos = median(.data[[fitness_col]], na.rm = TRUE),
      mean_fitness_nt_pos = mean(.data[[fitness_col]], na.rm = TRUE),
      sd_fitness_nt_pos = sd(.data[[fitness_col]], na.rm = TRUE),
      
      median_sigma_nt_pos =
        if (sigma_col %in% names(df))
          median(.data[[sigma_col]], na.rm = TRUE)
        else NA_real_,
      
      # ------------------------
      # Category counts
      # ------------------------
      
      genome_nt_pos_num_beneficial =
        sum(mutation_category_FDR_lethal_1nt == "Beneficial", na.rm = TRUE),
      
      genome_nt_pos_num_neutral =
        sum(mutation_category_FDR_lethal_1nt == "Neutral", na.rm = TRUE),
      
      genome_nt_pos_num_detrimental =
        sum(mutation_category_FDR_lethal_1nt == "Detrimental", na.rm = TRUE),
      
      genome_nt_pos_num_lethal =
        sum(mutation_category_FDR_lethal_1nt == "Lethal", na.rm = TRUE),
      
      # ------------------------
      # Category fractions
      # ------------------------
      
      genome_nt_pos_frac_beneficial =
        genome_nt_pos_num_beneficial / n_mutations_nt_pos,
      
      genome_nt_pos_frac_neutral =
        genome_nt_pos_num_neutral / n_mutations_nt_pos,
      
      genome_nt_pos_frac_detrimental =
        genome_nt_pos_num_detrimental / n_mutations_nt_pos,
      
      genome_nt_pos_frac_lethal =
        genome_nt_pos_num_lethal / n_mutations_nt_pos,
      
      # ------------------------
      # Essential logic (original)
      # ------------------------
      
      genome_nt_pos_essential_FDR = ifelse(
        any(mutation_category_FDR_1nt == "Detrimental", na.rm = TRUE),
        "Essential", "Non-essential"
      ),
      
      genome_nt_pos_essential_FDR_lethal = ifelse(
        any(mutation_category_FDR_lethal_1nt == "Lethal", na.rm = TRUE),
        "Essential", "Non-essential"
      ),
      
      genome_nt_pos_essential_FDR_deleterious = ifelse(
        any(mutation_category_FDR_cutoff_1nt_combined == "Deleterious", na.rm = TRUE),
        "Essential", "Non-essential"
      ),
      
      # ------------------------
      # Median fitness rule
      # ------------------------
      
      genome_nt_pos_essential_median_fitness = {
        med_fit <- median(.data[[fitness_col]], na.rm = TRUE)
        if (!is.na(med_fit) && med_fit < 0.5 * wt_reference) {
          "Essential"
        } else {
          "Non-essential"
        }
      },
      
      # ------------------------
      # FINAL classification (UPDATED)
      # ------------------------
      
      genome_nt_pos_nucleotide_category = dplyr::case_when(
        
        # strongest signal
        any(mutation_category_FDR_lethal_1nt == "Lethal", na.rm = TRUE) ~ "Essential",
        
        # sub-lethal constraint
        any(mutation_category_FDR_1nt == "Detrimental", na.rm = TRUE) ~ "Sensitive",
        
        # only safe tolerance call
        is_fully_sampled ~ "Tolerant",
        
        # incomplete sampling
        TRUE ~ "Unable to define"
      ),
      
      # ------------------------
      # Preserve invariant columns
      # ------------------------
      
      dplyr::across(
        -c(
          dplyr::any_of(c(pos_col, fitness_col, sigma_col)),
          dplyr::starts_with("mutation_category_")
        ),
        dplyr::first
      ),
      
      .groups = "drop"
    )
}


# Function to do Chi-square Test (with Cramér’s V)
run_chisq_test <- function(df,
                           group_col,
                           category_col,
                           dataset = NA,
                           gene = NA,
                           simulate_p = FALSE) {

  group_sym    <- rlang::ensym(group_col)
  category_sym <- rlang::ensym(category_col)

  dat <- df %>%
    dplyr::filter(!is.na(!!group_sym),
                  !is.na(!!category_sym))

  if (nrow(dat) == 0) {
    stop("No complete cases available for chi-square test.")
  }

  tab <- table(dat[[rlang::as_name(group_sym)]],
               dat[[rlang::as_name(category_sym)]])

  test <- suppressWarnings(
    chisq.test(tab, simulate.p.value = simulate_p)
  )

  # Effect size: Cramér’s V
  cramer_v <- sqrt(as.numeric(test$statistic) /
                   (sum(tab) * (min(dim(tab)) - 1)))

  # Convert tables safely
  observed_df <- as.data.frame(tab)
  colnames(observed_df) <- c("group", "category", "observed")
  
  expected_df <- as.data.frame(test$expected)
  colnames(expected_df) <- c("group", "category", "expected")
  
  residual_df <- as.data.frame(test$stdres)
  colnames(residual_df) <- c("group", "category", "std_resid")

  summary_tbl <- tibble::tibble(
    test_type     = "chi_square",
    dataset       = dataset,
    gene          = gene,
    group_var     = rlang::as_name(group_sym),
    category_var  = rlang::as_name(category_sym),
    n_total       = sum(tab),
    statistic     = unname(test$statistic),
    df            = unname(test$parameter),
    p_value       = test$p.value,
    effect_size   = unname(cramer_v),
    effect_name   = "cramers_v",
    method        = test$method,
    min_expected  = min(test$expected)
  )

  return(list(
    summary   = summary_tbl,
    observed  = observed_df,
    expected  = expected_df,
    residuals = residual_df
  ))
}


# Function to generate summary table
generate_mutation_category_summary <- function(
    df, 
    class_col,
    category_col,
    total_count = NULL
) {

  class_syms    <- rlang::syms(class_col)
  category_sym  <- rlang::ensym(category_col)

  df %>%
    dplyr::filter(!is.na(!!category_sym)) %>%
    dplyr::group_by(!!!class_syms, !!category_sym) %>%
    dplyr::summarise(count = dplyr::n(), .groups = "drop") %>%
    dplyr::group_by(!!!class_syms) %>%
    dplyr::mutate(
      proportion = if (is.null(total_count)) {
        count / sum(count)
      } else {
        count / total_count
      },
      percent = round(proportion * 100, 2)
    ) %>%
    dplyr::ungroup()
}


# Function to generate statistics results
run_fisher_from_summary <- function(summary_df,
                                    group_col,
                                    category_col,
                                    count_col = "count",
                                    test_category,
                                    reference_group = NULL,
                                    comparison_group = NULL,
                                    alternative = "two.sided") {
  
  group_col <- rlang::ensym(group_col)
  category_col <- rlang::ensym(category_col)
  count_col <- rlang::ensym(count_col)
  
  # ---- Collapse category into target vs other ----
  fisher_df <- summary_df %>%
    mutate(
      test_binary = ifelse(
        !!category_col == test_category,
        test_category,
        paste0("Non-", test_category)
      )
    ) %>%
    group_by(!!group_col, test_binary) %>%
    summarise(
      count = sum(!!count_col, na.rm = TRUE),
      .groups = "drop"
    )
  
  # ---- Optionally restrict to two groups ----
  if (!is.null(reference_group) & !is.null(comparison_group)) {
    fisher_df <- fisher_df %>%
      filter(
        !!group_col %in% c(reference_group, comparison_group)
      )
  }
  
  # ---- Convert to 2 x 2 table ----
  fisher_table <- fisher_df %>%
    pivot_wider(
      names_from = test_binary,
      values_from = count,
      values_fill = 0
    ) %>%
    column_to_rownames(rlang::as_string(group_col)) %>%
    as.matrix()
  
  # Make column order clear
  fisher_table <- fisher_table[, c(test_category, paste0("Non-", test_category))]
  
  # Make row order clear, if specified
  if (!is.null(reference_group) & !is.null(comparison_group)) {
    fisher_table <- fisher_table[c(comparison_group, reference_group), ]
  }
  
  # ---- Fisher's exact test ----
  fisher_result <- fisher.test(
    fisher_table,
    alternative = alternative
  )
  
  # ---- Output clean result ----
  result_df <- tibble(
    comparison_group = rownames(fisher_table)[1],
    reference_group = rownames(fisher_table)[2],
    test_category = test_category,
    comparison_target_count = fisher_table[1, test_category],
    comparison_other_count = fisher_table[1, paste0("Non-", test_category)],
    reference_target_count = fisher_table[2, test_category],
    reference_other_count = fisher_table[2, paste0("Non-", test_category)],
    comparison_percent = round(
      fisher_table[1, test_category] / sum(fisher_table[1, ]) * 100,
      2
    ),
    reference_percent = round(
      fisher_table[2, test_category] / sum(fisher_table[2, ]) * 100,
      2
    ),
    odds_ratio = unname(fisher_result$estimate),
    conf_low = fisher_result$conf.int[1],
    conf_high = fisher_result$conf.int[2],
    p_value = fisher_result$p.value
  )
  
  list(
    table = fisher_table,
    fisher_test = fisher_result,
    result = result_df
  )
}



```


### 2) Data plotting
```{r}
# Function to plot volcano plots
plot_volcano <- function(data,
                         x_col = "fitness_final",
                         y_col = "p_adj_two_samples_syn",
                         color_col = "mutation_category_FDR_cutoff",
                         x_lim = c(-1, 1.5),
                         y_lim = c(0, 2.7),
                         plot_title = "Volcano Plot") {

  x_sym <- rlang::sym(x_col)
  y_sym <- rlang::sym(y_col)
  color_sym <- rlang::sym(color_col)

  ggplot(data, aes(x = !!x_sym, y = -log10(!!y_sym), color = !!color_sym)) +
    geom_point(alpha = 0.6, size = 1) +
    scale_color_manual(values = c(
      "Beneficial"  = "#A1A9D0",  # soft blue
      "Neutral"     = "#D9D9D9",  # light gray
      "Detrimental" = "#F4A6A6",  # light red
      "Lethal"      = "#F0988C"   # muted dark red
    )) +
    xlim(x_lim) +
    ylim(y_lim) +
    labs(
      x = "Fitness",
      y = expression(-log[10](FDR)),
      title = plot_title,
      color = "Mutation Category"
    ) +
    theme_classic() +
    theme(
      plot.title = element_text(),
      legend.position = "none"
    )
}


# Function to make Donut plot 
plot_donut_percent <- function(df,
                       category_col = Category,     # bare column name
                       count_col    = count,        # if NULL, will try percent/proportion
                       percent_col  = NULL,         # optional: bare name of percent/proportion (0–1 or 0–100)
                       title        = "Composition",
                       label_acc    = 0.1,          # percent label precision
                       palette      = NULL) {       # optional vector of colors

  cat_sym <- enquo(category_col)
  cnt_sym <- enquo(count_col)
  pct_sym <- enquo(percent_col)

  # 1) collapse to category level and compute percent
  if (!quo_is_null(pct_sym)) {
    df_cat <- df %>%
      group_by(!!cat_sym) %>%
      summarise(percent = sum(!!pct_sym, na.rm = TRUE), .groups = "drop") %>%
      mutate(percent = ifelse(percent > 1, percent/100, percent))
  } else {
    df_cat <- df %>%
      group_by(!!cat_sym) %>%
      summarise(count = sum(!!cnt_sym, na.rm = TRUE), .groups = "drop") %>%
      mutate(percent = count / sum(count))
  }

  df_cat <- df_cat %>%
    mutate(label = percent_format(accuracy = label_acc)(percent))

  # 2) build donut
  p <- ggplot(df_cat, aes(x = 2, y = percent, fill = !!cat_sym)) +
    geom_col(color = "white", width = 0.9) +
    coord_polar(theta = "y") +
    xlim(0.5, 2.5) +  # creates the hole
    geom_text(aes(label = label), position = position_stack(vjust = 0.5), size = 4) +
    labs(title = title, x = NULL, y = NULL, fill = rlang::as_name(cat_sym)) +
    theme_void() +
    theme(legend.position = "none")

  if (!is.null(palette)) {
    p <- p + scale_fill_manual(values = palette)
  }

  p
}



# Function to plot special stacked bar plot
# 1) function to make stacked bar plot for 1nt mutations
bar_percent_stacked_genome <- function(
    summary_df,
    class_col = Class,
    category_col = Category,
    count_col = count,
    percent_col = percent,
    plot_title = "Mutation Category by Class",
    fill_legend_title = "Mutation Category",
    x_order = NULL,
    facet_col = NULL,
    custom_colors = NULL,
    show_segment_labels = FALSE,
    label_digits = 1,
    use_percent = TRUE,   # TRUE = 0–100, FALSE = 0–1
    stat_df = NULL,
    stat_p_col = "p_value",
    stat_format = c("p", "stars"),
    stat_position = c("subtitle", "caption", "none")
) {

  stat_format   <- match.arg(stat_format)
  stat_position <- match.arg(stat_position)

  # ---- Tidy eval ----
  class_sym    <- rlang::ensym(class_col)
  category_sym <- rlang::ensym(category_col)
  count_sym    <- rlang::ensym(count_col)
  percent_sym  <- rlang::ensym(percent_col)
  facet_sym    <- if (!is.null(facet_col)) rlang::ensym(facet_col) else NULL

  class_name    <- rlang::as_name(class_sym)
  category_name <- rlang::as_name(category_sym)
  percent_name  <- rlang::as_name(percent_sym)

  # ---- Default colors ----
  if (is.null(custom_colors)) {
    custom_colors <- c(
      "Beneficial"  = "#A1A9D0",
      "Neutral"     = "#D9D9D9",
      "Detrimental" = "#F4A6A6",
      "Lethal"      = "#F0988C"
    )
  }

  # ---- Clean & aggregate ----
  grouping_vars <- if (is.null(facet_sym)) {
    list(class_sym, category_sym)
  } else {
    list(facet_sym, class_sym, category_sym)
  }

  summary_df <- summary_df %>%
    dplyr::filter(!is.na(!!class_sym), !is.na(!!category_sym)) %>%
    dplyr::group_by(!!!grouping_vars) %>%
    dplyr::summarise(
      !!count_sym := sum(!!count_sym, na.rm = TRUE),
      .groups = "drop"
    )

  # ---- Recompute percentages ----
  group_class_vars <- if (is.null(facet_sym)) {
    list(class_sym)
  } else {
    list(facet_sym, class_sym)
  }

  summary_df <- summary_df %>%
    dplyr::group_by(!!!group_class_vars) %>%
    dplyr::mutate(
      !!percent_sym := if (use_percent) {
        round(!!count_sym / sum(!!count_sym) * 100, 2)
      } else {
        !!count_sym / sum(!!count_sym)
      }
    ) %>%
    dplyr::ungroup()

  # ---- Total counts ----
  total_df <- summary_df %>%
    dplyr::group_by(!!!group_class_vars) %>%
    dplyr::summarise(
      total_count = sum(!!count_sym, na.rm = TRUE),
      .groups = "drop"
    )

  # ---- Category order (robust) ----
  summary_df[[category_name]] <- factor(
    summary_df[[category_name]],
    levels = names(custom_colors),
    ordered = TRUE
  )

  # ---- X ordering ----
  if (!is.null(x_order)) {

    summary_df[[class_name]] <- factor(summary_df[[class_name]], levels = x_order)
    total_df[[class_name]]   <- factor(total_df[[class_name]], levels = x_order)

  } else if (is.null(facet_sym)) {

    class_order <- summary_df %>%
      dplyr::group_by(!!class_sym) %>%
      dplyr::summarise(
        order_val = sum(
          ifelse(
            .data[[category_name]] %in% c("Detrimental", "Lethal"),
            .data[[percent_name]],
            0
          ),
          na.rm = TRUE
        ),
        .groups = "drop"
      ) %>%
      dplyr::arrange(dplyr::desc(order_val)) %>%
      dplyr::pull(!!class_sym)

    summary_df[[class_name]] <- factor(summary_df[[class_name]], levels = class_order)
    total_df[[class_name]]   <- factor(total_df[[class_name]], levels = class_order)
  }

  # ---- Base plot ----
  p <- ggplot2::ggplot(
    summary_df,
    ggplot2::aes(
      x = .data[[class_name]],
      y = .data[[percent_name]],
      fill = .data[[category_name]]
    )
  ) +
    ggplot2::geom_bar(
      stat = "identity",
      position = "stack",
      width = 0.9,
      color = "black"
    ) +
    ggplot2::scale_fill_manual(
      values = custom_colors,
      name = fill_legend_title,
      drop = FALSE
    ) +
    ggplot2::labs(
      title = plot_title,
      x = "Class",
      y = ifelse(use_percent, "Proportion (%)", "Proportion")
    ) +
    ggplot2::theme_classic() +
    ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, hjust = 1))
  
  # ---- Y-axis (STRICT 0–100 + label space) ----
  p <- p +
    ggplot2::scale_y_continuous(
      #limits = c(0, 100),
      breaks = seq(0, 100, 25),
      expand = c(0, 0)
    ) +
    ggplot2::coord_cartesian(ylim = c(0, 110))

  # ---- Segment labels ----
  if (show_segment_labels) {
    p <- p +
      ggplot2::geom_text(
        ggplot2::aes(
          label = if (use_percent) {
            paste0(round(.data[[percent_name]], label_digits), "%")
          } else {
            round(.data[[percent_name]], label_digits)
          }
        ),
        position = ggplot2::position_stack(vjust = 0.5),
        size = 3
      )
  }

  # ---- Total labels ----
  y_max <- ifelse(use_percent, 100, 1)

  p <- p +
    ggplot2::geom_text(
      data = total_df,
      ggplot2::aes(
        x = .data[[class_name]],
        y = y_max * 1.05,
        label = scales::comma(total_count)
      ),
      inherit.aes = FALSE,
      size = 3.5
    ) +
    ggplot2::coord_cartesian(ylim = c(0, y_max * 1.1))

  # ---- Facet ----
  if (!is.null(facet_sym)) {
    p <- p + ggplot2::facet_wrap(vars(!!facet_sym), scales = "free_x")
  }

  # ---- Statistics ----
  if (!is.null(stat_df) && stat_position != "none" && is.null(facet_sym)) {

    p_val <- stat_df[[stat_p_col]][1]

    stat_label <- if (stat_format == "p") {
      ifelse(p_val < 2.2e-16,
             "p < 2.2e-16",
             paste0("p = ", signif(p_val, 3)))
    } else {
      dplyr::case_when(
        p_val < 0.001 ~ "***",
        p_val < 0.01  ~ "**",
        p_val < 0.05  ~ "*",
        TRUE ~ "ns"
      )
    }

    label_text <- paste0("Chi-square test: ", stat_label)

    if (stat_position == "subtitle") {
      p <- p + ggplot2::labs(subtitle = label_text)
    } else if (stat_position == "caption") {
      p <- p + ggplot2::labs(caption = label_text)
    }
  }

  return(p)
}


# 2) function to make stacked bar plot for 1nt in both mutation category and nt classification
# updated: add n= on top of bars
bar_percent_stacked_genome_nucleotide <- function(
  data,
  x_col,
  fill_col,
  palette = NULL,
  title = NULL,
  legend_title = "Factor Class",
  show_legend = TRUE,
  x_order = NULL
) {

  # Required packages:
  # dplyr, ggplot2, rlang

  # --- Capture columns safely ---
  x_sym    <- rlang::ensym(x_col)
  fill_sym <- rlang::ensym(fill_col)

  x_name   <- rlang::as_name(x_sym)
  fill_name <- rlang::as_name(fill_sym)

  # --- Step 1: Compute total counts per x group ---
  total_counts <- data %>%
    dplyr::group_by(!!x_sym) %>%
    dplyr::summarise(total = sum(count, na.rm = TRUE), .groups = "drop")

  # --- Step 2: Sum counts per x and fill category ---
  grouped_data <- data %>%
    dplyr::group_by(!!x_sym, !!fill_sym) %>%
    dplyr::summarise(count = sum(count, na.rm = TRUE), .groups = "drop")

  # --- Step 3: Calculate percentage ---
  percent_data <- grouped_data %>%
    dplyr::left_join(total_counts, by = x_name) %>%
    dplyr::mutate(percent = 100 * count / total) %>%
    dplyr::select(-total)

  # --- Step 4: Handle x ordering ---
  if (!is.null(x_order)) {
    percent_data <- percent_data %>%
      dplyr::mutate(!!x_sym := factor(.data[[x_name]], levels = x_order))
  }

  # --- Step 5: Enforce stack order from palette ---
  if (!is.null(palette) && !is.null(names(palette))) {
    percent_data[[fill_name]] <- factor(
      percent_data[[fill_name]],
      levels = names(palette)
    )
  }

  # --- Step 6: Compute total counts for bar-top labels ---
  bar_totals <- percent_data %>%
    dplyr::group_by(!!x_sym) %>%
    dplyr::summarise(
      n = sum(count, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    #dplyr::mutate(label = paste0("n = ", n))
    dplyr::mutate(label = n)

  # --- Step 7: Build plot ---
  p <- ggplot2::ggplot(percent_data,
              ggplot2::aes(x = !!x_sym,
                           y = percent,
                           fill = !!fill_sym)) +
    ggplot2::geom_bar(stat = "identity",
             position = "stack",
             width = 0.9,
             color = "black") +
    ggplot2::geom_text(
      data = bar_totals,
      ggplot2::aes(x = !!x_sym, y = 100, label = label),
      inherit.aes = FALSE,
      vjust = -0.5,
      size = 4
    ) +
    ggplot2::labs(
      title = title,
      x = NULL,
      y = "Percentage (%)",
      fill = if (show_legend) legend_title else NULL
    ) +
    ggplot2::scale_y_continuous(
      #limits = c(0, 110),   # room for top labels
      #expand = ggplot2::expansion(mult = c(0, 0.05))
      breaks = seq(0, 100, 25),
      labels = scales::label_number(scale = 1),
      expand = expansion(mult = c(0, 0.10))
    ) +
    ggplot2::theme_classic() +
    ggplot2::theme(
      legend.position = if (show_legend) "right" else "none",
      axis.text.x = ggplot2::element_text(angle = 0)
    )

  # --- Step 8: Apply custom palette ---
  if (!is.null(palette)) {
    p <- p +
      ggplot2::scale_fill_manual(
        values = palette,
        drop = FALSE,
        na.value = "white"
      )
  }

  return(p)
}


## Functions for heatmaps
# 1) Function to plot heatmaps for noncoding regions at Nt level
plot_heatmap_noncoding <- function(data,
                                   pos_col = "genome_nt_mut_pos",
                                   sub_col = "genome_nt_mut_sub",
                                   wt_col = "genome_nt_mut_wt",
                                   value_col = "fitness_final",
                                   plot_title = "",
                                   show_legend = TRUE,
                                   show_title = TRUE,
                                   x_label = "WT Nucleotides", # Add custom X label
                                   y_label = "Mutants" # Add custom Y label
                                   ) {
  
  # Convert to symbols
  pos_sym <- rlang::sym(pos_col)
  sub_sym <- rlang::sym(sub_col)
  wt_sym  <- rlang::sym(wt_col)
  val_sym <- rlang::sym(value_col)
  
  # Base plot
  p <- ggplot(data, aes(x = !!pos_sym, y = !!sub_sym, fill = !!val_sym)) +
    geom_tile(color = "white", width = 1, height = 1) +
    coord_fixed() +
    
    # Mark WT
    geom_point(
      aes(x = !!pos_sym, y = !!wt_sym),
      shape = 4, size = 0.8, color = "black", stroke = 0.6,
      inherit.aes = FALSE
    ) +
    
    # Color scale
    scale_fill_gradientn(
      colors = c("darkred", "red", "gray88", "blue"),
      values = scales::rescale(c(-1, 0, 1, 1.5)),
      limits = c(-1, 1.5),
      name = value_col,
      na.value = "white",
      guide = if (show_legend) guide_colorbar(
        barwidth = 10,
        barheight = 1,
        direction = "horizontal",
        ticks.colour = "black",
        title.position = "top",
        title.hjust = 0.5
      ) else "none"
    ) +
    
    # X axis: WT letters
    scale_x_continuous(
      breaks = data[[pos_col]],
      labels = data[[wt_col]],
      expand = c(0, 0)
    ) +
    
    # Y axis
    scale_y_discrete(expand = c(0, 0)) +
    
    # Theme
    theme_minimal(base_size = 12) +
    theme(
      axis.text.x = element_text(hjust = 1, vjust = 0.5),
      axis.text.y = element_text(vjust = 0.5),
      axis.title.x = element_text(face = "bold", size = 12),
      axis.title.y = element_text(face = "bold", size = 12),
      panel.grid = element_blank(),
      legend.position = if (show_legend) "right" else "none",
      plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
      plot.margin = margin(10, 10, 10, 10)
    )

  # Add title if needed
  if (show_title) {
    p <- p + ggtitle(plot_title)
  }

  # Set x and y axis labels if provided
  if (!is.null(x_label)) {
    p <- p + xlab(x_label)
  }
  if (!is.null(y_label)) {
    p <- p + ylab(y_label)
  }

  return(p)
}


# 2) Function for heatmaps with a color for the wt blocks 
plot_heatmap_noncoding_colored_wt <- function(data,
                                   pos_col = "genome_nt_mut_pos",
                                   sub_col = "genome_nt_mut_sub",
                                   wt_col = "genome_nt_mut_wt",
                                   value_col = "fitness_final",
                                   plot_title = "",
                                   show_legend = TRUE,
                                   show_title = TRUE,
                                   x_label = "WT Nucleotides",
                                   y_label = "Mutants",
                                   wt_value = 1.0,
                                   wt_shape = 21,
                                   wt_size = 0.8) {
  
  library(dplyr)
  library(ggplot2)
  library(rlang)
  
  # ---- symbols ----
  pos_sym <- sym(pos_col)
  sub_sym <- sym(sub_col)
  wt_sym  <- sym(wt_col)
  val_sym <- sym(value_col)
  
  # ---- 1) Create WT rows ----
  wt_rows <- data %>%
    distinct(!!pos_sym, !!wt_sym) %>%
    mutate(
      !!sub_sym := .data[[wt_col]],
      !!val_sym := wt_value,
      is_wt = TRUE
    )
  
  # ---- 2) Combine data ----
  data_plot <- bind_rows(
    data %>% mutate(is_wt = FALSE),
    wt_rows
  )
  
  # ---- 3) X-axis mapping (stable & unique) ----
  x_map <- data %>%
    distinct(!!pos_sym, !!wt_sym) %>%
    arrange(!!pos_sym)
  
  # ---- 4) Plot ----
  p <- ggplot(data_plot, aes(
    x = !!pos_sym,
    y = !!sub_sym,
    fill = !!val_sym
  )) +
    
    geom_tile(color = "white", width = 1, height = 1) +
    
    # ---- WT overlay ----
    geom_point(
      data = dplyr::filter(data_plot, is_wt),
      aes(x = !!pos_sym, y = !!sub_sym),
      shape = wt_shape,
      size = wt_size,
      stroke = 0.5,
      color = "black",
      fill = "white",
      inherit.aes = FALSE
    ) +
    
    coord_fixed() +
    
    # ---- Color scale ----
    scale_fill_gradientn(
      colors = c("darkred", "red", "gray88", "blue"),
      values = scales::rescale(c(-1, 0, 1, 1.5)),
      limits = c(-1, 1.5),
      name = value_col,
      na.value = "white",
      guide = if (show_legend) guide_colorbar(
        barwidth = 10,
        barheight = 1,
        direction = "horizontal",
        title.position = "top",
        title.hjust = 0.5
      ) else "none"
    ) +
    
    # ---- X axis ----
    scale_x_continuous(
      breaks = x_map[[pos_col]],
      labels = x_map[[wt_col]],
      expand = c(0, 0)
    ) +
    
    # ---- Y axis ----
    scale_y_discrete(expand = c(0, 0)) +
    
    # ---- Theme ----
    theme_minimal(base_size = 12) +
    theme(
      axis.text.x = element_text(hjust = 1, vjust = 0.5),
      axis.text.y = element_text(vjust = 0.5),
      axis.title.x = element_text(face = "bold"),
      axis.title.y = element_text(face = "bold"),
      panel.grid = element_blank(),
      legend.position = if (show_legend) "right" else "none",
      plot.title = element_text(hjust = 0.5, face = "bold"),
      plot.margin = margin(10, 10, 10, 10)
    )
  
  # ---- Title ----
  if (show_title) {
    p <- p + ggtitle(plot_title)
  }
  
  # ---- Axis labels ----
  if (!is.null(x_label)) {
    p <- p + xlab(x_label)
  }
  if (!is.null(y_label)) {
    p <- p + ylab(y_label)
  }
  
  return(p)
}





```



## Load data
### 1) load all and 1nt variants
```{r}
## all variants
load("./data/fitness_data/phix_all_dms100_sub_libraries_normalized_unique_stat.RData")
head(as.data.frame(phix_all_dms_sub_libraries_normalized_unique_stat), 0)
nrow(phix_all_dms_sub_libraries_normalized_unique_stat)
# [1] 44195


## 1nt Data
## Load 1nt DMS data after Statistic tests
load("./data/fitness_data/phix_all_dms100_sub_libraries_normalized_unique_stat_1nt.RData")
head(as.data.frame(phix_all_dms_sub_libraries_normalized_unique_stat_1nt), 0)
nrow(phix_all_dms_sub_libraries_normalized_unique_stat_1nt)
# [1] 16098



```


### 2) get 1nt Synonymous reference
  - for collapsing 1nt mutations data
  - for plotting volcano plots

```{r}
## 1nt Data
filtered_data_1nt_synonymous <- phix_all_dms_sub_libraries_normalized_unique_stat %>%
  filter(genome_nt_mut_ham == 1, phix_aa_mut_class == "Synonymous", !is.na(fitness_final))

# get weighted.mean of all 1nt synonymous mutations;
phix_1nt_synonymous_combined_weights <- 1 / (filtered_data_1nt_synonymous$sigma_final^2)  # Weights are the inverse of variance
phix_1nt_synonymous_combined_fitness_final <- sum(filtered_data_1nt_synonymous$fitness_final * phix_1nt_synonymous_combined_weights) / sum(phix_1nt_synonymous_combined_weights) 
# [1] 0.978221
phix_1nt_synonymous_combined_sigma_final <- sqrt(1/sum(phix_1nt_synonymous_combined_weights)) 
# [1] 0.0002813241



```


### 3) look-up table for 1aa synonymous mutations
To quantify how many 1aa synonymous mutations have codons fall into nt-level regulatory elements

```{r}
# check data 
head(as.data.frame(phix_all_dms_sub_libraries_normalized_unique_stat), 0)

# look-up table
phix_1aa_syn_genome_nt_look_up <- phix_all_dms_sub_libraries_normalized_unique_stat %>%
  filter(!is.na(genome_nt_pos_annotation), phix_aa_mut_class == "Synonymous") %>%
  select(phix_aa_mut_class, phix_aa_mut_ID, genome_nt_pos_RBS, genome_nt_pos_RBS_gene, genome_nt_pos_promoter, genome_nt_pos_promoter_gene, genome_nt_pos_promoter_anti_sense, genome_nt_pos_terminator, genome_nt_pos_terminator_gene, genome_nt_pos_ori, genome_nt_pos_annotation)

nrow(phix_1aa_syn_genome_nt_look_up)
# [1] 123

phix_1aa_syn_genome_nt_look_up_unique <- phix_1aa_syn_genome_nt_look_up %>%
  separate_rows(phix_aa_mut_ID, sep = ",") %>%
  mutate(phix_aa_mut_ID = trimws(phix_aa_mut_ID)) %>%
  distinct()

nrow(phix_1aa_syn_genome_nt_look_up_unique)
# [1] 49


## Saving the advanced proteome data.frame
#save(phix_1aa_syn_genome_nt_look_up_unique, file = "./data/proteins_data/plotting_data/phix_1aa_syn_genome_nt_look_up_unique.RData")
write_tsv(phix_1aa_syn_genome_nt_look_up_unique, file = "./data/proteins_data/plotting_data/phix_1aa_syn_genome_nt_look_up_unique.tsv")



```



## Process 1nt variants
1) update 1nt mut data and annotations
2) process 1nt mut data to get nucleotide residue level data
  - FUNCTION: collapse_nt_to_position
3) statistic tests 

```{r}
# Check data
head(as.data.frame(phix_all_dms_sub_libraries_normalized_unique_stat_1nt), 0)
table(phix_all_dms_sub_libraries_normalized_unique_stat_1nt$mutation_category_FDR_lethal_1nt)


# 1) updated 1nt variants data 
phix_all_dms_sub_unique_1nt_annotated <- annotate_1nt_mutations(phix_all_dms_sub_libraries_normalized_unique_stat_1nt)


table(phix_all_dms_sub_unique_1nt_annotated$genome_nt_pos_nucleotide_class)
#            Coding          Noncoding Overlapping Coding 
#             13021                636               2441 

table(phix_all_dms_sub_unique_1nt_annotated$mutation_category_FDR_1nt)
# Beneficial     Neutral Detrimental 
#        176        7693        8229 

table(phix_all_dms_sub_unique_1nt_annotated$mutation_category_FDR_1nt[!is.na(phix_all_dms_sub_unique_1nt_annotated$genome_nt_pos_noncoding_regions)])
# Beneficial     Neutral Detrimental 
#         41         473         122 

table(phix_all_dms_sub_unique_1nt_annotated$genome_nt_pos_annotation)
#       Ori (+)           Promoter                RBS       RBS,Promoter     RBS,Terminator         Terminator Terminator,Ori (-)  Unknown Noncoding 
#            90                381                120                  6                 33                411                130                269 



# 2) get 1nt residues data 
phix_all_dms_sub_unique_1nt_pos_annotated <- phix_all_dms_sub_unique_1nt_annotated %>%
  select(dplyr::any_of(nt_position_cols)) %>%
  collapse_nt_to_position(wt_reference = phix_1nt_synonymous_combined_fitness_final)


table(phix_all_dms_sub_unique_1nt_pos_annotated$genome_nt_pos_nucleotide_class)
#         Coding          Noncoding Overlapping Coding 
#           4355                217                814 

table(phix_all_dms_sub_unique_1nt_pos_annotated$genome_nt_pos_nucleotide_category)
#       Essential        Sensitive         Tolerant Unable to define 
#            2588             1102             1664               32 

write_tsv(phix_all_dms_sub_unique_1nt_pos_annotated, file = "./data/fitness_data/phix_all_dms_sub_unique_1nt_pos_annotated.tsv")
# renamed 'Supplementary_file_phix_1nt_residues.tsv'




```



## Statistics
- global category: Chi-square
```{r}
# checking data 
head(as.data.frame(phix_all_dms_sub_unique_1nt_annotated), 2)
head(as.data.frame(phix_all_dms_sub_unique_1nt_pos_annotated), 0)


# 1) mutations
phix_genome_chisq_test <- run_chisq_test(
  phix_all_dms_sub_unique_1nt_annotated,
  #phix_all_dms_sub_unique_1nt_annotated[phix_all_dms_sub_unique_1nt_annotated$genome_nt_pos_noncoding == "Noncoding",],
  #phix_all_dms_sub_unique_1nt_pos_annotated,
  
  #genome_nt_pos_wt,
  genome_nt_pos_nucleotide_class,
  
  #mutation_category_FDR_lethal_1nt, 
  mutation_category_FDR_1nt,
  
  dataset="Genome 1nt Mutations", 
  
  gene="all"
  )

phix_genome_chisq_test$summary
view(phix_genome_chisq_test$summary)


# 2) residues
phix_genome_chisq_test <- run_chisq_test(
  phix_all_dms_sub_unique_1nt_pos_annotated,
  
  #aa_mut_gene, 
  #aa_mut_class, 
  genome_nt_pos_nucleotide_class,
  
  genome_nt_pos_nucleotide_category,
  
  dataset="Genome 1nt Residues", 
  
  gene="all"
  )

phix_genome_chisq_test$summary
view(phix_genome_chisq_test$summary)



```



## Summary 
Summary 1nt mutation and residue data
  - FUNCTION: 'generate_mutation_category_summary'

### 1) 1nt mutations
  - $mutation_category_FDR_lethal_1nt
```{r}
# Check data
head(as.data.frame(phix_all_dms_sub_unique_1nt_annotated), 0)
#head(as.data.frame(phix_all_dms_sub_unique_1nt_pos_annotated), 0)
table(phix_all_dms_sub_unique_1nt_annotated$mutation_category_FDR_1nt)
# Beneficial     Neutral Detrimental 
#        176        7693        8229 

# Run the Summary Function 
# 1) all 1nt mut
summary_1nt_mut <- generate_mutation_category_summary(
  phix_all_dms_sub_unique_1nt_annotated,
  
  class_col = c("genome_nt_pos_nucleotide_class"),
  #class_col = c("genome_nt_pos_coding"),
  #class_col = c("genome_nt_mut_wt"),
  #class_col = c("genome_nt_mut_base_type"),
  #class_col = c("genome_nt_mut_wt_type", "genome_nt_mut_sub_type", "genome_nt_mut_base_type"),
  #class_col = c("genome_nt_mut_wt", "genome_nt_pos_nucleotide_class", "genome_nt_mut_wt_type", "genome_nt_mut_sub_type", "genome_nt_mut_base_type"),
  
  category_col = "mutation_category_FDR_1nt"
  #category_col = "mutation_category_FDR_lethal_1nt"
  #category_col = "mutation_category_FDR_cutoff_1nt"
  #category_col = "mutation_category_FDR_cutoff_1nt_combined"
)

#write_tsv(summary_1nt_mut, file = "./data/summary_tables/summary_1nt_mut.tsv")

fisher_random <- run_fisher_from_summary(
  summary_df = summary_1nt_mut,
  group_col = genome_nt_pos_nucleotide_class,
  category_col = mutation_category_FDR_1nt,
  count_col = count,
  test_category = "Detrimental",
  comparison_group = "Coding",
  reference_group = "Overlapping Coding",
  alternative = "two.sided"
)

fisher_random$table
fisher_random$result



# Check noncoding data
table(phix_all_dms_sub_unique_1nt_annotated$genome_nt_pos_annotation_combined[phix_all_dms_sub_unique_1nt_annotated$genome_nt_pos_nucleotide_class == "Noncoding" & phix_all_dms_sub_unique_1nt_annotated$mutation_category_FDR_1nt == "Detrimental"])
# Multi-Annotation          Promoter               RBS        Terminator Unknown Noncoding 
#               38                35                12                14                23 

table(phix_all_dms_sub_unique_1nt_annotated$mutation_category_FDR_1nt[phix_all_dms_sub_unique_1nt_annotated$genome_nt_pos_nucleotide_class == "Noncoding" & phix_all_dms_sub_unique_1nt_annotated$genome_nt_mut_pos %in% c(2353:2373)])
## nt 3927-3932
# Beneficial     Neutral Detrimental 
#          0           1          15 

## nt 3950-3955
# Beneficial     Neutral Detrimental 
#          0           7          11 

## nt 3945-3948
# Beneficial     Neutral Detrimental 
#          0           1           9 

## nt 3967-3975, 3' stem of TH
# Beneficial     Neutral Detrimental 
#          0          10          17 

## nt 3956-3963, 5' stem of TH
# Beneficial     Neutral Detrimental 
#          0          20           3 

## nt 2308-2327, 5' stem of TF
# Beneficial     Neutral Detrimental 
#         38          15           7 

## nt 2328-2351, loop + 3' stem of TF
# Beneficial     Neutral Detrimental 
#          0          58          12 

## nt T2321C
# Beneficial     Neutral Detrimental 
#          0           1           0 

## nt 2353-2373, potential TF2
# Beneficial     Neutral Detrimental 
#          0          47          15 


# 2) noncoding 1nt mut 
summary_1nt_noncoding_mut <- generate_mutation_category_summary(
  phix_all_dms_sub_unique_1nt_annotated[phix_all_dms_sub_unique_1nt_annotated$genome_nt_pos_nucleotide_class == "Noncoding",],
  #phix_all_dms_sub_unique_1nt_annotated[phix_all_dms_sub_unique_1nt_annotated$genome_nt_pos_nucleotide_class == "Noncoding" & phix_all_dms_sub_unique_1nt_annotated$mutation_category_FDR_1nt == "Detrimental",],
  #phix_all_dms_sub_unique_1nt_annotated[phix_all_dms_sub_unique_1nt_annotated$genome_nt_pos_nucleotide_class == "Noncoding" & phix_all_dms_sub_unique_1nt_annotated$mutation_category_FDR_1nt == "Beneficial",],
  
  #phix_all_dms_sub_unique_1nt_annotated[phix_all_dms_sub_unique_1nt_annotated$genome_nt_pos_nucleotide_class == "Noncoding" & phix_all_dms_sub_unique_1nt_annotated$genome_nt_mut_pos %in% c(3950:3955),],
  #phix_all_dms_sub_unique_1nt_annotated[phix_all_dms_sub_unique_1nt_annotated$genome_nt_pos_nucleotide_class == "Noncoding" & phix_all_dms_sub_unique_1nt_annotated$genome_nt_pos_annotation_combined == "Promoter",],
  
  #class_col = c("phix_aa_mut_genes"),
  #class_col = c("genome_nt_mut_wt"),
  class_col = c("genome_nt_pos_annotation"),
  #class_col = c("genome_nt_mut_wt", "genome_nt_pos_annotation_combined"),
  #class_col = c("genome_nt_pos_nucleotide_class"),
  #class_col = c("genome_nt_pos_noncoding_regions"),
  #class_col = c("genome_nt_pos_annotation_combined"),
  #class_col = c("phix_aa_mut_genes", "genome_nt_pos_noncoding_regions", "genome_nt_pos_annotation_combined"),
  #class_col = c("phix_aa_mut_genes", "genome_nt_mut_wt", "genome_nt_pos_nucleotide_class", "genome_nt_mut_wt_type", "genome_nt_mut_sub_type", "genome_nt_mut_base_type"),
  
  #class_col = c("mutation_category_FDR_1nt"),
  
  #category_col = "genome_nt_pos_annotation"
  #category_col = "genome_nt_pos_annotation_combined"
  
  #category_col = "mutation_category_FDR_1nt"
  category_col = "mutation_category_FDR_lethal_1nt"
  #category_col = "mutation_category_FDR_cutoff_1nt"
  #category_col = "mutation_category_FDR_cutoff_1nt_combined"
)

#write_tsv(summary_1nt_noncoding_mut, file = "./data/summary_tables/summary_1nt_noncoding_mut.tsv")


# 3) detrimental noncoding 1nt mut
#summary_1nt_noncoding_mut_deleterious <- generate_mutation_category_summary(
  #phix_all_dms_sub_unique_1nt_annotated[phix_all_dms_sub_unique_1nt_annotated$genome_nt_pos_nucleotide_class == "Noncoding" & phix_all_dms_sub_unique_1nt_annotated$mutation_category_FDR_lethal_1nt %in% c("Detrimental", "Lethal"),],
  
  #class_col = c("phix_aa_mut_genes"),
  #class_col = c("genome_nt_pos_annotation"),
  #class_col = c("genome_nt_pos_annotation_combined"),
  #class_col = c("phix_aa_mut_genes", "genome_nt_pos_annotation"),
  #class_col = c("genome_nt_mut_wt", "phix_aa_mut_genes", "genome_nt_pos_annotation", "genome_nt_pos_noncoding_regions", "genome_nt_pos_annotation_combined"),
  
  #total_count = 21223,
 
  #category_col = "mutation_category_FDR_1nt"
  #category_col = "mutation_category_FDR_lethal_1nt"
  #category_col = "mutation_category_FDR_cutoff_1nt"
  #category_col = "mutation_category_FDR_cutoff_1nt_combined"
)






```



### 2) 1nt residues
  - $genome_nt_pos_nucleotide_category
```{r}
# Check data
#head(as.data.frame(phix_all_dms_sub_unique_1nt_annotated), 0)
head(as.data.frame(phix_all_dms_sub_unique_1nt_pos_annotated), 0)
table(phix_all_dms_sub_unique_1nt_pos_annotated$genome_nt_pos_nucleotide_category)
#       Essential        Sensitive         Tolerant Unable to define 
#            2588             1102             1664               32 

# 1) all 1nt nucleotide
summary_1nt_pos_essential_category <- generate_mutation_category_summary(
  phix_all_dms_sub_unique_1nt_pos_annotated,
  
  class_col = "genome_nt_pos_nucleotide_class",
  #class_col = c("genome_nt_pos_nucleotide_class", "genome_nt_pos_annotation_combined"),
  #class_col = c("phix_aa_mut_genes", "genome_nt_pos_annotation", "genome_nt_pos_factor_plot"),
  
  #total_count = 21223,
 
  category_col = "genome_nt_pos_nucleotide_category"
)

#write_tsv(summary_1nt_pos_essential_category, file = "./data/summary_tables/summary_1nt_pos_essential_category.tsv")


# 2) noncoding 1nt nucleotide 
summary_1nt_pos_noncoding_essential_category <- generate_mutation_category_summary(
  phix_all_dms_sub_unique_1nt_pos_annotated[!is.na(phix_all_dms_sub_unique_1nt_pos_annotated$genome_nt_pos_noncoding),],
  
  class_col = c("phix_aa_mut_genes"),
  #class_col = "genome_nt_pos_nucleotide_class",
  
  #total_count = 21223,
 
  category_col = "genome_nt_pos_nucleotide_category"
)



```



## Plots
### 1) Volcano plots (Statistics)
Fitness effects and statistical significance
```{r}
## check data
head(as.data.frame(phix_all_dms_sub_unique_1nt_plotting), 0)
table(phix_all_dms_sub_unique_1nt_plotting$mutation_category_FDR_1nt)

head(as.data.frame(phix_all_dms_sub_unique_AA_plotting), 0)


## Volcano plot - All Nucleotide Substitution Vatiants
volcano_sub <- plot_volcano(
  phix_all_dms_sub_unique_1nt_annotated,
  
  x_col = "fitness_final", 
  
  y_col = "p_value_1nt_two_sample_ttest_syn_FDR", 
  
  #color_col = "mutation_category_FDR",
  #color_col = "mutation_category_FDR_lethal",
  
  #color_col = "mutation_category_FDR_1nt",
  color_col = "mutation_category_FDR_lethal_1nt",
  
  x_lim = c(-1, 1.5), 
  y_lim = c(0, 2.7), 
  
  #plot_title = "All Substitutions (n=44195)"
  #plot_title = "Nonoverlapping AA Substitutions (AA ham < 2) (n=35227)"
  plot_title = "Single Nucleotide Variants (n=16098)"
  #plot_title = "Amino Acid Variants (n=40358)"
  ) + 
  geom_vline(xintercept = phix_1nt_synonymous_combined_fitness_final, linetype = "dashed", color = "black") + # Lethal - Detrimental
  geom_vline(xintercept = 0.5*phix_1nt_synonymous_combined_fitness_final, linetype = "dashed", color = "black") + # Lethal - Detrimental
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", color = "black")  # FDR threshold line
print(volcano_sub)



# Saving
#ggsave(volcano_sub, filename = "./plots/volcano_1nt.pdf", width = 3, height = 3)



```



### 2) donut plot
Mutation category, proportion of detrimental mutations (Figure 1)
```{r}
## Plotting
donut_plot_phix_mut <- plot_donut_percent(
  summary_1nt_mut,
  #summary_phix_sub_mut_1nt, 
  
  #summary_phix_sub_mut_AA,
  #summary_phix_sub_mut_AA_nonover,
  
  category_col = mutation_category_FDR_lethal_1nt,
  
  count_col = count,
  title = NULL,
  palette = c(
    "Beneficial"="#A1A9D0",
    "Neutral"="#D9D9D9",
    "Detrimental"="#F4A6A6",
    "Lethal"="#F0988C"
    )
  )
print(donut_plot_phix_mut)


## Saving
ggsave(donut_plot_phix_mut, filename = "./plots/donut_plot_phix_mut_1nt.pdf", width = 3, height = 3)


```

Nucleotide residue category, proportion of essential nucleotide residues (Figure S1)
```{r}
# checking 
head(as.data.frame(summary_1nt_pos_essential_category), 0)
table(summary_1nt_pos_essential_category$genome_nt_pos_nucleotide_category)


# plotting
donut_plot_genome_nt <- plot_donut_percent(
  df = summary_1nt_pos_essential_category,
  #df = summary_1nt_pos_essential_category[summary_1nt_pos_essential_category$genome_nt_pos_nucleotide_class == "Noncoding",],
  
  #category_col = genome_nt_pos_factor_plot, 
  #category_col = genome_nt_pos_annotation_combined,
  category_col = genome_nt_pos_nucleotide_category,
  
  count_col = count,
  
  #title = "Amino Acid Mutations Category",
  
  #palette = c(
    #"Beneficial"="#A1A9D0",
    #"Neutral"="#D9D9D9",
    #"Detrimental"="#F4A6A6",
    #"Lethal"="#F0988C"
    #)
  
  palette = c(
    "Unable to define" = "#2B82AC",  
    "Tolerant"         = "#DFCBBC",  
    "Sensitive"        = "#EDB01D",  
    "Essential"        = "#CD6143"
    )
  )
print(donut_plot_genome_nt)


## Saving
ggsave(donut_plot_genome_nt, filename = "./plots/donut_plot_genome_nt_category.pdf", width = 3, height = 3)


```



### 3) stacked bar plot
Mutation category, proportion of detrimental mutations across genomic regions (Figure 1)
```{r}
# check data 
head(as.data.frame(summary_1nt_noncoding_mut), 0)
table(summary_1nt_noncoding_mut$mutation_category_FDR_lethal_1nt)

## Plotting
bar_stacked_proteome_1nt_mut <- bar_percent_stacked_genome(
  #summary_1nt_mut,
  summary_1nt_noncoding_mut,
  
  #class_col = genome_nt_mut_wt,
  #class_col = genome_nt_pos_nucleotide_class,
  #class_col = genome_nt_pos_noncoding_regions,
  class_col = genome_nt_pos_annotation,

  category_col = mutation_category_FDR_lethal_1nt,
  #category_col = mutation_category_FDR_lethal,
  
  count_col = count,
  percent_col = percent,
  
  #plot_title = "Substitution Mutations category by Secondary Structures",
  fill_legend_title = "Mutation Category",
  
  #x_order = c("multi_orf", "single_orf"),
  #facet_col = "aa_mut_gene",
  
  #stat_df = NULL,
  #stat_p_col = "p_value",
  #stat_group_col = NULL,
  #stat_format = c("p", "stars"),
  #stat_position = c("subtitle", "caption", "none")
  
  ) +
  theme(legend.position = "none")
print(bar_stacked_proteome_1nt_mut)


## Saving
#ggsave(bar_stacked_proteome_1nt_mut, filename = "./plots/bar_stacked_genome_1nt_mut.pdf", width = 3, height = 3)
#ggsave(bar_stacked_proteome_1nt_mut, filename = "./plots/bar_stacked_noncoding_1nt_mut.pdf", width = 5, height = 4)






```

Nucleotide residue category, proportion of essential nucleotide residues across genomic regions (Figure S1)
```{r}
## Check Data
# checking 
head(as.data.frame(summary_1nt_pos_essential_category), 0)
table(summary_1nt_pos_essential_category$genome_nt_pos_nucleotide_category)


## Plot the Nucleotide Category by proteins
bar_stacked_genome_nt_category <- bar_percent_stacked_genome_nucleotide(
  data = summary_1nt_pos_essential_category,
  #data = summary_1nt_pos_essential_category[summary_1nt_pos_essential_category$genome_nt_pos_nucleotide_class == "Noncoding",],
  
  #x_col = phix_aa_mut_genes,
  x_col = genome_nt_pos_nucleotide_class,
  #x_col = genome_nt_pos_annotation_combined,
  
  #fill_col = genome_nt_pos_annotation_combined,
  fill_col = genome_nt_pos_nucleotide_category,
  
  #factor_col = "genome_nt_pos_factor_plot",
  
  palette = c(
    "Unable to define" = "#2B82AC",  
    "Tolerant"         = "#DFCBBC",  
    "Sensitive"        = "#EDB01D",  
    "Essential"        = "#CD6143"
    ),
  
  x_order = c("Overlapping Coding", "Coding", "Noncoding"),
  
  show_legend = FALSE
)
print(bar_stacked_genome_nt_category)


## Saving
ggsave(bar_stacked_genome_nt_category, filename = "./plots/bar_stacked_genome_nt_category.pdf", width = 2, height = 2)



```



### 4) Noncoding region heatmaps
```{r}
## Subset Noncoding 1nt DMS data 
head(as.data.frame(phix_all_dms_sub_unique_1nt_annotated), 0)
table(phix_all_dms_sub_unique_1nt_annotated$genome_nt_pos_noncoding)
# Noncoding 
#       636 

phix_dms_sub_unique_1nt_noncoding_cols <- phix_all_dms_sub_unique_1nt_annotated %>%  # 651 rows
  filter(!is.na(genome_nt_pos_noncoding)) %>%
  select(
    genome_nt_mut_pos,
    genome_nt_mut_sub,
    genome_nt_mut_wt,
    genome_mut_ID_nt,
    genome_nt_pos_structural,
    genome_nt_pos_overlap_orf,
    genome_nt_pos_coding,
    fitness_final,
    sigma_final,
    phix_aa_mut_genes,
    genome_nt_pos_ori,
    genome_nt_pos_RBS,
    genome_nt_pos_RBS_gene,
    genome_nt_pos_promoter,
    genome_nt_pos_promoter_gene,
    genome_nt_pos_promoter_anti_sense,
    genome_nt_pos_terminator,
    genome_nt_pos_terminator_gene,
    genome_nt_pos_annotation,
    genome_nt_pos_annotation_combined
  ) %>%
  mutate(genome_nt_mut_pos = as.numeric(genome_nt_mut_pos))

table(phix_dms_sub_unique_1nt_noncoding_cols$phix_aa_mut_genes)
# Noncoding#1 J-F Noncoding#2 F-G Noncoding#3 G-H Noncoding#4 H-A 
#             108             325              24             179 


# Subset different Noncoding data.frames 
phix_dms_sub_unique_1nt_noncoding1_cols <- phix_dms_sub_unique_1nt_noncoding_cols %>%
  filter(phix_aa_mut_genes == "Noncoding#1 J-F")
phix_dms_sub_unique_1nt_noncoding2_cols <- phix_dms_sub_unique_1nt_noncoding_cols %>%
  filter(phix_aa_mut_genes == "Noncoding#2 F-G")
phix_dms_sub_unique_1nt_noncoding3_cols <- phix_dms_sub_unique_1nt_noncoding_cols %>%
  filter(phix_aa_mut_genes == "Noncoding#3 G-H")
phix_dms_sub_unique_1nt_noncoding4_cols <- phix_dms_sub_unique_1nt_noncoding_cols %>%
  filter(phix_aa_mut_genes == "Noncoding#4 H-A")


# Plotting Noncoding Heatmaps
heatmap_non1 <- plot_heatmap_noncoding_colored_wt(phix_dms_sub_unique_1nt_noncoding1_cols, 
                                       plot_title = "Fitness Heatmap - Noncoding region between gene J-F",
                                       show_legend = FALSE, show_title = FALSE, wt_value = 0.9807835,wt_shape = 4)
print(heatmap_non1) # 36 positions
ggsave(heatmap_non1, filename = "./plots/heatmap_non1_substitutions.pdf", width = 4, height = 2)

heatmap_non2 <- plot_heatmap_noncoding_colored_wt(phix_dms_sub_unique_1nt_noncoding2_cols, 
                                       plot_title = "Fitness Heatmap - Noncoding region between gene F-G",
                                       show_legend = FALSE, show_title = FALSE, wt_value = 0.9807835,wt_shape = 4)
print(heatmap_non2) # 110 positions
ggsave(heatmap_non2, filename = "./plots/heatmap_non2_substitutions.pdf", width = 12, height = 2)

heatmap_non3 <- plot_heatmap_noncoding_colored_wt(phix_dms_sub_unique_1nt_noncoding3_cols, 
                                       plot_title = "Fitness Heatmap - Noncoding region between gene G-H",
                                       show_legend = FALSE, show_title = FALSE, wt_value = 0.9807835,wt_shape = 4)
print(heatmap_non3) # 8 positions
ggsave(heatmap_non3, filename = "./plots/heatmap_non3_substitutions.pdf", width = 1.5, height = 2)

heatmap_non4 <- plot_heatmap_noncoding_colored_wt(phix_dms_sub_unique_1nt_noncoding4_cols, 
                                       plot_title = "Fitness Heatmap - Noncoding region between gene H-A",
                                       show_legend = TRUE, show_title = FALSE, wt_value = 0.9807835,wt_shape = 4)
print(heatmap_non4) # 63 positions 
ggsave(heatmap_non4, filename = "./plots/heatmap_non4_substitutions.pdf", width = 9, height = 2)



```



### 5) Circular genomic heatmap 
Outside annotation - add Heatmap - add residue axis

#### 1) Coding regions
```{r}
## 
# Load necessary libraries
#library(circlize)

# 1. Build genome annotation data frame
phix_genome_length <- 5386
phix_coding <- data.frame(
  pos = 1:phix_genome_length,
  coding = 0,
  overlapping = 0
)

# 2. Annotate coding regions
phix_coding$coding[phix_coding$pos >= 965 & phix_coding$pos <= 1000] <- 1
phix_coding$coding[phix_coding$pos >= 2285 & phix_coding$pos <= 2394] <- 1
phix_coding$coding[phix_coding$pos >= 2923 & phix_coding$pos <= 2930] <- 1
phix_coding$coding[phix_coding$pos >= 3918 & phix_coding$pos <= 3980] <- 1

# 3. Annotate overlapping regions
phix_coding$overlapping[phix_coding$pos < 222] <- 2
phix_coding$overlapping[phix_coding$pos > 389 & phix_coding$pos < 394] <- 2
phix_coding$overlapping[phix_coding$pos > 567 & phix_coding$pos < 844] <- 2
phix_coding$overlapping[phix_coding$pos == 848] <- 2
phix_coding$overlapping[phix_coding$pos > 5074] <- 2

# Quick checks (optional)
message("Coding summary:"); print(table(phix_coding$coding))
message("Overlapping summary:"); print(table(phix_coding$overlapping))

# 4. Prepare for circos heatmap
phix_coding_matrix <- as.matrix(phix_coding[, c("coding", "overlapping")])
rownames(phix_coding_matrix) <- phix_coding$pos

# 5. Plot circular heatmap
circos.clear()
circos.par(start.degree = 90, gap.degree = 0)
col_phix_coding <- colorRamp2(c(0, 1, 2), c("white", "gray", "black"))

circos.heatmap(
  phix_coding_matrix,
  cluster = FALSE,
  col = col_phix_coding,
  track.height = 0.05
)



```


#### 2) 1nt Mut Fitness Heatmaps
```{r}
## Subset columns from the 1nt substitution data for circular heatmap
#head(as.data.frame(phix_all_dms_sub_unique_1nt_plotting), 0)
head(as.data.frame(phix_all_dms_sub_unique_1nt_annotated), 0)
phix_dms_sub_unique_1nt_circle <- phix_all_dms_sub_unique_1nt_pos_annotated %>%
  select(genome_nt_mut_pos1, genome_nt_mut_sub, fitness_final)
range(phix_dms_sub_unique_1nt_circle$fitness_final)

# Reshape and convert to matrix for circular heatmap
phix_matrix <- phix_dms_sub_unique_1nt_circle %>%
  #filter(genome_nt_mut_pos1 %in% c(3918:5386)) %>%
  dcast(genome_nt_mut_pos1 ~ genome_nt_mut_sub, value.var = "fitness_final") %>%
  arrange(genome_nt_mut_pos1) %>%
  column_to_rownames("genome_nt_mut_pos1") %>%
  as.matrix()

# Color scale: darkred → red → gray88 → blue
col_phix <- colorRamp2(c(-1, 0, 1, 1.5), c("darkred", "red", "gray88", "blue"))
col_phix <- colorRamp2(
  breaks = c(-1, 0, 1, 1.5),
  colors = c("darkred", "red", "gray88", "blue")
)

# Set circos layout parameters
circos.par(start.degree = 90, gap.degree = 0)

# Plot circular heatmap
circos.heatmap(
  phix_matrix,
  cluster = FALSE,
  col = col_phix
)

# Add genome coordinate axis every 500 nt
circos.axis(
  h = "bottom",
  major.at = seq(0, 5386, by = 500),
  labels = seq(0, 5386, by = 500),
  labels.cex = 0.6,
  labels.facing = "inside",
  direction = "inside",
  minor.ticks = 0,
  labels.pos.adjust = TRUE
)

#Add correct legend with matching limits
lgd <- Legend(
  title = "Fitness",
  col_fun = col_phix,
  at = c(-1, 0, 1, 1.5),  # Explicit breakpoints
  labels = c("-1", "0", "1", "1.5"),
  direction = "horizontal"
)
grid.newpage()
draw(lgd)

circos.clear()


```


#### 3) ORF annotations
```{r}
library(circlize)

# 1. Clear any existing plot and define gene annotations
circos.clear()

phix_genome_annotation <- data.frame(
  phix_genome_gene = factor(
    c("A", "D", "F", "G", "H", "A*", "K", "E", "J", "B", "C"),
    levels = c("A", "D", "F", "G", "H", "A*", "K", "E", "J", "B", "C")
  ),
  phix_gene_end = c(1542, 459, 1284, 528, 987, 1026, 171, 276, 117, 363, 261)
)

phix_genome_annotation$sector <- "s"
phix_genome_annotation$sector[c(1:5)]    <- "s1"  # A, D, F, G, H
phix_genome_annotation$sector[c(6, 8:9)] <- "s2"  # A*, E, J
phix_genome_annotation$sector[c(10:11)]  <- "s3"  # B, C
phix_genome_annotation$sector[7]         <- "s4"  # K

xlim_list <- list(
  s1 = cbind(0, phix_genome_annotation$phix_gene_end[phix_genome_annotation$sector == "s1"]),
  s2 = cbind(0, phix_genome_annotation$phix_gene_end[phix_genome_annotation$sector == "s2"]),
  s3 = cbind(0, phix_genome_annotation$phix_gene_end[phix_genome_annotation$sector == "s3"]),
  s4 = cbind(0, phix_genome_annotation$phix_gene_end[phix_genome_annotation$sector == "s4"])
)

# Track 1: Initialize and plot for s1
# 1st track: A, D, F, G, H;
circos.par(start.degree =184, gap.degree = c(14.4, 10.1, 7.35, 0.5, 4.2)) # Group#1: , gap.degree = c(14.4, 2.4, 7.35, 0.5, 4.2)
circos.initialize(factors = phix_genome_annotation$phix_genome_gene[phix_genome_annotation$sector == "s1"], xlim = xlim_s1)
circos.track(ylim = c(0, 1), panel.fun = function(x, y) {
  circos.arrow(CELL_META$xlim[1], CELL_META$xlim[2], 
               arrow.head.width = CELL_META$yrange * 0.5, arrow.head.length = cm_x(0.2), 
               col = "#66C2A5")
  circos.text(CELL_META$xcenter, CELL_META$ycenter, CELL_META$sector.index, 
              facing = "downward")
}, bg.border = NA, track.height = 0.25)

# Track 2: Initialize and plot for s2
# 2nd track: A*, E, J;
circos.par(start.degree =150, cell.padding = c(0.02, 0, 0.02, 0), gap.degree = c(28.8, 0.5, 236)) # Group#2: , gap.degree = c(47.5, 236)
circos.initialize(factors = phix_genome_annotation$phix_genome_gene[phix_genome_annotation$sector == "s2"], xlim = xlim_s2)
circos.track(ylim = c(0, 1), panel.fun = function(x, y) {
  circos.arrow(CELL_META$xlim[1], CELL_META$xlim[2], 
               arrow.head.width = CELL_META$yrange * 0.5, arrow.head.length = cm_x(0.2), 
               col = "#66C2A5")
  circos.text(CELL_META$xcenter, CELL_META$ycenter, CELL_META$sector.index, 
              facing = "downward")
}, bg.border = NA, track.height = 0.25)

# Track 3: Initialize and plot for s3
# 3rd track: B, C;
circos.par(start.degree =110.8, cell.padding = c(0.02, 0, 0.02, 0), gap.degree = c(5.4, 312.8)) # Group#3: , gap.degree = c(5.4, 312.8)
circos.initialize(factors = phix_genome_annotation$phix_genome_gene[phix_genome_annotation$sector == "s3"], xlim = xlim_s3)
circos.track(ylim = c(0, 1), panel.fun = function(x, y) {
  circos.arrow(CELL_META$xlim[1], CELL_META$xlim[2], 
               arrow.head.width = CELL_META$yrange*0.5, arrow.head.length = cm_x(0.2), 
               col = "#66C2A5")
  circos.text(CELL_META$xcenter, CELL_META$ycenter, CELL_META$sector.index, 
              facing = "downward")
}, bg.border = NA, track.height = 0.25)

# Track 4: Initialize and plot for s4
# 4th track: K;
circos.par(start.degree =87, gap.degree = c(348.5)) # Group#3: , gap.degree = c(23.1, 306.9)
circos.initialize(factors = phix_genome_annotation$phix_genome_gene[phix_genome_annotation$sector == "s4"], xlim = xlim_s4)
circos.track(ylim = c(0, 1), panel.fun = function(x, y) {
  circos.arrow(CELL_META$xlim[1], CELL_META$xlim[2], 
               arrow.head.width = CELL_META$yrange*0.5, arrow.head.length = cm_x(0.2), 
               col = "#66C2A5")
  circos.text(CELL_META$xcenter, CELL_META$ycenter, CELL_META$sector.index, 
              facing = "downward")
}, bg.border = NA, track.height = 0.25)

# Done
circos.clear()


```




## Save script checkpoint
save a snapshot of all objects produced by the script

```{r}
# checkpoint - 03 statistics, summary of mutation category, and make plots
# save.image(file = "data/checkpoints/03_single_nucleotide_variants.RData")


```
















