---
title: "phix_proteins_AF3_modelling_data_analysis"
output: html_document
date: "2026-05-03"
editor_options: 
  chunk_output_type: console
---

## Set working working directory
```{r}
here::here()

```



## Load Functions

### 1) process atom distance data
```{r}
# Function to parse contact files
parse_contact_file <- function(file_path, structure_file, gene_label) {
  
  df_raw <- read_table(
    file_path,
    col_names = FALSE,
    col_types = cols(.default = "c")
  )
  
  df_raw <- df_raw %>%
    rename(
      atom1 = X1,
      atom2 = X2,
      distance = X3
    ) %>%
    mutate(
      distance = as.numeric(distance)
    )
  
  df_parsed <- df_raw %>%
    mutate(
      chain1 = str_extract(atom1, "(?<=/)[^:]+"),
      res1   = as.integer(str_extract(atom1, "(?<=:)[0-9]+")),
      atom1_name = str_extract(atom1, "(?<=@).+"),
      
      chain2 = str_extract(atom2, "(?<=/)[^:]+"),
      res2   = as.integer(str_extract(atom2, "(?<=:)[0-9]+")),
      atom2_name = str_extract(atom2, "(?<=@).+")
    ) %>%
    filter(chain1 != chain2) %>%
    mutate(
      structure_file = structure_file,
      gene = gene_label
    )
  
  return(df_parsed)
}


# Function to read model confidence from .json files
read_AF3_summary_jsons <- function(root_path) {
  
  files <- list.files(
    root_path,
    #pattern = "_summary\\.json$",
    pattern = "_summary_confidences\\.json$",
    full.names = TRUE,
    recursive = TRUE
  )
  
  if (!length(files)) stop("No *_summary.json files found.")
  
  parse_one <- function(f) {
    
    file_name <- basename(f)
    
    # Remove "_summary.json" to get model name
    model_name <- sub("_summary\\.json$", "", file_name)
    
    j <- try(jsonlite::fromJSON(f), silent = TRUE)
    if (inherits(j, "try-error")) j <- list()
    
    tibble::tibble(
      model = model_name,
      iptm  = if (!is.null(j$iptm)) as.numeric(j$iptm) else NA_real_,
      ptm   = if (!is.null(j$ptm))  as.numeric(j$ptm)  else NA_real_,
      ranking_score =
        if (!is.null(j$ranking_score)) as.numeric(j$ranking_score)
        else if (!is.null(j$aggregate_score)) as.numeric(j$aggregate_score)
        else NA_real_,
      has_clash =
        if (!is.null(j$has_clash)) j$has_clash
        else if (!is.null(j$has_inter_chain_clashes)) j$has_inter_chain_clashes
        else NA,
      fraction_disordered =
        if (!is.null(j$fraction_disordered)) as.numeric(j$fraction_disordered)
        else NA_real_,
      num_recycles =
        if (!is.null(j$num_recycles)) as.numeric(j$num_recycles)
        else NA_real_
    )
  }
  
  purrr::map_dfr(files, parse_one)
}


# Function to extract chain-level and pair-level AlphaFold 3 metrics
read_AF3_chain_metrics <- function(root_path) {
  
  files <- list.files(
    root_path,
    pattern = "_summary_confidences\\.json$",
    full.names = TRUE,
    recursive = TRUE
  )
  
  if (length(files) == 0) {
    stop("No *_summary_confidences.json files found.")
  }
  
  parse_one <- function(f) {
    
    file_name <- basename(f)
    group_name <- str_remove(file_name, "_summary_confidences\\.json$")
    
    j <- try(fromJSON(f), silent = TRUE)
    if (inherits(j, "try-error")) return(NULL)
    
    # -------------------------
    # Model-level
    # -------------------------
    
    model_df <- tibble(
      group = group_name,
      iptm  = if (!is.null(j$iptm)) as.numeric(j$iptm) else NA_real_,
      ptm   = if (!is.null(j$ptm))  as.numeric(j$ptm)  else NA_real_,
      ranking_score =
        if (!is.null(j$ranking_score)) as.numeric(j$ranking_score)
        else if (!is.null(j$aggregate_score)) as.numeric(j$aggregate_score)
        else NA_real_
    )
    
    # -------------------------
    # Chain-level
    # -------------------------
    
    chain_df <- NULL
    
    if (!is.null(j$chain_ptm)) {
      
      chain_df <- tibble(
        group = group_name,
        chain_index = seq_along(j$chain_ptm),
        chain_ptm = as.numeric(j$chain_ptm),
        chain_iptm =
          if (!is.null(j$chain_iptm))
            as.numeric(j$chain_iptm)
          else NA_real_
      )
    }
    
    # -------------------------
    # Pair-level
    # -------------------------
    
    pair_df <- NULL
    
    if (!is.null(j$chain_pair_iptm)) {
      
      iptm_mat <- as.matrix(j$chain_pair_iptm)
      pae_mat  <- as.matrix(j$chain_pair_pae_min)
      
      iptm_long <- as.data.frame(iptm_mat) %>%
        mutate(chain_i = row_number()) %>%
        pivot_longer(
          -chain_i,
          names_to = "chain_j",
          values_to = "pair_iptm"
        ) %>%
        mutate(chain_j = as.integer(str_remove(chain_j, "V")))
      
      pae_long <- as.data.frame(pae_mat) %>%
        mutate(chain_i = row_number()) %>%
        pivot_longer(
          -chain_i,
          names_to = "chain_j",
          values_to = "pair_pae_min"
        ) %>%
        mutate(chain_j = as.integer(str_remove(chain_j, "V")))
      
      pair_df <- left_join(
        iptm_long,
        pae_long,
        by = c("chain_i", "chain_j")
      ) %>%
        mutate(group = group_name) %>%
        select(group, chain_i, chain_j,
               pair_iptm, pair_pae_min)
    }
    
    list(
      model = model_df,
      chain = chain_df,
      pair  = pair_df
    )
  }
  
  parsed <- map(files, parse_one)
  
  list(
    model_level = bind_rows(map(parsed, "model")),
    chain_level = bind_rows(map(parsed, "chain")),
    pair_level  = bind_rows(map(parsed, "pair"))
  )
}



# Function to map gene names to chain indexes
# 1) map pTM scores to chains
map_chain_confidence <- function(chain_level_df, chain_roles_df) {

  chain_level_df %>%
    left_join(
      chain_roles_df %>%
        select(group, chain_index, gene, chain_id),
      by = c("group", "chain_index")
    )
}


# 2) map pairwise ipTM scores
map_pairwise_confidence <- function(pair_df, chain_roles_df) {

  pair_df %>%
    left_join(
      chain_roles_df %>%
        select(group, chain_index, gene),
      by = c("group", "chain_i" = "chain_index")
    ) %>%
    rename(gene_i = gene) %>%
    left_join(
      chain_roles_df %>%
        select(group, chain_index, gene),
      by = c("group", "chain_j" = "chain_index")
    ) %>%
    rename(gene_j = gene)
}



# Function to read ipSAE scores
read_ipSAE_results <- function(root_path,
                               pattern = "_\\d+_\\d+\\.txt$") {
  
  library(dplyr)
  library(stringr)
  library(purrr)
  library(readr)
  library(tibble)
  
  files <- list.files(
    root_path,
    pattern = pattern,
    full.names = TRUE,
    recursive = TRUE
  )
  
  if (length(files) == 0) {
    stop("No ipSAE result files found.")
  }
  
  parse_one <- function(f) {
    
    file_name <- basename(f)
    
    # Extract cutoff values
    cutoff_match <- str_match(file_name, "_(\\d+)_(\\d+)\\.txt$")
    cutoff1 <- as.integer(cutoff_match[,2])
    cutoff2 <- as.integer(cutoff_match[,3])
    
    # Extract model group (remove _15_15.txt)
    group <- str_remove(file_name, "_\\d+_\\d+\\.txt$")
    
    # Read whitespace-delimited table
    df <- read_table(f, show_col_types = FALSE)
    
    df %>%
      mutate(
        group = group,
        cutoff1 = cutoff1,
        cutoff2 = cutoff2,
        source_file = file_name
      ) %>%
      select(group, cutoff1, cutoff2, everything())
  }
  
  map_dfr(files, parse_one)
}




# Function to get model name or ids for specific interactions
get_models_with_partner <- function(residue_contacts_df, partner_label) {
  residue_contacts_df %>%
    dplyr::filter(!is.na(partner_gene),
                  partner_gene == partner_label) %>%
    dplyr::pull(model_name) %>%
    unique()
}



```


### 2) annotate and process fitness combined data
```{r}
# Function to parse model names at both model and hypothesis levels 
# 1) parsing model names in column 'group'
parse_model_group <- function(df, group_col = "group") {
  
  library(dplyr)
  library(stringr)
  
  df %>%
    mutate(
      hypothesis = str_remove(.data[[group_col]],
                              "_seed[0-9]+_model_[0-9]+$"),
      
      seed = as.integer(str_extract(.data[[group_col]],
                                    "(?<=_seed)[0-9]+")),
      
      model = as.integer(str_extract(.data[[group_col]],
                                     "(?<=_model_)[0-9]+")),
      
      system = str_extract(hypothesis, "^[^_]+"),
      rest   = str_remove(hypothesis, "^[^_]+_"),
      
      nt = str_extract(rest, "[0-9]+nt"),
      nt_numeric = as.integer(str_remove(nt, "nt"))
    ) %>%
    
    mutate(
      dna_type = case_when(
        str_detect(rest, "^noDNA") ~ "noDNA",
        !is.na(nt) ~ str_remove(rest, paste0("_?", nt, ".*$")),
        TRUE ~ NA_character_
      ),
      
      modifier = case_when(
        str_detect(rest, "^noDNA") ~ str_remove(rest, "^noDNA_?"),
        !is.na(nt) ~ str_remove(rest, paste0(".*", nt, "_?")),
        TRUE ~ NA_character_
      ),
      
      modifier = na_if(modifier, "")
    ) %>%
    
    mutate(
      modifier_plot = case_when(
        is.na(modifier) ~ "DNA-only",
        modifier == "2rep" ~ "+ 2Rep",
        modifier == "gpc" ~ "+ C",
        modifier == "2rep_gpc" ~ "+ 2Rep + C",
        TRUE ~ modifier
      ),
      
      system_short = case_when(
        system == "gpa" ~ "A",
        system == "gpc" ~ "C",
        TRUE ~ system
      ),
      
      hypothesis_short = ifelse(
        is.na(modifier),
        paste0(system_short, "-", dna_type),
        paste0(system_short, "-", dna_type, "-", modifier)
      ),
      
      group_short = paste0(
        hypothesis_short,
        "-s", seed,
        "-m", model
      )
    ) %>%
    
    select(-rest)
}


# 2) function to get interaction information from the column names of computational scores
parse_confidence_feature <- function(df) {
  
  genes <- c("A", "C", "DNA", "MG", "rep")
  
  df %>%
    mutate(
      parts = strsplit(feature, "_"),
      n = lengths(parts),
      
      last = sapply(parts, function(x) x[length(x)]),
      second_last = sapply(parts, function(x) {
        if (length(x) >= 2) x[length(x)-1] else NA_character_
      }),
      
      is_pair  = last %in% genes & second_last %in% genes,
      is_chain = last %in% genes & !is_pair,
      
      interaction = ifelse(
        is_pair,
        paste(second_last, last, sep = "-"),
        NA_character_
      ),
      
      gene = ifelse(
        is_chain,
        last,
        NA_character_
      ),
      
      metric = case_when(
        is_pair  ~ sapply(parts, function(x) paste(x[1:(length(x)-2)], collapse = "_")),
        is_chain ~ sapply(parts, function(x) paste(x[1:(length(x)-1)], collapse = "_")),
        TRUE ~ feature
      ),
      
      level = case_when(
        is_pair  ~ "pair",
        is_chain ~ "chain",
        TRUE ~ "other"
      )
    ) %>%
    select(-parts, -n, -last, -second_last, -is_pair, -is_chain)
}


# 3) parsing model names 
parse_model_names <- function(model_names) {
  
  #library(dplyr)
  #library(stringr)
  #library(tibble)
  
  tibble(model_name = model_names) %>%
    
    # ---- Extract hypothesis ----
    mutate(
      hypothesis = str_remove(
        model_name,
        "_seed\\d+_model_\\d+$"
      )
    ) %>%
    
    # ---- Extract seed and model (if present) ----
    mutate(
      seed = as.integer(str_extract(model_name, "(?<=seed)\\d+")),
      af3_model = as.integer(str_extract(model_name, "(?<=model_)\\d+$"))
    ) %>%
    
    # ---- Parse hypothesis components ----
    mutate(
      parts = str_split(hypothesis, "_"),
      system = sapply(parts, `[`, 1),
      nt = str_extract(hypothesis, "[0-9]+nt"),
      nt_numeric = as.integer(str_remove(nt, "nt")),
      
      dna_type = sapply(parts, function(x) {
        nt_index <- which(str_detect(x, "[0-9]+nt"))
        if (length(nt_index) == 1 && nt_index > 2) {
          paste(x[2:(nt_index - 1)], collapse = "_")
        } else {
          x[2]
        }
      }),
      
      modifier = sapply(parts, function(x) {
        nt_index <- which(str_detect(x, "[0-9]+nt"))
        if (length(nt_index) == 1 && nt_index < length(x)) {
          paste(x[(nt_index + 1):length(x)], collapse = "_")
        } else {
          NA_character_
        }
      })
    ) %>%
    select(-parts)
}


# 4) parsing hypothesis names
parse_hypothesis_names <- function(.data,
                                   hypothesis_col = "hypothesis") {
  
  #library(dplyr)
  #library(stringr)
  
  .data %>%
    mutate(
      parts = str_split(.data[[hypothesis_col]], "_"),
      
      system = sapply(parts, `[`, 1),
      
      nt = str_extract(.data[[hypothesis_col]], "[0-9]+nt"),
      nt_numeric = as.integer(str_remove(nt, "nt")),
      
      dna_type = sapply(parts, function(x) {
        nt_index <- which(str_detect(x, "[0-9]+nt"))
        if (length(nt_index) == 1 && nt_index > 2) {
          paste(x[2:(nt_index - 1)], collapse = "_")
        } else {
          x[2]
        }
      }),
      
      modifier = sapply(parts, function(x) {
        nt_index <- which(str_detect(x, "[0-9]+nt"))
        if (length(nt_index) == 1 && nt_index < length(x)) {
          paste(x[(nt_index + 1):length(x)], collapse = "_")
        } else {
          NA_character_
        }
      })
    ) %>%
    select(-parts)
}


# 5) Function to parse the column names with enrichment score 
parse_enrichment_feature <- function(df) {
  
  library(dplyr)
  library(stringr)
  
  pattern <- "^(.*)_([A-Za-z]+)_(res|mut)_interact_([A-Za-z]+)$"
  
  df %>%
    mutate(
      metric_prefix = str_match(feature, pattern)[,2],
      gene          = str_match(feature, pattern)[,3],
      level         = str_match(feature, pattern)[,4],
      partner       = str_match(feature, pattern)[,5],
      
      interaction = ifelse(
        !is.na(gene) & !is.na(partner),
        paste(gene, partner, sep = "-"),
        NA_character_
      )
    )
}




# Function to assemble annotated Fitness data and summarise frequently predicted residues
assemble_residue_long <- function(annotated_list,
                                  residue_col = "aa_mut_pos",
                                  interact_col = "interact",
                                  essential_col = "essential_flag",
                                  wt_col = "aa_mut_wt",
                                  models_keep = NULL) {
  
  if (!is.null(models_keep)) {
    annotated_list <- annotated_list[names(annotated_list) %in% models_keep]
  }
  
  df <- bind_rows(annotated_list, .id = "model_name") %>%
    select(
      model_name,
      residue = all_of(residue_col),
      interact = all_of(interact_col),
      essential = all_of(essential_col),
      wt = all_of(wt_col)
    ) %>%
    mutate(
      interact = as.integer(interact),
      essential = as.integer(essential)
    )
  
  parsed <- parse_model_names(unique(df$model_name))
  
  df %>%
    left_join(parsed, by = "model_name")
}



# Function to calculate residue frequency and remove non-interacting residues
compute_residue_frequency <- function(residue_long_df,
                                      group_vars = c("hypothesis", "residue", "wt"),
                                      remove_never_interacting = TRUE) {
  
  #library(dplyr)
  
  freq_df <- residue_long_df %>%
    
    group_by(across(all_of(c(group_vars, "residue", "wt")))) %>%
    
    summarise(
      n_models = n(),
      freq_interact = mean(interact),
      freq_essential_interact = mean(interact * essential),
      .groups = "drop"
    )
  
  if (remove_never_interacting) {
    
    residues_keep <- freq_df %>%
      group_by(residue) %>%
      summarise(max_freq = max(freq_interact), .groups = "drop") %>%
      filter(max_freq > 0) %>%
      pull(residue)
    
    freq_df <- freq_df %>%
      filter(residue %in% residues_keep)
  }
  
  freq_df
}



# Function to annotate A-C interaction predictions
add_hypothesis_short_plot <- function(
  df,
  hypothesis_col = "hypothesis_short",
  new_col = "hypothesis_short_plot"
) {
  
  df %>%
    dplyr::mutate(
      !!new_col := dplyr::case_when(
        .data[[hypothesis_col]] == "A-noDNA-gpc" ~ "A-C",
        .data[[hypothesis_col]] == "A-ssDNA-gpc" ~ "A-C-DNA",
        .data[[hypothesis_col]] == "A-ssDNA-2rep_gpc" ~ "A-C-DNA-Rep",
        TRUE ~ .data[[hypothesis_col]]
      )
    )
}



# Function to summarize model confidence
summarise_confidence_column <- function(
  df,
  value_col,
  group_col = NULL
) {
  
  library(dplyr)
  
  if (is.null(group_col)) {
    
    df %>%
      summarise(
        metric = value_col,
        n = sum(!is.na(.data[[value_col]])),
        min = min(.data[[value_col]], na.rm = TRUE),
        max = max(.data[[value_col]], na.rm = TRUE),
        median = median(.data[[value_col]], na.rm = TRUE),
        mean = mean(.data[[value_col]], na.rm = TRUE),
        sd = sd(.data[[value_col]], na.rm = TRUE)
      )
    
  } else {
    
    df %>%
      group_by(.data[[group_col]]) %>%
      summarise(
        metric = value_col,
        n = sum(!is.na(.data[[value_col]])),
        min = min(.data[[value_col]], na.rm = TRUE),
        max = max(.data[[value_col]], na.rm = TRUE),
        median = median(.data[[value_col]], na.rm = TRUE),
        mean = mean(.data[[value_col]], na.rm = TRUE),
        sd = sd(.data[[value_col]], na.rm = TRUE),
        .groups = "drop"
      )
  }
}




```


### 3) process fitness data
```{r}
# Function to run Odds Ratio Calculation
# 1) from annotated fitness datasets
run_odds_ratio <- function(df,
                           exposure_col,
                           exposure_focus,
                           outcome_col,
                           outcome_focus,
                           dataset = NA,
                           group = NA) {

  exposure_sym <- rlang::ensym(exposure_col)
  outcome_sym  <- rlang::ensym(outcome_col)

  exposure_name <- rlang::as_name(exposure_sym)
  outcome_name  <- rlang::as_name(outcome_sym)

  dat <- df %>%
    dplyr::filter(!is.na(.data[[exposure_name]]),
                  !is.na(.data[[outcome_name]])) %>%
    dplyr::mutate(
      exposure = .data[[exposure_name]] == exposure_focus,
      outcome  = .data[[outcome_name]] %in% outcome_focus
    )

  if (nrow(dat) == 0) {
    stop("No complete cases available.")
  }

  tab <- table(dat$exposure, dat$outcome)

  # If not 2x2, return NA row instead of stopping
  if (!all(dim(tab) == c(2, 2))) {
    return(tibble::tibble(
      test_type   = "fisher_odds_ratio",
      dataset     = dataset,
      group       = group,
      exposure_var= exposure_name,
      outcome_var = outcome_name,
      exposure_focus = exposure_focus,
      outcome_focus  = paste(outcome_focus, collapse = ","),
      n_total     = sum(tab),
      a_exposed_outcome     = NA_integer_,
      b_exposed_no_outcome  = NA_integer_,
      c_unexp_outcome       = NA_integer_,
      d_unexp_no_outcome    = NA_integer_,
      odds_ratio  = NA_real_,
      conf_low    = NA_real_,
      conf_high   = NA_real_,
      log_or_ln   = NA_real_,
      log_or_log10= NA_real_,
      log_conf_low_ln    = NA_real_,
      log_conf_high_ln   = NA_real_,
      log_conf_low_log10 = NA_real_,
      log_conf_high_log10= NA_real_,
      p_value     = NA_real_,
      method      = "Not 2x2 — skipped"
    ))
  }

  ft <- fisher.test(tab)

  # Extract once
  or_val  <- unname(ft$estimate)
  ci_low  <- ft$conf.int[1]
  ci_high <- ft$conf.int[2]

  tibble::tibble(
    test_type   = "fisher_odds_ratio",
    dataset     = dataset,
    group       = group,
    exposure_var= exposure_name,
    outcome_var = outcome_name,
    exposure_focus = exposure_focus,
    outcome_focus  = paste(outcome_focus, collapse = ","),
    n_total     = sum(tab),
    a_exposed_outcome     = tab[2, 2],
    b_exposed_no_outcome  = tab[2, 1],
    c_unexp_outcome       = tab[1, 2],
    d_unexp_no_outcome    = tab[1, 1],
    odds_ratio  = or_val,
    conf_low    = ci_low,
    conf_high   = ci_high,

    # Natural log
    log_or_ln   = log(or_val),
    log_conf_low_ln  = log(ci_low),
    log_conf_high_ln = log(ci_high),

    # Log10
    log_or_log10     = log10(or_val),
    log_conf_low_log10  = log10(ci_low),
    log_conf_high_log10 = log10(ci_high),

    p_value     = ft$p.value,
    method      = ft$method
  )
}

# 2) from combined counts
run_odds_ratio_from_counts <- function(a, b, c, d,
                                       dataset = NA,
                                       group = NA) {
  
  library(tibble)
  
  # Build table correctly:
  # rows = exposure (FALSE, TRUE)
  # cols = outcome (FALSE, TRUE)
  
  tab <- matrix(
    c(d, c,
      b, a),
    nrow = 2,
    byrow = TRUE
  )
  
  ft <- fisher.test(tab)
  
  or_val  <- unname(ft$estimate)
  ci_low  <- ft$conf.int[1]
  ci_high <- ft$conf.int[2]
  
  tibble(
    test_type   = "fisher_odds_ratio_combined",
    dataset     = dataset,
    group       = group,
    n_total     = sum(tab),
    a_exposed_outcome     = a,
    b_exposed_no_outcome  = b,
    c_unexp_outcome       = c,
    d_unexp_no_outcome    = d,
    odds_ratio  = or_val,
    conf_low    = ci_low,
    conf_high   = ci_high,
    
    log_or_ln   = log(or_val),
    log_conf_low_ln  = log(ci_low),
    log_conf_high_ln = log(ci_high),
    
    log_or_log10     = log10(or_val),
    log_conf_low_log10  = log10(ci_low),
    log_conf_high_log10 = log10(ci_high),
    
    p_value     = ft$p.value,
    method      = ft$method
  )
}



# Function to have a wrapper calculate enrichment by Odds Ratio
compute_enrichment <- function(df,
                                           exposure_col,
                                           exposure_focus,
                                           outcome_col,
                                           outcome_focus,
                                           dataset_label,
                                           model_id,
                                           suffix) {
  
  res <- run_odds_ratio(
    df = df,
    exposure_col = {{ exposure_col }},
    exposure_focus = exposure_focus,
    outcome_col = {{ outcome_col }},
    outcome_focus = outcome_focus,
    dataset = dataset_label,
    group = model_id
  )
  
  res %>%
    dplyr::rename(
      !!paste0("odds_ratio_", suffix) := odds_ratio,
      !!paste0("log_or_ln_", suffix) := log_or_ln,
      !!paste0("log_or_log10_", suffix) := log_or_log10,
      !!paste0("conf_low_", suffix) := conf_low,
      !!paste0("conf_high_", suffix) := conf_high,
      !!paste0("log_conf_low_ln_", suffix) := log_conf_low_ln,
      !!paste0("log_conf_high_ln_", suffix) := log_conf_high_ln,
      !!paste0("log_conf_low_log10_", suffix) := log_conf_low_log10,
      !!paste0("log_conf_high_log10_", suffix) := log_conf_high_log10,
      !!paste0("p_value_", suffix) := p_value,
      !!paste0("n_total_", suffix) := n_total,
      !!paste0("a_exposed_outcome_", suffix) := a_exposed_outcome,
      !!paste0("b_exposed_no_outcome_", suffix) := b_exposed_no_outcome,
      !!paste0("c_unexp_outcome_", suffix) := c_unexp_outcome,
      !!paste0("d_unexp_no_outcome_", suffix) := d_unexp_no_outcome
    )
}



# Function to combine the calculated enrichment scores by models
combine_enrichment_tables <- function(..., join_by = "group") {
  
  tables <- list(...)
  
  # function to keep only join key + suffixed metric columns
  clean_one <- function(df) {
    df %>%
      dplyr::select(
        dplyr::all_of(join_by),
        dplyr::starts_with("n_total_"),
        dplyr::starts_with("a_exposed_outcome_"),
        dplyr::starts_with("b_exposed_no_outcome_"),
        dplyr::starts_with("c_unexp_outcome_"),
        dplyr::starts_with("d_unexp_no_outcome_"),
        dplyr::starts_with("odds_ratio_"),
        dplyr::starts_with("conf_low_"),
        dplyr::starts_with("conf_high_"),
        dplyr::starts_with("log_or_"),
        dplyr::starts_with("log_conf_"),
        dplyr::starts_with("p_value_")
      )
  }
  
  tables_clean <- lapply(tables, clean_one)
  
  # iterative left join
  Reduce(function(x, y) {
    dplyr::left_join(x, y, by = join_by)
  }, tables_clean)
}


# Function to annotate predicted contacts
annotate_fitness_with_contacts <- function(fitness_df,
                                           residue_contacts_df,
                                           model_id) {
  
  interacting_residues <- residue_contacts_df %>%
    dplyr::filter(model_name == model_id) %>%
    dplyr::pull(residue) %>%
    unique()
  
  fitness_df %>%
    dplyr::mutate(
      interact = aa_mut_pos %in% interacting_residues,
      model = model_id
    )
}


# Function to annotate predicted contacts by interaction partners
annotate_fitness_with_partner <- function(fitness_df,
                                          residue_contacts_df,
                                          model_id,
                                          partner_focus) {
  
  interacting_residues <- residue_contacts_df %>%
    dplyr::filter(
      model_name == model_id,
      partner_gene == partner_focus
    ) %>%
    dplyr::pull(residue) %>%
    unique()
  
  fitness_df %>%
    dplyr::mutate(
      interact = aa_mut_pos %in% interacting_residues,
      model = model_id
    )
}



# Function to merge AF3 confidence metrics with model scores
add_confidence_metrics <- function(.data, confidence_df) {
  
  dplyr::left_join(
    .data,
    confidence_df,
    by = "group"
  )
}


# Function to add interface metrics to combined datasets 
add_interface_metrics <- function(df,
                                  A_suffix,
                                  C_suffix = NULL) {
  
  if (!is.null(C_suffix)) {
    # Protein–protein case
    df %>%
      mutate(
        n_A_interface =
          .data[[paste0("a_exposed_outcome_", A_suffix)]] +
          .data[[paste0("b_exposed_no_outcome_", A_suffix)]],
        
        n_C_interface =
          .data[[paste0("a_exposed_outcome_", C_suffix)]] +
          .data[[paste0("b_exposed_no_outcome_", C_suffix)]],
        
        interface_geom_mean =
          sqrt(n_A_interface * n_C_interface),
        
        symmetry_ratio =
          pmin(n_A_interface, n_C_interface) /
          pmax(n_A_interface, n_C_interface)
      )
  } else {
    # Single protein (e.g., A–DNA)
    df %>%
      mutate(
        n_A_interface =
          .data[[paste0("a_exposed_outcome_", A_suffix)]] +
          .data[[paste0("b_exposed_no_outcome_", A_suffix)]]
      )
  }
}



# Function to generate summary table
generate_mutation_category_summary <- function(
    df, 
    class_col,
    category_col,
    total_count = NULL
) {

  class_syms    <- rlang::syms(class_col)
  category_sym  <- rlang::ensym(category_col)

  df %>%
    dplyr::filter(!is.na(!!category_sym)) %>%
    dplyr::group_by(!!!class_syms, !!category_sym) %>%
    dplyr::summarise(count = dplyr::n(), .groups = "drop") %>%
    dplyr::group_by(!!!class_syms) %>%
    dplyr::mutate(
      proportion = if (is.null(total_count)) {
        count / sum(count)
      } else {
        count / total_count
      },
      percent = round(proportion * 100, 2)
    ) %>%
    dplyr::ungroup()
}







```


### 4) plotting 
```{r}
# Function to plot violin plots 
plot_confidence_violin <- function(
  data,
  x_col = "interaction",
  y_col = "value",
  x_order = NULL,
  x_lab = NULL,
  y_lab = NULL,
  plot_title = NULL,
  y_limits = NULL,
  facet_col = NULL,
  facet_scales = "fixed",
  add_thresholds = FALSE,
  threshold_lines = NULL,
  fill_color = "steelblue"
) {
  
  library(ggplot2)
  library(dplyr)
  
  plot_df <- data %>%
    filter(
      !is.na(.data[[y_col]]),
      is.finite(.data[[y_col]])
    )
  
  # Optional custom x-axis order
  if (!is.null(x_order)) {
    plot_df <- plot_df %>%
      mutate(
        !!x_col := factor(.data[[x_col]], levels = x_order)
      ) %>%
      filter(!is.na(.data[[x_col]]))
  }
  
  p <- ggplot(
    plot_df,
    aes(
      x = .data[[x_col]],
      y = .data[[y_col]]
    )
  ) +
    #geom_violin(trim = FALSE, fill = fill_color, alpha = 0.6) +
    geom_violin(trim = FALSE, scale = "width", fill = fill_color, alpha = 0.6) +
    geom_boxplot(width = 0.1, outlier.shape = NA) +
    theme_classic() +
    labs(
      x = x_lab,
      y = y_lab,
      title = plot_title
    ) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1)
    )
  
  # Optional y limits
  if (!is.null(y_limits)) {
    p <- p + coord_cartesian(ylim = y_limits)
  }
  
  # Optional facet
  if (!is.null(facet_col)) {
    p <- p +
      facet_wrap(
        as.formula(paste("~", facet_col)),
        scales = facet_scales
      )
  }
  
  # Optional thresholds
  if (add_thresholds && !is.null(threshold_lines)) {
    for (thr in threshold_lines) {
      p <- p +
        geom_hline(
          yintercept = thr,
          linetype = "dashed",
          color = "black"
        )
    }
  }
  
  return(p)
}


# Function to plot forest showing OR for AF3 predictions
plot_forest_AF3 <- function(
  data,
  label_col,
  est_col,
  low_col,
  high_col,
  p_col = NULL,
  facet_col = NULL,
  title = "Forest plot of log10(OR)",
  ref_line = 0,
  star_y = 1.0,
  facet_scales = "free_x",
  facet_nrow = NULL,
  facet_ncol = 1,
  flip_axes = TRUE,              # 🔹 flip x/y axes
  label_col_order = NULL,        # 🔹 custom order for labels
  show_value_labels = TRUE,      # 🔹 NEW: show numeric label for est
  label_size = 3.3,              # 🔹 size for numeric labels
  label_vjust = -0.6             # 🔹 vertical adjustment (flip_axes=TRUE)
) {
  #library(ggplot2)
  #library(dplyr)
  #library(rlang)

  # --- Prepare data ---
  df <- data %>%
    mutate(
      label = .data[[label_col]],
      est   = .data[[est_col]],
      low   = .data[[low_col]],
      high  = .data[[high_col]],
      pval  = if (!is.null(p_col)) .data[[p_col]] else NA_real_,
      facet = if (!is.null(facet_col)) .data[[facet_col]] else "All"
    ) %>%
    filter(is.finite(est), is.finite(low), is.finite(high))

  # --- Optional label ordering ---
  if (!is.null(label_col_order)) {
    df <- df %>%
      mutate(label = factor(label, levels = label_col_order))
  }

  # --- Compute significance stars ---
  if (!is.null(p_col)) {
    df <- df %>%
      mutate(
        signif_label = case_when(
          is.na(pval)   ~ "",
          pval <= 0.001 ~ "***",
          pval <= 0.01  ~ "**",
          pval <= 0.05  ~ "*",
          TRUE          ~ ""
        )
      )
  } else {
    df <- df %>% mutate(signif_label = "")
  }

  # --- Base mappings depending on flip_axes ---
  if (flip_axes) {
    p <- ggplot(df, aes(x = label, y = est))
    ref_line_aes <- aes(yintercept = ref_line)
    err_aes <- aes(ymin = low, ymax = high)
    star_aes <- aes(y = star_y, x = label, label = signif_label)
    val_label_aes <- aes(label = round(est, 2), y = est, x = label)
  } else {
    p <- ggplot(df, aes(y = label, x = est))
    ref_line_aes <- aes(xintercept = ref_line)
    err_aes <- aes(xmin = low, xmax = high)
    star_aes <- aes(x = star_y, y = label, label = signif_label)
    val_label_aes <- aes(label = round(est, 2), x = est, y = label)
  }

  # --- Build plot base ---
  p <- p +
    {if (flip_axes)
      geom_hline(ref_line_aes, linetype = "dashed", color = "gray60")
    else
      geom_vline(ref_line_aes, linetype = "dashed", color = "gray60")} +
    {if (flip_axes)
      geom_errorbar(err_aes, width = 0, color = "black")
    else
      geom_errorbarh(err_aes, height = 0, color = "black")} +
    geom_point(size = 2.5, color = "black") +
    theme_classic() +
    labs(
      x = if (flip_axes) NULL else expression(log[10] * "(Odds Ratio)"),
      #y = if (flip_axes) expression(log[10] * "(Odds Ratio)") else NULL,
      y = "Odds Ratio",
      title = title
    ) +
    theme(
      axis.text.x = element_text(size = 9, angle = if (flip_axes) 45 else 0, hjust = 1),
      axis.text.y = element_text(size = 9),
      strip.background = element_blank(),
      strip.text = element_text(face = "bold", size = 11)
    )

  # --- Add numeric labels above/beside points ---
  if (show_value_labels) {
    if (flip_axes) {
      p <- p +
        geom_text(
          val_label_aes,
          vjust = label_vjust,
          size = label_size,
          fontface = "plain"
        )
    } else {
      p <- p +
        geom_text(
          val_label_aes,
          hjust = -0.3,
          size = label_size,
          fontface = "plain"
        )
    }
  }

  # --- Add significance stars ---
  if (!is.null(p_col)) {
    p <- p +
      geom_text(
        data = df,
        star_aes,
        hjust = 0.5,
        size = 3.5,
        fontface = "bold"
      )
  }

  # --- Facet support ---
  if (!is.null(facet_col)) {
    p <- p +
      facet_wrap(
        vars(facet),
        scales = facet_scales,
        nrow = facet_nrow,
        ncol = facet_ncol
      )
  }

  return(p)
}

plot_forest_AF3 <- function(
  data,
  label_col,
  est_col,
  low_col,
  high_col,
  p_col = NULL,
  facet_col = NULL,
  title = "Forest plot of log10(OR)",
  ref_line = 0,
  star_y = 1.0,
  facet_scales = "free_x",
  facet_nrow = NULL,
  facet_ncol = 1,
  flip_axes = TRUE,
  label_col_order = NULL,
  facet_col_order = NULL,        # NEW: custom order for facet groups
  show_value_labels = TRUE,
  label_size = 3.3,
  label_vjust = -0.6
) {
  
  library(ggplot2)
  library(dplyr)
  
  # --- Prepare data ---
  df <- data %>%
    mutate(
      label = .data[[label_col]],
      est   = .data[[est_col]],
      low   = .data[[low_col]],
      high  = .data[[high_col]],
      pval  = if (!is.null(p_col)) .data[[p_col]] else NA_real_,
      facet = if (!is.null(facet_col)) .data[[facet_col]] else "All"
    ) %>%
    filter(
      is.finite(est),
      is.finite(low),
      is.finite(high)
    )
  
  # --- Optional label ordering ---
  if (!is.null(label_col_order)) {
    df <- df %>%
      mutate(
        label = factor(label, levels = label_col_order)
      )
  }
  
  # --- Optional facet ordering ---
  if (!is.null(facet_col_order)) {
    df <- df %>%
      mutate(
        facet = factor(facet, levels = facet_col_order)
      ) %>%
      filter(!is.na(facet))
  }
  
  # --- Compute significance stars ---
  if (!is.null(p_col)) {
    df <- df %>%
      mutate(
        signif_label = case_when(
          is.na(pval)    ~ "",
          pval <= 0.001  ~ "***",
          pval <= 0.01   ~ "**",
          pval <= 0.05   ~ "*",
          TRUE           ~ ""
        )
      )
  } else {
    df <- df %>%
      mutate(signif_label = "")
  }
  
  # --- Base mappings depending on flip_axes ---
  if (flip_axes) {
    p <- ggplot(df, aes(x = label, y = est))
    ref_line_aes <- aes(yintercept = ref_line)
    err_aes <- aes(ymin = low, ymax = high)
    star_aes <- aes(y = star_y, x = label, label = signif_label)
    val_label_aes <- aes(label = round(est, 2), y = est, x = label)
  } else {
    p <- ggplot(df, aes(y = label, x = est))
    ref_line_aes <- aes(xintercept = ref_line)
    err_aes <- aes(xmin = low, xmax = high)
    star_aes <- aes(x = star_y, y = label, label = signif_label)
    val_label_aes <- aes(label = round(est, 2), x = est, y = label)
  }
  
  # --- Build plot base ---
  p <- p +
    {
      if (flip_axes) {
        geom_hline(
          ref_line_aes,
          linetype = "dashed",
          color = "gray60"
        )
      } else {
        geom_vline(
          ref_line_aes,
          linetype = "dashed",
          color = "gray60"
        )
      }
    } +
    {
      if (flip_axes) {
        geom_errorbar(
          err_aes,
          width = 0,
          color = "black"
        )
      } else {
        geom_errorbarh(
          err_aes,
          height = 0,
          color = "black"
        )
      }
    } +
    geom_point(size = 2.5, color = "black") +
    theme_classic() +
    labs(
      x = if (flip_axes) NULL else expression(log[10] * "(Odds Ratio)"),
      y = if (flip_axes) "Odds Ratio" else NULL,
      title = title
    ) +
    theme(
      axis.text.x = element_text(
        size = 9,
        angle = if (flip_axes) 45 else 0,
        hjust = 1
      ),
      axis.text.y = element_text(size = 9),
      strip.background = element_blank(),
      strip.text = element_text(
        face = "bold",
        size = 11
      )
    )
  
  # --- Add numeric labels above/beside points ---
  if (show_value_labels) {
    if (flip_axes) {
      p <- p +
        geom_text(
          val_label_aes,
          vjust = label_vjust,
          size = label_size,
          fontface = "plain"
        )
    } else {
      p <- p +
        geom_text(
          val_label_aes,
          hjust = -0.3,
          size = label_size,
          fontface = "plain"
        )
    }
  }
  
  # --- Add significance stars ---
  if (!is.null(p_col)) {
    p <- p +
      geom_text(
        data = df,
        star_aes,
        hjust = 0.5,
        size = 3.5,
        fontface = "bold"
      )
  }
  
  # --- Facet support ---
  if (!is.null(facet_col)) {
    p <- p +
      facet_wrap(
        vars(facet),
        scales = facet_scales,
        nrow = facet_nrow,
        ncol = facet_ncol
      )
  }
  
  return(p)
}



# Function to plot dots bar, directly showing scores
plot_model_score_dots <- function(
  df,
  x_col = "hypothesis_short",
  y_col,
  p_col = NULL,
  or_col = NULL,
  model_col = "group",
  p_cutoff = 0.05,
  or_cutoff = 1,
  highlight_significant = TRUE,
  threshold_value = NULL,
  threshold_label = NULL,
  add_threshold_line = TRUE,
  count_above_threshold = FALSE,
  add_median = TRUE,
  add_n_label = TRUE,
  y_lab = NULL,
  x_lab = NULL,
  plot_title = NULL,
  y_limits = NULL,
  y_break_by = NULL,
  jitter_width = 0.15,
  point_size = 1.6
) {
  library(dplyr)
  library(ggplot2)
  
  # ---- safety checks ----
  required_cols <- c(x_col, y_col, model_col)
  
  if (!is.null(p_col)) {
    required_cols <- c(required_cols, p_col)
  }
  
  if (!is.null(or_col)) {
    required_cols <- c(required_cols, or_col)
  }
  
  missing_cols <- setdiff(required_cols, names(df))
  
  if (length(missing_cols) > 0) {
    stop(
      "These columns are missing from df: ",
      paste(missing_cols, collapse = ", ")
    )
  }
  
  # ---- prepare plotting dataframe ----
  plot_df <- df %>%
    mutate(
      plot_x = .data[[x_col]],
      plot_y = as.numeric(.data[[y_col]]),
      model_id = .data[[model_col]]
    ) %>%
    filter(
      !is.na(plot_x),
      !is.na(plot_y),
      is.finite(plot_y)
    )
  
  # ---- add p-value column if provided ----
  if (!is.null(p_col)) {
    plot_df <- plot_df %>%
      mutate(plot_p = as.numeric(.data[[p_col]]))
  } else {
    plot_df <- plot_df %>%
      mutate(plot_p = NA_real_)
  }
  
  # ---- add OR column if provided ----
  if (!is.null(or_col)) {
    plot_df <- plot_df %>%
      mutate(plot_or = as.numeric(.data[[or_col]]))
  } else {
    plot_df <- plot_df %>%
      mutate(plot_or = NA_real_)
  }
  
  # ---- define significance ----
  if (!is.null(p_col) && !is.null(or_col)) {
    plot_df <- plot_df %>%
      mutate(
        is_significant = !is.na(plot_p) &
          !is.na(plot_or) &
          plot_p <= p_cutoff &
          plot_or > or_cutoff
      )
  } else if (!is.null(p_col)) {
    plot_df <- plot_df %>%
      mutate(
        is_significant = !is.na(plot_p) &
          plot_p <= p_cutoff
      )
  } else {
    plot_df <- plot_df %>%
      mutate(is_significant = FALSE)
  }
  
  # ---- threshold counting ----
  plot_df <- plot_df %>%
    mutate(
      above_threshold = if (!is.null(threshold_value)) {
        plot_y > threshold_value
      } else {
        FALSE
      }
    )
  
  # ---- labels ----
  n_label_df <- plot_df %>%
    group_by(plot_x) %>%
    summarise(
      n_models = n_distinct(model_id),
      n_significant = sum(is_significant, na.rm = TRUE),
      n_above_threshold = sum(above_threshold, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    mutate(
      n_label = case_when(
        highlight_significant && !is.null(p_col) &&
          count_above_threshold && !is.null(threshold_value) ~
          paste0("n=", n_models, "; sig=", n_significant, "; >thr=", n_above_threshold),
        
        highlight_significant && !is.null(p_col) ~
          paste0("n=", n_models, "; sig=", n_significant),
        
        count_above_threshold && !is.null(threshold_value) ~
          paste0("n=", n_models, "; >thr=", n_above_threshold),
        
        TRUE ~ paste0("n=", n_models)
      )
    )
  
  # ---- y limits ----
  if (is.null(y_limits)) {
    y_min <- min(plot_df$plot_y, na.rm = TRUE)
    y_max <- max(plot_df$plot_y, na.rm = TRUE)
    
    if (!is.null(threshold_value)) {
      y_min <- min(y_min, threshold_value, na.rm = TRUE)
      y_max <- max(y_max, threshold_value, na.rm = TRUE)
    }
    
    y_range <- y_max - y_min
    
    if (y_range == 0) {
      y_range <- 0.1
    }
    
    y_limits <- c(
      y_min,
      y_max + 0.12 * y_range
    )
  }
  
  n_label_df <- n_label_df %>%
    mutate(y_pos = y_limits[2])
  
  # ---- plot ----
  p <- ggplot(
    plot_df,
    aes(
      x = plot_x,
      y = plot_y
    )
  )
  
  if (add_threshold_line && !is.null(threshold_value)) {
    p <- p +
      geom_hline(
        yintercept = threshold_value,
        linetype = "dashed",
        color = "red",
        linewidth = 0.45
      )
  }
  
  if (add_median) {
    p <- p +
      stat_summary(
        fun = median,
        geom = "crossbar",
        width = 0.45,
        color = "red",
        linewidth = 0.35,
        linetype = "solid"
      )
  }
  
  if (highlight_significant && !is.null(p_col)) {
    p <- p +
      geom_jitter(
        data = plot_df %>% filter(!is_significant),
        width = jitter_width,
        height = 0,
        size = point_size,
        alpha = 0.5,
        color = "grey50"
      ) +
      geom_jitter(
        data = plot_df %>% filter(is_significant),
        width = jitter_width,
        height = 0,
        size = point_size + 0.5,
        alpha = 0.95,
        color = "red"
      )
  } else {
    p <- p +
      geom_jitter(
        width = jitter_width,
        height = 0,
        size = point_size,
        alpha = 0.75,
        color = "black"
      )
  }
  
  if (!is.null(threshold_label) && !is.null(threshold_value)) {
    p <- p +
      annotate(
        "text",
        x = Inf,
        y = threshold_value,
        label = threshold_label,
        hjust = 1.05,
        vjust = -0.4,
        size = 3,
        color = "red"
      )
  }
  
  if (add_n_label) {
    p <- p +
      geom_text(
        data = n_label_df,
        aes(
          x = plot_x,
          y = y_pos,
          label = n_label
        ),
        inherit.aes = FALSE,
        size = 3,
        angle = 0,
        hjust = 0.5
      )
  }
  
  if (!is.null(y_break_by)) {
    p <- p +
      scale_y_continuous(
        breaks = seq(
          floor(y_limits[1] / y_break_by) * y_break_by,
          ceiling(y_limits[2] / y_break_by) * y_break_by,
          by = y_break_by
        )
      )
  }
  
  p <- p +
    coord_cartesian(
      ylim = y_limits,
      clip = "off"
    ) +
    theme_classic() +
    labs(
      x = x_lab,
      y = y_lab,
      title = plot_title
    ) +
    theme(
      axis.text.x = element_text(
        angle = 45,
        hjust = 1,
        vjust = 1
      ),
      plot.margin = margin(t = 15, r = 10, b = 10, l = 10)
    )
  
  return(p)
}




```




## Load Fitness Data
20260302 Updated, use proteome 1aa mutations fitness data to rank Alphafold 3 models of protein complexes;

### 1) load Proteome Data
```{r}
# Load Protein Fitness Data
load("./data/proteins_data/plotting_data/phix_proteome_dms_mut_Fitness_unique_advanced.RData") 
head(as.data.frame(phix_proteome_dms_mut_Fitness_unique_advanced), 0)
nrow(phix_proteome_dms_mut_Fitness_unique_advanced) # 48136x


load("./data/proteins_data/plotting_data/phix_proteome_dms_aa_pos_Fitness_advanced.RData") 
head(as.data.frame(phix_proteome_dms_aa_pos_Fitness_advanced), 0)
nrow(phix_proteome_dms_aa_pos_Fitness_advanced) # 2327x



```



### 2) Proteins A and C Data 
Remeber to remove A-DNA mutations/residues when studying A-C interactions.
- for enrichment test, remove core residues in protein A. 
- all residues data are provided for annotating interface residues.
```{r}
# check data 
head(as.data.frame(phix_proteome_dms_mut_Fitness_unique_advanced), 0)
table(phix_proteome_dms_mut_Fitness_unique_advanced$aa_mut_gene)


## AA Mutations Data
# Protein A
phix_protein_A_dms_sub_mut_Fitness_AF3_analysis <- phix_proteome_dms_mut_Fitness_unique_advanced %>%
  filter(aa_mut_gene == "A", aa_mut_class == "Missense", phix_aa_mut_orf_class == "single_orf", monomer_structural_class_predicted != "Core")
  #filter(aa_mut_gene == "A", aa_mut_class == "Missense", phix_aa_mut_orf_class == "single_orf", )
nrow(phix_protein_A_dms_sub_mut_Fitness_AF3_analysis)
# [1] 4832

# Protein C
phix_protein_C_dms_sub_mut_Fitness_AF3_analysis <- phix_proteome_dms_mut_Fitness_unique_advanced %>%
  filter(aa_mut_gene == "C", aa_mut_class == "Missense", phix_aa_mut_orf_class == "single_orf")
nrow(phix_protein_C_dms_sub_mut_Fitness_AF3_analysis)
# [1] 1113



## AA Residues Data
# Protein A
phix_protein_A_dms_sub_aa_Fitness_AF3_analysis <- phix_proteome_dms_aa_pos_Fitness_advanced %>%
  filter(aa_mut_gene == "A")
nrow(phix_protein_A_dms_sub_aa_Fitness_AF3_analysis)
# [1] 513

# Protein C
phix_protein_C_dms_sub_aa_Fitness_AF3_analysis <- phix_proteome_dms_aa_pos_Fitness_advanced %>%
  filter(aa_mut_gene == "C")
nrow(phix_protein_C_dms_sub_aa_Fitness_AF3_analysis)
# [1] 86




```




## Load AF3 Predictions Data
### 1) Metadata
- mapping raw prediction models with working files
```{r}
## root
project_root <- "/Users/hwei/Desktop/phix_WGM/structures/project_af3"
#project_root <- "./structures/project_af3"


## Metadata
# 1) structure map, link structures to raw predictions
prediction_map <- read_csv("./structures/project_af3/metadata/model_mapping.csv")
# Rows: 2700 Columns: 12


# 2) chain map, link proteins to chains 
chain_roles <- read_csv("./structures/project_af3/metadata/chain_roles.csv")
# Rows: 6350 Columns: 7

chain_roles <- chain_roles %>%
  mutate(
    group = sub("\\.cif$", "", structure_file)
  )

chain_lookup <- chain_roles %>%
  select(structure_file, chain_id, partner_gene = gene)


# 3) contact files
contact_map <- read_csv("./structures/project_af3/metadata/contact_mapping.csv")
# Rows: 2000 Columns: 4

contact_map <- contact_map %>%
  mutate(
    full_path = file.path(project_root, result_file)
  )
# file.exists(contact_map$full_path[1]) 
# [1] TRUE
file.exists(contact_map$full_path[1:3])
# [1] TRUE TRUE TRUE




```

- get atom contacts, run through all predicted models
```{r}
# library(purrr)
atomic_contacts <- purrr::pmap_dfr(
  list(
    contact_map$full_path,
    contact_map$structure_file,
    contact_map$gene
  ),
  parse_contact_file
) %>%
  dplyr::mutate(
    model_name = sub("\\.cif$", "", structure_file)
  )


# check results
head(atomic_contacts, 0)
n_distinct(atomic_contacts$structure_file)
# [1] 300


# join for parter information
atomic_contacts <- atomic_contacts %>%
  left_join(
    chain_lookup,
    by = c("structure_file", "chain2" = "chain_id")
  )

# check results
atomic_contacts %>%
  #dplyr::filter(gene == "A") %>%
  dplyr::filter(gene == "C") %>%
  dplyr::count(partner_gene)

# 1) A interact
#1 C             57488
#2 DNA          306202
#3 MG             4301
#4 rep          139742

# 2) C interact
#1 A            57488
#2 DNA           4632
#3 MG              25
#4 rep          17370

# save parsed atom distance data
#saveRDS(atomic_contacts, "structures/project_af3/analysis/atomic_contacts.rds")
write_tsv(atomic_contacts, file = "./data/summary_tables/atomic_contacts.tsv")



```

- residue-level interface annotation, collapse to residue contacts
```{r}
# collapse atoms to residues 
residue_contacts <- atomic_contacts %>%
  group_by(model_name, structure_file, gene, res1, partner_gene) %>%
  summarise(
    n_atomic_contacts = n(),
    .groups = "drop"
  ) %>%
  rename(residue = res1) %>%
  mutate(interacts = TRUE)
glimpse(residue_contacts)

# check results
head(as.data.frame(residue_contacts), 0)
nrow(residue_contacts)
# [1] 32102

n_distinct(residue_contacts$structure_file)
# [1] 300

table(residue_contacts$gene)
#     A      C 
# 27555   4547 

# save residue contacts
#saveRDS(residue_contacts, "structures/project_af3/analysis/residue_contacts.rds")
write_tsv(residue_contacts, file = "./data/summary_tables/residue_contacts.tsv")



```

- 1) separate contact information by proteins
```{r}
# separate residue contacts
residue_contacts_A <- residue_contacts %>% filter(gene == "A")
residue_contacts_C <- residue_contacts %>% filter(gene == "C")

# separate models in residue contacts
model_ids_A <- unique(residue_contacts_A$model_name)
model_ids_C <- unique(residue_contacts_C$model_name)


```

- 2) further separate contact information by interactions
```{r}
# get interacting partners
unique(residue_contacts_A$partner_gene)
unique(residue_contacts_C$partner_gene)

# get models include specific interactions
models_ids_A_interact_C   <- get_models_with_partner(residue_contacts_A, "C")
models_ids_A_interact_DNA <- get_models_with_partner(residue_contacts_A, "DNA")
models_ids_A_interact_MG <- get_models_with_partner(residue_contacts_A, "MG")
models_ids_A_interact_rep <- get_models_with_partner(residue_contacts_A, "rep")


models_ids_C_interact_A   <- get_models_with_partner(residue_contacts_C, "A")
models_ids_C_interact_DNA   <- get_models_with_partner(residue_contacts_C, "DNA")
models_ids_C_interact_MG   <- get_models_with_partner(residue_contacts_C, "MG")
models_ids_C_interact_rep   <- get_models_with_partner(residue_contacts_C, "rep")


# models_ids_A_interact_rep

# check model number
length(models_ids_A_interact_C)
# [1] 149
length(models_ids_A_interact_DNA)
# [1] 250
length(models_ids_A_interact_MG)
# [1] 288
length(models_ids_A_interact_rep)
# [1] 100



```



### 2) Confidences data
- FUNCTION: read_AF3_chain_metrics
- FUNCTION: read_AF3_summary_jsons

```{r}
# 1) model-level confidence summary .json files
confidence_table <- read_AF3_summary_jsons(root_path = "structures/project_af3/working_models")

confidence_table_to_bind <- confidence_table %>%
  mutate(
    model = basename(model),  # in case full paths are present
    
    group = str_remove(
      model,
      "_summary_confidences\\.json$"
    )
  ) %>%
  select(-model)


# 2) chain- and chain-pair- levels confidence summary .json files 
af3_metrics <- read_AF3_chain_metrics("structures/project_af3/working_models")
#af3_metrics$chain_level

# prepare chains information
head(as.data.frame(chain_roles), 5)

chain_roles_mapped <- chain_roles %>%
  mutate(chain_index = match(chain_id, sort(unique(chain_id))))


# 3) chain pTM scores
chain_mapped <- map_chain_confidence(
  af3_metrics$chain_level,
  chain_roles_mapped
)
head(as.data.frame(chain_mapped), 0)


## chain pTM
pTM_chain_wide <- chain_mapped %>%
  group_by(group, gene) %>%
  summarise(
    chain_ptm = mean(chain_ptm, na.rm = F),
    .groups = "drop"
  ) %>%
  mutate(col_name = paste0("pTM_", gene)) %>%
  select(group, col_name, chain_ptm) %>%
  pivot_wider(
    names_from = col_name,
    values_from = chain_ptm
  )
head(as.data.frame(pTM_chain_wide), 0)

pTM_chain_long <- pTM_chain_wide %>%
  pivot_longer(
    cols = starts_with("pTM_"),
    names_to = "molecule",
    values_to = "pTM"
  )  %>%
  parse_model_group()


## chain ipTM
ipTM_chain_wide <- chain_mapped %>%
  group_by(group, gene) %>%
  summarise(
    chain_iptm = mean(chain_iptm, na.rm = F),
    .groups = "drop"
  ) %>%
  mutate(col_name = paste0("ipTM_", gene)) %>%
  select(group, col_name, chain_iptm) %>%
  pivot_wider(
    names_from = col_name,
    values_from = chain_iptm
  )
head(as.data.frame(ipTM_chain_wide), 0)

ipTM_chain_long <- ipTM_chain_wide %>%
  pivot_longer(
    cols = starts_with("ipTM_"),
    names_to = "molecule",
    values_to = "ipTM"
  )  %>%
  parse_model_group()


# 4) pairwise ipTM scores
ipTM_pair_mapped <- map_pairwise_confidence(
  af3_metrics$pair_level,
  chain_roles_mapped
)
head(as.data.frame(ipTM_pair_mapped), 5)

# remove directionality and get the interface confidence 
ipTM_pair_unique <- ipTM_pair_mapped %>%
  filter(gene_i != gene_j) %>%
  mutate(
    gene1 = pmin(gene_i, gene_j),
    gene2 = pmax(gene_i, gene_j)
  ) %>%
  group_by(group, gene1, gene2) %>%
  summarise(
    iptm_pair = if (all(is.na(pair_iptm))) NA_real_
                else max(pair_iptm, na.rm = TRUE),
    pae_min_pair = if (all(is.na(pair_pae_min))) NA_real_
                   else min(pair_pae_min, na.rm = TRUE),
    .groups = "drop"
  )

ipTM_pair_wide <- ipTM_pair_unique %>%
  mutate(pair_name = paste0("ipTM_", gene1, "_", gene2)) %>%
  select(group, pair_name, iptm_pair) %>%
  pivot_wider(
    names_from = pair_name,
    values_from = iptm_pair
  )
head(as.data.frame(ipTM_pair_wide), 0)

ipTM_pair_long <- ipTM_pair_wide %>%
  pivot_longer(
    cols = starts_with("ipTM_"),
    names_to = "interface",
    values_to = "ipTM"
  )  %>%
  parse_model_group()


# 5) ipSAE scores
#view(chain_roles)
ipSAE_raw <- read_ipSAE_results("structures/project_af3/analysis/ipSAE")

ipSAE_metrics_to_keep <- c(
  "ipSAE", "ipSAE_d0chn", "ipSAE_d0dom",
  "ipTM_af", "ipTM_d0chn",
  "pDockQ", "pDockQ2", "LIS"
)

ipSAE_pair_mapped <- ipSAE_raw %>%
  left_join(
    chain_roles %>% select(group, chain_id, gene),
    by = c("group", "Chn1" = "chain_id")
  ) %>%
  rename(gene1 = gene) %>%
  left_join(
    chain_roles %>% select(group, chain_id, gene),
    by = c("group", "Chn2" = "chain_id")
  ) %>%
  rename(gene2 = gene)
head(as.data.frame(ipSAE_pair_mapped), 0)


ipSAE_pair_unique <- ipSAE_pair_mapped %>%
  filter(Type == "max", gene1 != gene2) %>%
  mutate(
    geneA = pmin(gene1, gene2),
    geneB = pmax(gene1, gene2)
  ) %>%
  group_by(group, geneA, geneB) %>%
  summarise(
    across(
      all_of(ipSAE_metrics_to_keep),
      ~ if (all(is.na(.))) NA_real_ else max(., na.rm = TRUE)
    ),
    .groups = "drop"
  ) %>%
  rename_with(
    ~ paste0(.x, "_ipSAE_15"),
    .cols = c(ipTM_af, ipTM_d0chn, pDockQ, pDockQ2, LIS)
  )
head(as.data.frame(ipSAE_pair_unique), 0)


ipSAE_pair_wide <- ipSAE_pair_unique %>%
  mutate(
    pair_id = paste(geneA, geneB, sep = "_")
  ) %>%
  pivot_longer(
    cols = -c(group, geneA, geneB, pair_id),
    names_to = "metric",
    values_to = "value"
  ) %>%
  mutate(
    metric_pair = paste0(metric, "_", pair_id)
  ) %>%
  select(group, metric_pair, value) %>%
  pivot_wider(
    names_from = metric_pair,
    values_from = value
  )
head(as.data.frame(ipSAE_pair_wide), 0)


# 6) combine computational confidence for each model
computational_confidence_table <- pTM_chain_wide %>%
  left_join(ipTM_chain_wide, by = "group") %>%
  left_join(ipTM_pair_wide,  by = "group") %>%
  left_join(ipSAE_pair_wide, by = "group")

head(as.data.frame(computational_confidence_table), 0)
#write_tsv(computational_confidence_table, file = "./data/summary_tables/computational_confidence_table.tsv")

# long format
computational_confidence_long <- computational_confidence_table %>%
  pivot_longer(
    cols = -group,
    names_to = "feature",
    values_to = "value"
  ) %>%
  parse_model_group() %>%
  parse_confidence_feature() %>%
  add_hypothesis_short_plot()

#write_tsv(computational_confidence_long, file = "./data/summary_tables/computational_confidence_long.tsv")


# 7) process summary tables
computational_confidence_table_plot <- computational_confidence_table %>%
  parse_model_group() %>%
  add_hypothesis_short_plot()


head(as.data.frame(computational_confidence_table_plot), 2)
table(computational_confidence_table_plot$hypothesis_short)
#         A-dsDNA      A-noDNA-gpc          A-ssDNA     A-ssDNA-2rep A-ssDNA-2rep_gpc      A-ssDNA-gpc 
#              50               50               50               50               50               50 
table(computational_confidence_table_plot$modifier)
#    2rep 2rep_gpc      gpc 
#      50       50      100 


# Quick statistics of for certain proteins or interactions
summarise_confidence_column(
  #computational_confidence_table_plot,
  #computational_confidence_table_plot[computational_confidence_table_plot$hypothesis_short == "A-dsDNA",],
  #computational_confidence_table_plot[computational_confidence_table_plot$hypothesis_short == "A-ssDNA",],
  
  #computational_confidence_table_plot[!is.na(computational_confidence_table_plot$modifier),],
  #computational_confidence_table_plot[computational_confidence_table_plot$hypothesis_short == "A-noDNA-gpc",],
  #computational_confidence_table_plot[computational_confidence_table_plot$hypothesis_short == "A-ssDNA-gpc",],
  computational_confidence_table_plot[computational_confidence_table_plot$hypothesis_short == "A-ssDNA-2rep_gpc",],
  
  #value_col = "pTM_A",
  #value_col = "ipTM_A_DNA",
  
  #value_col = "pTM_A"
  #value_col = "pTM_C"
  value_col = "ipTM_A_C"
)





```


- plotting computational confidence
- Violin plots (all metrics) (Figure S5)
```{r}
# check data
head(as.data.frame(computational_confidence_long), 0)
table(computational_confidence_long$hypothesis_short)
#         A-dsDNA      A-noDNA-gpc          A-ssDNA     A-ssDNA-2rep A-ssDNA-2rep_gpc      A-ssDNA-gpc 
#            3400             3400             3400             3400             3400             3400 


# plotting
violin_AF3_confidence <- plot_confidence_violin(
  #data = computational_confidence_long,
  #data = computational_confidence_long[computational_confidence_long$interaction == "A-C",],
  #data = computational_confidence_long[computational_confidence_long$feature == "pTM_A",],
  #data = computational_confidence_long[computational_confidence_long$feature == "pTM_A" & !is.na(computational_confidence_long$modifier),],
  #data = computational_confidence_long[computational_confidence_long$feature %in% c("ipTM_A_DNA", "pTM_A") & is.na(computational_confidence_long$modifier),],
  
  #data = computational_confidence_long[!is.na(computational_confidence_long$modifier),],
  #data = computational_confidence_long[computational_confidence_long$modifier %in% c("gpc", "2rep_gpc"),],
  #data = computational_confidence_long[computational_confidence_long$modifier %in% c("gpc", "2rep_gpc") & computational_confidence_long$feature %in% c("ipSAE_A_C", "ipSAE_d0chn_A_C", "ipSAE_d0dom_A_C", "ipTM_A_C", "pTM_A", "pTM_C"),],
  data = computational_confidence_long[computational_confidence_long$modifier %in% c("gpc", "2rep_gpc") & computational_confidence_long$feature %in% c("ipTM_A_C", "pTM_A", "pTM_C"),],
  
  #x_col = "interaction",
  #x_col = "metric",
  #x_col = "hypothesis_short",
  
  x_col = "hypothesis_short_plot",
  y_col = "value",
  #x_order = c("A-ssDNA", "A-dsDNA", "A-C", "A-C-DNA", "A-C-DNA-Rep"),
  
  x_lab = "Models",
  y_lab = "Scores",
  
  plot_title = NULL,
  #plot_title = "Computational confidence of A-C interfaces",
  #plot_title = "Computational confidence for A-C models",
  #plot_title = "pTM scores for protein A structure in different models",
  add_thresholds = TRUE,
  
  #threshold_lines = c(0.6, 0.8),
  threshold_lines = c(0.8),
  
  facet_col = "feature",
  
  fill_color = "steelblue"
) 
print(violin_AF3_confidence)


# saving
ggsave(violin_AF3_confidence, filename = "./plots/violin_AF3_confidence_ipTM_A_C.pdf", width = 5, height = 3)
ggsave(violin_AF3_confidence, filename = "./plots/violin_AF3_confidence_pTM_A.pdf", width = 3, height = 3)

ggsave(violin_AF3_confidence, filename = "./plots/violin_AF3_confidence_A_DNA.pdf", width = 3, height = 2.5)
ggsave(violin_AF3_confidence, filename = "./plots/violin_AF3_confidence_A_C.pdf", width = 5, height = 2.5)





```



### 3) RMSD data
- Align protein A models to a A-ssDNA reference
- Summary median RMSD
```{r}
#library(dplyr)
#library(readr)
#library(purrr)


# ---- load all RMSD csv files ----
rmsd_files <- list.files(
  "./structures/project_af3/analysis/RMSD",
  pattern = "\\.csv$",
  full.names = TRUE
)

rmsd_long <- map_dfr(rmsd_files, function(f) {
  
  read_csv(f, show_col_types = FALSE) %>%
    mutate(
      source_file = basename(f),
      comparison = basename(f) %>%
        str_remove("\\.csv$")
    )
}) %>%
  mutate(
    pruned_atom_pairs = as.numeric(pruned_atom_pairs),
    pruned_rmsd_to_reference = as.numeric(pruned_rmsd_to_reference),
    all_atom_pairs = as.numeric(all_atom_pairs),
    all_pair_rmsd_to_reference = as.numeric(all_pair_rmsd_to_reference)
  )

rmsd_long <- rmsd_long %>%
  parse_model_group(group_col = "model") %>%
  add_hypothesis_short_plot(hypothesis_col = "hypothesis_short")

#write_tsv(rmsd_long, file = "./data/summary_tables/rmsd_long.tsv")



# plotting
violin_AF3_RMSD <- plot_confidence_violin(
  data = rmsd_long,
  
  #x_col = "hypothesis_short",
  x_col = "hypothesis_short_plot",
  
  #y_col = "rmsd_to_reference",
  y_col = "all_pair_rmsd_to_reference",
  
  x_lab = "Model",
  y_lab = "RMSD",
  x_order = c("A-ssDNA", "A-dsDNA", "A-C", "A-C-DNA", "A-C-DNA-Rep"),
  
  plot_title = "Computational confidence for A-C models",
  
  add_thresholds = FALSE,
  #y_limits = c(0, 1.5),
  
  #threshold_lines = c(0.6, 0.8),
  #threshold_lines = c(0.8),
  
  #facet_col = "feature",
  
  fill_color = "black"
) 
print(violin_AF3_RMSD)

# saving
ggsave(violin_AF3_RMSD, filename = "./plots/violin_AF3_RMSD.pdf", width = 3, height = 2.5)



```




## Evaluate structure models
- A-DNA and A-C pipelines
  1) Annotate all A interface residues first, then specific interactions;
  2) Run A-DNA models first, get A-DNA interface residues;
  3) Exclude A-DNA residues, evaluate A-C interfaces;

- define essential residues in binary 'TRUE' or 'FALSE' in $essential_flag
- define deleterious&lethal mutations in binary 'TRUE' or 'FALSE' in $deleterious_mut_flag




## Evaluate A-DNA models
### 1) annotating A-DNA models
In both Residue- and Mutation- levels.
```{r}
# check data
head(as.data.frame(phix_protein_A_dms_sub_mut_Fitness_AF3_analysis), 0)
head(as.data.frame(phix_protein_C_dms_sub_mut_Fitness_AF3_analysis), 0)

head(as.data.frame(phix_protein_A_dms_sub_aa_Fitness_AF3_analysis), 0)
head(as.data.frame(phix_protein_C_dms_sub_aa_Fitness_AF3_analysis), 0)


## Protein A
# 1) All interactions (Everything)
annotated_list_A <- setNames(
  lapply(model_ids_A, function(m) {
    annotate_fitness_with_contacts(
      fitness_df = phix_protein_A_dms_sub_aa_Fitness_AF3_analysis,
      residue_contacts_df = residue_contacts_A,
      model_id = m
    ) %>%
      dplyr::mutate(
        essential_flag = residue_category == "Essential"
      )
  }),
  model_ids_A
)

annotated_list_A_mut <- setNames(
  lapply(model_ids_A, function(m) {
    annotate_fitness_with_contacts(
      fitness_df = phix_protein_A_dms_sub_mut_Fitness_AF3_analysis,
      residue_contacts_df = residue_contacts_A,
      model_id = m
    ) %>%
      dplyr::mutate(
        #essential_flag = residue_category == "Essential"
        deleterious_mut_flag = mutation_category_FDR_lethal_1aa %in% c("Deleterious", "Lethal")
      )
  }),
  model_ids_A
)


# 2) Protein A-DNA interactions (Protein-DNA)
annotated_list_A_interact_DNA <- setNames(
  lapply(models_ids_A_interact_DNA, function(m) {
    annotate_fitness_with_partner(
      fitness_df = phix_protein_A_dms_sub_aa_Fitness_AF3_analysis,
      residue_contacts_df = residue_contacts_A,
      model_id = m,
      partner_focus = "DNA"
    ) %>%
      dplyr::mutate(
        essential_flag = residue_category == "Essential"
      )
  }),
  models_ids_A_interact_DNA
)

annotated_list_A_mut_interact_DNA <- setNames(
  lapply(models_ids_A_interact_DNA, function(m) {
    annotate_fitness_with_partner(
      fitness_df = phix_protein_A_dms_sub_mut_Fitness_AF3_analysis,
      residue_contacts_df = residue_contacts_A,
      model_id = m,
      partner_focus = "DNA"
    ) %>%
      dplyr::mutate(
        #essential_flag = residue_category == "Essential"
        deleterious_mut_flag = mutation_category_FDR_lethal_1aa %in% c("Deleterious", "Lethal")
      )
  }),
  models_ids_A_interact_DNA
)




```



### 2) scoring A-DNA models
```{r}
## Protein A 
# 1) All interactions (Everything)
model_scores_A_residue <- dplyr::bind_rows(
  lapply(names(annotated_list_A), function(m) {
    compute_enrichment(
      df = annotated_list_A[[m]],
      exposure_col = interact,
      exposure_focus = TRUE,
      outcome_col = essential_flag,
      outcome_focus = TRUE,
      dataset_label = "Protein_A_residue_ANY",
      model_id = m,
      suffix = "A_res_interact_ANY"
    )
  })
)

model_scores_A_mutation <- dplyr::bind_rows(
  lapply(names(annotated_list_A_mut), function(m) {
    compute_enrichment(
      df = annotated_list_A_mut[[m]],
      exposure_col = interact,
      exposure_focus = TRUE,
      outcome_col = deleterious_mut_flag,
      outcome_focus = TRUE,
      dataset_label = "Protein_A_mutations_ANY",
      model_id = m,
      suffix = "A_mut_interact_ANY"
    )
  })
)

model_scores_A_interact_ALL <- combine_enrichment_tables(
  model_scores_A_residue,
  model_scores_A_mutation
) %>%
  parse_model_group() %>%
  add_confidence_metrics(confidence_table_to_bind) %>%
  add_confidence_metrics(computational_confidence_table)



# 2) Protein A-DNA interactions (Protein-DNA)
model_scores_A_residue_interact_DNA <- dplyr::bind_rows(
  lapply(names(annotated_list_A_interact_DNA), function(m) {
    compute_enrichment(
      df = annotated_list_A_interact_DNA[[m]],
      exposure_col = interact,
      exposure_focus = TRUE,
      outcome_col = essential_flag,
      outcome_focus = TRUE,
      dataset_label = "Protein_A_residue_interact_DNA",
      model_id = m,
      suffix = "A_res_interact_DNA"
    )
  })
)


model_scores_A_mutation_interact_DNA <- dplyr::bind_rows(
  lapply(names(annotated_list_A_mut_interact_DNA), function(m) {
    compute_enrichment(
      df = annotated_list_A_mut_interact_DNA[[m]],
      exposure_col = interact,
      exposure_focus = TRUE,
      outcome_col = deleterious_mut_flag,
      outcome_focus = TRUE,
      dataset_label = "Protein_A_mutations_interact_DNA",
      model_id = m,
      suffix = "A_mut_interact_DNA"
    )
  })
)

head(as.data.frame(model_scores_A_mutation_interact_DNA), 0)
model_scores_A_mutation_interact_DNA_FDR <- model_scores_A_mutation_interact_DNA %>%
  dplyr::mutate(
    p_value_adj_BH_A_mut_interact_DNA = p.adjust(
      p_value_A_mut_interact_DNA,
      method = "BH"
    ),
    
    significant_enriched_A_mut_interact_DNA_p05 =
      odds_ratio_A_mut_interact_DNA > 1 &
      p_value_A_mut_interact_DNA < 0.05,
    
    significant_enriched_A_mut_interact_DNA_FDR10 =
      odds_ratio_A_mut_interact_DNA > 1 &
      p_value_adj_BH_A_mut_interact_DNA < 0.10,
    
    significant_enriched_A_mut_interact_DNA_FDR05 =
      odds_ratio_A_mut_interact_DNA > 1 &
      p_value_adj_BH_A_mut_interact_DNA < 0.05
  )
table(model_scores_A_mutation_interact_DNA_FDR$significant_enriched_A_mut_interact_DNA_p05)
# FALSE  TRUE 
#   15    235 
table(model_scores_A_mutation_interact_DNA_FDR$significant_enriched_A_mut_interact_DNA_FDR10)
# FALSE  TRUE 
#   13    237 
table(model_scores_A_mutation_interact_DNA_FDR$significant_enriched_A_mut_interact_DNA_FDR05)
# FALSE  TRUE 
#   16    234 


model_scores_A_interact_DNA <- combine_enrichment_tables(
  model_scores_A_residue_interact_DNA,
  model_scores_A_mutation_interact_DNA
) %>%
  parse_model_group() %>%
  add_hypothesis_short_plot() %>%
  add_confidence_metrics(confidence_table_to_bind) %>%
  add_confidence_metrics(computational_confidence_table)

model_scores_A_interact_DNA_FDR <- combine_enrichment_tables(
  #model_scores_A_residue_interact_DNA,
  model_scores_A_mutation_interact_DNA_FDR
) %>%
  parse_model_group() %>%
  add_hypothesis_short_plot() %>%
  add_confidence_metrics(confidence_table_to_bind) %>%
  add_confidence_metrics(computational_confidence_table)

#write_tsv(model_scores_A_interact_DNA_FDR, file = "./data/summary_tables/model_scores_A_interact_DNA_FDR.tsv")



```



### 3) evaluate A-DNA models (Figure 5)
- select A-DNA models, ranked by internal computational metrics, ipTM
- get A-DNA interface residues, group into A-ssDNA, A-dsDNA, shared by both
- calculate enrichments of detrimental mutations
```{r}
## Check data
head(as.data.frame(model_scores_A_interact_DNA), 0)
table(model_scores_A_interact_DNA$dna_type)


# get the model names from the scoring table
model_scores_A_interact_DNA  %>%
  dplyr::filter(is.na(modifier)) %>%
  dplyr::filter(dna_type == "dsDNA") %>%
  #dplyr::filter(dna_type == "ssDNA") %>%
  #dplyr::arrange(dplyr::desc(ranking_score)) %>%
  dplyr::arrange(dplyr::desc(iptm)) %>%
  dplyr::slice(1) %>%
  #dplyr::pull(group, ranking_score)
  #dplyr::pull(group, pTM_A)
  dplyr::pull(group, ipTM_A_DNA)

  
# Top-ranked models
# A-dsDNA
# gpa_dsDNA_30nt_seed2_model_0
# model iptm == 0.78
# model pTM_A == 0.86

# A-ssDNA
# gpa_ssDNA_30nt_seed5_model_0
# model iptm == 0.83
# model pTM_A == 0.85


# 1) get interface residues
# A-ssDNA model
A_ssDNA_residues_selected_model <- annotated_list_A_interact_DNA[["gpa_ssDNA_30nt_seed5_model_0"]] %>%
  dplyr::filter(interact == TRUE) %>%
  dplyr::pull(aa_mut_pos) %>%
  unique() %>%
  sort()

A_ssDNA_residues_selected_model %>% paste(collapse = ",")
# [1] "28,31,32,33,45,49,53,144,145,146,156,167,168,171,172,174,202,204,221,225,240,241,242,243,244,249,254,256,257,258,259,261,263,265,288,291,303,305,333,335,338,339,342,343,347,350,383,385,386,387,388,412,413,414,417,420,423,427,446,449,450,451,452,453,456,469,472,473,476,478,499,502,503,506"

length(A_ssDNA_residues_selected_model) 
# [1] 74

# A-dsDNA model
A_dsDNA_residues_selected_model <- annotated_list_A_interact_DNA[["gpa_dsDNA_30nt_seed2_model_0"]] %>%
  dplyr::filter(interact == TRUE) %>%
  dplyr::pull(aa_mut_pos) %>%
  unique() %>%
  sort()

A_dsDNA_residues_selected_model %>% paste(collapse = ",")
# [1] "28,31,32,33,45,49,137,144,145,146,157,158,159,160,161,162,165,168,169,171,172,173,174,175,176,179,183,217,220,221,225,229,239,240,241,242,243,244,249,351,359,383,384,385,386,387,388,389,412,413,414,419,420,423,427,446,447,449,450,451,452,453,454,455,456,457,458,459,461,462,463,466,467,469,470,471,472,473,474,476,478"

length(A_dsDNA_residues_selected_model) 
# [1] 81



# 2) get the union A-DNA residues
A_DNA_residues_selected_model <- union(
  A_ssDNA_residues_selected_model,
  A_dsDNA_residues_selected_model
)
A_DNA_residues_selected_model %>% paste(collapse = ",")

length(A_DNA_residues_selected_model) 
# [1] 111



# 3) group interface residues
ss <- A_ssDNA_residues_selected_model
ds <- A_dsDNA_residues_selected_model

ss_only <- setdiff(ss, ds)
ss_only %>% paste(collapse = ",")
length(ss_only)
# [1] 30

ds_only <- setdiff(ds, ss)
ds_only %>% paste(collapse = ",")
length(ds_only)
# [1] 37

shared <- intersect(ss, ds)
shared %>% paste(collapse = ",")
length(shared)
# [1] 44



# 4) enrichment tests for each
phix_protein_A_dms_sub_mut_AF3_DNA_interface <- phix_protein_A_dms_sub_mut_Fitness_AF3_analysis %>%
  mutate(
    interact_shared = aa_mut_pos %in% shared,
    interact_ss_only = aa_mut_pos %in% ss_only,
    interact_ds_only = aa_mut_pos %in% ds_only
  )

enrich_ss <- run_odds_ratio(
  df = phix_protein_A_dms_sub_mut_AF3_DNA_interface,
  exposure_col = interact_ss_only,
  exposure_focus = TRUE,
  outcome_col = mutation_category_FDR_1aa,
  outcome_focus = "Detrimental",
  dataset = "A_DNA",
  group = "ssDNA-specific \n(30aa)"
)

enrich_ds <- run_odds_ratio(
  df = phix_protein_A_dms_sub_mut_AF3_DNA_interface,
  exposure_col = interact_ds_only,
  exposure_focus = TRUE,
  outcome_col = mutation_category_FDR_1aa,
  outcome_focus = "Detrimental",
  dataset = "A_DNA",
  group = "dsDNA-specific \n(37aa)"
)

enrich_shared <- run_odds_ratio(
  df = phix_protein_A_dms_sub_mut_AF3_DNA_interface,
  exposure_col = interact_shared,
  exposure_focus = TRUE,
  outcome_col = mutation_category_FDR_1aa,
  outcome_focus = "Detrimental",
  dataset = "A_DNA",
  group = "Shared \n(44aa)"
)

enrichment_comparison <- dplyr::bind_rows(
    enrich_ss,
    enrich_ds,
    enrich_shared
  ) %>%
  dplyr::mutate(
    label = group
  )

#write_tsv(enrichment_comparison, file = "./data/summary_tables/enrichment_comparison_A_interact_DNA.tsv")




# 5) plottings 
head(as.data.frame(enrichment_comparison), 0)
table(enrichment_comparison$label)
# dsDNA-specific \n(37aa)         Shared \n(44aa) ssDNA-specific \n(30aa) 
#                       1                       1                       1 

forest_plot_AF3_model_A <- plot_forest_AF3(
  data = enrichment_comparison,
  label_col = "label",
  est_col = "odds_ratio",
  low_col = "conf_low",
  high_col = "conf_high",
  
  #est_col = "log_or_log10",
  #low_col = "log_conf_low_log10",
  #high_col = "log_conf_high_log10",
  
  label_col_order = c("ssDNA-specific \n(30aa)", "dsDNA-specific \n(37aa)", "Shared \n(44aa)"),
  
  p_col = "p_value",
  title = "Enrichment of Detrimental Mutations \nin DNA Interfaces",
  ref_line = 0
)
print(forest_plot_AF3_model_A)


# saving 
ggsave(forest_plot_AF3_model_A, filename = "./plots/forest_plot_AF3_model_A.pdf", width = 2.5, height = 3)





```




## Evaluate A-C models
### 1) annotating A-C models
Removed validated A-DNA interface residues;
```{r}
# check data
head(as.data.frame(phix_protein_A_dms_sub_mut_Fitness_AF3_analysis), 0)
head(as.data.frame(phix_protein_C_dms_sub_mut_Fitness_AF3_analysis), 0)

head(as.data.frame(phix_protein_A_dms_sub_aa_Fitness_AF3_analysis), 0)
head(as.data.frame(phix_protein_C_dms_sub_aa_Fitness_AF3_analysis), 0)


## Protein A
# Protein A-C interactions (Protein-protein)
annotated_list_A_interact_C <- setNames(
  lapply(models_ids_A_interact_C, function(m) {
    annotate_fitness_with_partner(
      fitness_df = phix_protein_A_dms_sub_aa_Fitness_AF3_analysis,
      residue_contacts_df = residue_contacts_A,
      model_id = m,
      partner_focus = "C"
    ) %>%
      
      # ---- add DNA annotation ----
      dplyr::mutate(
        interact_DNA = aa_mut_pos %in% A_DNA_residues_selected_model
      ) %>%

      # ---- exclude DNA residues ----
      dplyr::filter(
        interact == TRUE | interact_DNA == FALSE
        #!interact_DNA
      ) %>%

      # ---- define essential ----
      dplyr::mutate(
        essential_flag = residue_category == "Essential"
      )
    
  }),
  models_ids_A_interact_C
)
#view(annotated_list_A_interact_C[["gpa_ssDNA_30nt_2rep_gpc_seed8_model_3"]])
#table(annotated_list_A_interact_C[["gpa_ssDNA_30nt_2rep_gpc_seed8_model_3"]]$interact)



annotated_list_A_mut_interact_C <- setNames(
  lapply(models_ids_A_interact_C, function(m) {
    annotate_fitness_with_partner(
      fitness_df = phix_protein_A_dms_sub_mut_Fitness_AF3_analysis,
      residue_contacts_df = residue_contacts_A,
      model_id = m,
      partner_focus = "C"
    ) %>%
      
      # ---- add DNA annotation ----
      dplyr::mutate(
        interact_DNA = aa_mut_pos %in% A_DNA_residues_selected_model
      ) %>%

      # ---- exclude DNA-only residues ----
      dplyr::filter(
        interact == TRUE | interact_DNA == FALSE,
        !interact_DNA
      ) %>%

      # ---- define essential ----
      dplyr::mutate(
        #essential_flag = residue_category == "Essential"
        deleterious_mut_flag = mutation_category_FDR_lethal_1aa %in% c("Deleterious", "Lethal")
      )
    
  }),
  models_ids_A_interact_C
)



## Protein C
# 1) All interactions (Everything)
annotated_list_C <- setNames(
  lapply(model_ids_C, function(m) {
    annotate_fitness_with_contacts(
      fitness_df = phix_protein_C_dms_sub_aa_Fitness_AF3_analysis,
      residue_contacts_df = residue_contacts_C,
      model_id = m
    ) %>%
      dplyr::mutate(
        essential_flag = residue_category == "Essential"
      )
  }),
  model_ids_C
)

annotated_list_C_mut <- setNames(
  lapply(model_ids_C, function(m) {
    annotate_fitness_with_contacts(
      fitness_df = phix_protein_C_dms_sub_mut_Fitness_AF3_analysis,
      residue_contacts_df = residue_contacts_C,
      model_id = m
    ) %>%
      dplyr::mutate(
        #essential_flag = residue_category == "Essential"
        deleterious_mut_flag = mutation_category_FDR_lethal_1aa %in% c("Deleterious", "Lethal")
      )
  }),
  model_ids_C
)


# 2) Protein C-A interactions (Protein-protein)
annotated_list_C_interact_A <- setNames(
  lapply(models_ids_C_interact_A, function(m) {
    annotate_fitness_with_partner(
      fitness_df = phix_protein_C_dms_sub_aa_Fitness_AF3_analysis,
      residue_contacts_df = residue_contacts_C,
      model_id = m,
      partner_focus = "A"
    ) %>%
      dplyr::mutate(
        essential_flag = residue_category == "Essential"
      )
  }),
  models_ids_C_interact_A
)

annotated_list_C_mut_interact_A <- setNames(
  lapply(models_ids_C_interact_A, function(m) {
    annotate_fitness_with_partner(
      fitness_df = phix_protein_C_dms_sub_mut_Fitness_AF3_analysis,
      residue_contacts_df = residue_contacts_C,
      model_id = m,
      partner_focus = "A"
    ) %>%
      dplyr::mutate(
        #essential_flag = residue_category == "Essential"
        deleterious_mut_flag = mutation_category_FDR_lethal_1aa %in% c("Deleterious", "Lethal")
      )
  }),
  models_ids_C_interact_A
)


```



### 2) scoring A-C models
```{r}
## Protein A 
# A-C interactions
model_scores_A_residue_interact_C <- dplyr::bind_rows(
  lapply(names(annotated_list_A_interact_C), function(m) {
    compute_enrichment(
      df = annotated_list_A_interact_C[[m]],
      exposure_col = interact,
      exposure_focus = TRUE,
      outcome_col = essential_flag,
      outcome_focus = TRUE,
      dataset_label = "Protein_A_residue_interact_C",
      model_id = m,
      suffix = "A_res_interact_C"
    )
  })
)

model_scores_A_mutation_interact_C <- dplyr::bind_rows(
  lapply(names(annotated_list_A_mut_interact_C), function(m) {
    compute_enrichment(
      df = annotated_list_A_mut_interact_C[[m]],
      exposure_col = interact,
      exposure_focus = TRUE,
      outcome_col = deleterious_mut_flag,
      outcome_focus = TRUE,
      dataset_label = "Protein_A_mutations_interact_C",
      model_id = m,
      suffix = "A_mut_interact_C"
    )
  })
)

head(as.data.frame(model_scores_A_mutation_interact_C), 0)
model_scores_A_mutation_interact_C_FDR <- model_scores_A_mutation_interact_C %>%
  dplyr::mutate(
    p_value_adj_BH_A_mut_interact_C = p.adjust(
      p_value_A_mut_interact_C,
      method = "BH"
    ),
    
    significant_enriched_A_mut_interact_C_p05 =
      odds_ratio_A_mut_interact_C > 1 &
      p_value_A_mut_interact_C < 0.05,
    
    significant_enriched_A_mut_interact_C_FDR10 =
      odds_ratio_A_mut_interact_C > 1 &
      p_value_adj_BH_A_mut_interact_C < 0.10,
    
    significant_enriched_A_mut_interact_C_FDR05 =
      odds_ratio_A_mut_interact_C > 1 &
      p_value_adj_BH_A_mut_interact_C < 0.05
  )
table(model_scores_A_mutation_interact_C_FDR$significant_enriched_A_mut_interact_C_p05)
# FALSE  TRUE 
#   130    19 
table(model_scores_A_mutation_interact_C_FDR$significant_enriched_A_mut_interact_C_FDR10)
# FALSE  TRUE 
#   130    19 
table(model_scores_A_mutation_interact_C_FDR$significant_enriched_A_mut_interact_C_FDR05)
# FALSE  TRUE 
#   133    16 



## Protein C
# 1) All interactions (Everything)
model_scores_C_residue <- dplyr::bind_rows(
  lapply(names(annotated_list_C), function(m) {
    compute_enrichment(
      df = annotated_list_C[[m]],
      exposure_col = interact,
      exposure_focus = TRUE,
      outcome_col = essential_flag,
      outcome_focus = TRUE,
      dataset_label = "Protein_C_residue_ANY",
      model_id = m,
      suffix = "C_res_interact_ANY"
    )
  })
)

model_scores_C_mutation <- dplyr::bind_rows(
  lapply(names(annotated_list_C_mut), function(m) {
    compute_enrichment(
      df = annotated_list_C_mut[[m]],
      exposure_col = interact,
      exposure_focus = TRUE,
      outcome_col = deleterious_mut_flag,
      outcome_focus = TRUE,
      dataset_label = "Protein_C_mutations_ANY",
      model_id = m,
      suffix = "C_mut_interact_ANY"
    )
  })
)


# 2) Protein C-A interactions (Protein-protein)
model_scores_C_residue_interact_A <- dplyr::bind_rows(
  lapply(names(annotated_list_C_interact_A), function(m) {
    compute_enrichment(
      df = annotated_list_C_interact_A[[m]],
      exposure_col = interact,
      exposure_focus = TRUE,
      outcome_col = essential_flag,
      outcome_focus = TRUE,
      dataset_label = "Protein_C_residue_interact_A",
      model_id = m,
      suffix = "C_res_interact_A"
    )
  })
)

model_scores_C_mutation_interact_A <- dplyr::bind_rows(
  lapply(names(annotated_list_C_mut_interact_A), function(m) {
    compute_enrichment(
      df = annotated_list_C_mut_interact_A[[m]],
      exposure_col = interact,
      exposure_focus = TRUE,
      outcome_col = deleterious_mut_flag,
      outcome_focus = TRUE,
      dataset_label = "Protein_C_mutations_interact_A",
      model_id = m,
      suffix = "C_mut_interact_A"
    )
  })
)

head(as.data.frame(model_scores_C_mutation_interact_A), 0)
model_scores_C_mutation_interact_A_FDR <- model_scores_C_mutation_interact_A %>%
  dplyr::mutate(
    p_value_adj_BH_C_mut_interact_A = p.adjust(
      p_value_C_mut_interact_A,
      method = "BH"
    ),
    
    significant_enriched_C_mut_interact_A_p05 =
      odds_ratio_C_mut_interact_A > 1 &
      p_value_C_mut_interact_A < 0.05,
    
    significant_enriched_C_mut_interact_A_FDR10 =
      odds_ratio_C_mut_interact_A > 1 &
      p_value_adj_BH_C_mut_interact_A < 0.10,
    
    significant_enriched_C_mut_interact_A_FDR05 =
      odds_ratio_C_mut_interact_A > 1 &
      p_value_adj_BH_C_mut_interact_A < 0.05
  )
table(model_scores_C_mutation_interact_A_FDR$significant_enriched_C_mut_interact_A_p05)
# FALSE  TRUE 
#   116    33 
table(model_scores_C_mutation_interact_A_FDR$significant_enriched_C_mut_interact_A_FDR10)
# FALSE  TRUE 
#   114    35 
table(model_scores_C_mutation_interact_A_FDR$significant_enriched_C_mut_interact_A_FDR05)
# FALSE  TRUE 
#   116    33 



## Combine enrichments from Proteins A and C

#model_scores_A_interact_C <- combine_enrichment_tables(
  #model_scores_A_residue_interact_C,
  #model_scores_A_mutation_interact_C,
  #model_scores_C_residue_interact_A,
  #model_scores_C_mutation_interact_A
#) %>%
  #parse_model_group() %>%
  #add_hypothesis_short_plot() %>%
  #add_confidence_metrics(confidence_table_to_bind) %>%
  #add_confidence_metrics(computational_confidence_table)

head(as.data.frame(model_scores_A_mutation_interact_C_FDR), 0)
head(as.data.frame(model_scores_C_mutation_interact_A_FDR), 0)

model_scores_A_interact_C_FDR <- combine_enrichment_tables(
  model_scores_A_residue_interact_C,
  model_scores_A_mutation_interact_C_FDR,
  model_scores_C_residue_interact_A,
  model_scores_C_mutation_interact_A_FDR
) %>%
  parse_model_group() %>%
  add_hypothesis_short_plot() %>%
  add_confidence_metrics(confidence_table_to_bind) %>%
  add_confidence_metrics(computational_confidence_table)

head(as.data.frame(model_scores_A_interact_C_FDR), 0)

#write_tsv(model_scores_A_interact_C_FDR, file = "./data/summary_tables/model_scores_A_interact_C_FDR.tsv")


```



### 3) evaluate A-C models (Figure 5)
- select A-C models, ranked by internal computational metrics, ipTM
- get A-C interface residues
- calculate enrichments of detrimental mutations on both sides
```{r}
# check data
head(as.data.frame(model_scores_A_interact_C_FDR), 0)
table(model_scores_A_interact_C_FDR$hypothesis)
#          gpa_noDNA_gpc gpa_ssDNA_30nt_2rep_gpc      gpa_ssDNA_30nt_gpc 
#                     50                      49                      50 

# Quick check the number of models
model_scores_A_interact_C_FDR %>%
  dplyr::mutate(
    plot_p = as.numeric(p_value_adj_BH_C_mut_interact_A),
    plot_or = as.numeric(odds_ratio_C_mut_interact_A),
    is_significant = plot_p <= 0.05 & plot_or > 1
  ) %>%
  dplyr::count(hypothesis_short_plot, is_significant)


# 1) filter models by enrichment of detrimental mutations
model_scores_A_interact_C_FDR_test <- model_scores_A_interact_C_FDR %>%
  dplyr::filter(hypothesis %in% c("gpa_noDNA_gpc", "gpa_ssDNA_30nt_2rep_gpc", "gpa_ssDNA_30nt_gpc")) %>%
  dplyr::mutate(
    C_enriched_sig = (
      odds_ratio_C_mut_interact_A > 1 &
      p_value_adj_BH_C_mut_interact_A < 0.05
      #p_value_adj_BH_C_mut_interact_A < 0.1
    ),
    
    A_enriched_sig = (
      odds_ratio_A_mut_interact_C > 1 &
      p_value_adj_BH_A_mut_interact_C < 0.05
      #p_value_adj_BH_A_mut_interact_C < 0.1
    ),
    
    both_enriched_sig = (
      odds_ratio_A_mut_interact_C > 1 &
      odds_ratio_C_mut_interact_A > 1 &
      p_value_adj_BH_A_mut_interact_C < 0.05 &
      p_value_adj_BH_C_mut_interact_A < 0.05
      #p_value_adj_BH_A_mut_interact_C < 0.1 &
      #p_value_adj_BH_C_mut_interact_A < 0.1
    ),
    
    highlight_class = dplyr::case_when(
      both_enriched_sig ~ "Both",
      A_enriched_sig ~ "A only",
      C_enriched_sig ~ "C only",
      TRUE ~ "None"
    ),
    
    C_A_interface = a_exposed_outcome_C_res_interact_A + b_exposed_no_outcome_C_res_interact_A,
    A_C_interface = a_exposed_outcome_A_res_interact_C + b_exposed_no_outcome_A_res_interact_C
  ) 

head(as.data.frame(model_scores_A_interact_C_FDR_test), 0)
table(model_scores_A_interact_C_FDR_test$highlight_class)
# A only   Both C only   None 
#     11      5     28    105 

#write_tsv(model_scores_A_interact_C_FDR_test, file = "./data/summary_tables/model_scores_A_interact_C_FDR_test.tsv")


# 2) rank the models ranking scores (higher is better)
model_scores_A_interact_C_FDR_test_selected_models <- model_scores_A_interact_C_FDR_test %>%
  dplyr::filter(
    highlight_class == "Both"
  ) %>%
  #dplyr::arrange(desc(C_A_interface)) 
  dplyr::arrange(desc(ipTM_A_C)) 

# Top-ranked model
# gpa_ssDNA_30nt_2rep_gpc_seed8_model_3

model_scores_A_interact_C_FDR_test_selected_models$group
# [1] "gpa_ssDNA_30nt_2rep_gpc_seed8_model_3" "gpa_ssDNA_30nt_2rep_gpc_seed9_model_0" "gpa_ssDNA_30nt_2rep_gpc_seed9_model_4" "gpa_ssDNA_30nt_2rep_gpc_seed2_model_3" "gpa_ssDNA_30nt_2rep_gpc_seed4_model_2"

## Top-ranked model: "gpa_ssDNA_30nt_2rep_gpc_seed8_model_3"


# 3) transform the scoring table to long format
head(as.data.frame(model_scores_A_interact_C_FDR_test_selected_models), 0)

model_scores_FDR_long <- model_scores_A_interact_C_FDR_test_selected_models %>%
  
  dplyr::transmute(
    group,
    
    # ---- A ----
    OR_A   = odds_ratio_A_mut_interact_C,
    low_A  = conf_low_A_mut_interact_C,
    high_A = conf_high_A_mut_interact_C,
    p_A    = p_value_A_mut_interact_C,
    
    # ---- C ----
    OR_C   = odds_ratio_C_mut_interact_A,
    low_C  = conf_low_C_mut_interact_A,
    high_C = conf_high_C_mut_interact_A,
    p_C    = p_value_C_mut_interact_A
  ) %>%
  
  tidyr::pivot_longer(
    cols = -group,
    names_to = c(".value", "protein"),
    names_pattern = "(OR|low|high|p)_(A|C)"
  )


model_scores_FDR_long <- model_scores_A_interact_C_FDR_test_selected_models %>%
  
  dplyr::transmute(
    group,
    hypothesis,
    hypothesis_short,
    hypothesis_short_plot,
    seed,
    model,
    
    A_C_interface,
    C_A_interface,
    
    # ---- A side: A mutations at A-C interface ----
    OR_A   = odds_ratio_A_mut_interact_C,
    low_A  = conf_low_A_mut_interact_C,
    high_A = conf_high_A_mut_interact_C,
    p_A    = p_value_A_mut_interact_C,
    p_adj_A = p_value_adj_BH_A_mut_interact_C,
    
    # ---- C side: C mutations at C-A interface ----
    OR_C   = odds_ratio_C_mut_interact_A,
    low_C  = conf_low_C_mut_interact_A,
    high_C = conf_high_C_mut_interact_A,
    p_C    = p_value_C_mut_interact_A,
    p_adj_C = p_value_adj_BH_C_mut_interact_A
  ) %>%
  
  tidyr::pivot_longer(
    cols = c(
      OR_A, low_A, high_A, p_A, p_adj_A,
      OR_C, low_C, high_C, p_C, p_adj_C
    ),
    names_to = c(".value", "protein"),
    names_pattern = "(OR|low|high|p|p_adj)_(A|C)"
  ) %>%
  
  dplyr::mutate(
    interface = dplyr::case_when(
      protein == "A" ~ "A-C",
      protein == "C" ~ "C-A",
      TRUE ~ NA_character_
    ),
    
    interface_size = dplyr::case_when(
      protein == "A" ~ A_C_interface,
      protein == "C" ~ C_A_interface,
      TRUE ~ NA_real_
    ),
    
    interface_plot = paste0(
      interface,
      "\n(",
      interface_size,
      "aa)"
    ),
    
    log10_OR = log10(OR),
    log10_low = log10(low),
    log10_high = log10(high),
    
    enriched_sig_FDR10 = OR > 1 & p_adj < 0.10,
    enriched_sig_FDR05 = OR > 1 & p_adj < 0.05
  )

#write_tsv(model_scores_FDR_long, file = "./data/summary_tables/mutagenesis_supported_A_interact_C.tsv")


# 4) get interface residues
A_C_residues_selected_model <- annotated_list_A_interact_C[["gpa_ssDNA_30nt_2rep_gpc_seed8_model_3"]] %>%
  dplyr::filter(interact == TRUE) %>%
  dplyr::pull(aa_mut_pos) %>%
  unique() %>%
  sort()
length(A_C_residues_selected_model)
A_C_residues_selected_model %>% paste(collapse = ",")


C_A_residues_selected_model <- annotated_list_C_interact_A[["gpa_ssDNA_30nt_2rep_gpc_seed8_model_3"]] %>%
  dplyr::filter(interact == TRUE) %>%
  dplyr::pull(aa_mut_pos) %>%
  unique() %>%
  sort()
length(C_A_residues_selected_model)
C_A_residues_selected_model %>% paste(collapse = ",")


# get residues
A_C_model_interface_residue_summary <- collapse_model_interacting_residues(
  residue_contacts,
  #model_name_keep = "gpa_ssDNA_30nt_seed5_model_0"
  #model_name_keep = c("gpa_ssDNA_30nt_seed5_model_0", "gpa_dsDNA_30nt_seed2_model_0")
  
  model_name_keep = "gpa_ssDNA_30nt_2rep_gpc_seed8_model_3"
)

head(as.data.frame(A_C_model_interface_residue_summary), 2)


# 5) plottings 
head(as.data.frame(model_scores_FDR_long), 0)
table(model_scores_FDR_long$group)
# [1] "gpa_ssDNA_30nt_2rep_gpc_seed8_model_3" "gpa_ssDNA_30nt_2rep_gpc_seed9_model_0" "gpa_ssDNA_30nt_2rep_gpc_seed9_model_4" "gpa_ssDNA_30nt_2rep_gpc_seed2_model_3" "gpa_ssDNA_30nt_2rep_gpc_seed4_model_2"


forest_plot_AF3_model_A_C <- plot_forest_AF3(
  #data = model_scores_FDR_long,
  #data = model_scores_FDR_long[model_scores_FDR_long$group == "gpa_ssDNA_30nt_2rep_gpc_seed8_model_3",],
  #data = model_scores_FDR_long[model_scores_FDR_long$group == "gpa_ssDNA_30nt_2rep_gpc_seed9_model_0",],
  #data = model_scores_FDR_long[model_scores_FDR_long$group == "gpa_ssDNA_30nt_2rep_gpc_seed9_model_4",],
  #data = model_scores_FDR_long[model_scores_FDR_long$group == "gpa_ssDNA_30nt_2rep_gpc_seed2_model_3",],
  #data = model_scores_FDR_long[model_scores_FDR_long$group == "gpa_ssDNA_30nt_2rep_gpc_seed4_model_2",],
  
  data = model_scores_FDR_long[model_scores_FDR_long$group != "gpa_ssDNA_30nt_2rep_gpc_seed8_model_3",],
  
  #label_col = "protein",
  label_col = "interface_plot",
  
  est_col = "OR",
  low_col = "low",
  high_col = "high",
  p_col = "p",
  
  #facet_col = "dataset",
  facet_col = "group",
  
  facet_col_order = c("gpa_ssDNA_30nt_2rep_gpc_seed9_model_0", "gpa_ssDNA_30nt_2rep_gpc_seed9_model_4", "gpa_ssDNA_30nt_2rep_gpc_seed2_model_3", "gpa_ssDNA_30nt_2rep_gpc_seed4_model_2"),
  
  facet_nrow = 1,
  facet_ncol = 4,
  
  title = "Enrichment of Detrimental Mutations \nin DNA Interfaces",
  ref_line = 0
)
print(forest_plot_AF3_model_A_C)


# saving 
ggsave(forest_plot_AF3_model_A_C, filename = "./plots/forest_plot_AF3_model_C_A_s8m3.pdf", width = 2.5, height = 2.5)
ggsave(forest_plot_AF3_model_A_C, filename = "./plots/forest_plot_AF3_model_C_A_additional_models.pdf", width = 6, height = 3)






```




## Summary evaluations
### 1) Combine computational and experimental scores
```{r}

# combine enrichment scores
experimental_enrichment_scores <- combine_enrichment_tables(
  #model_scores_A_residue,
  model_scores_A_mutation,
  #model_scores_A_residue_interact_DNA,
  model_scores_A_mutation_interact_DNA,
  #model_scores_A_residue_interact_C,
  model_scores_A_mutation_interact_C,
  #model_scores_A_residue_interact_MG,
  model_scores_A_mutation_interact_MG,
  #model_scores_A_residue_interact_rep,
  model_scores_A_mutation_interact_rep,
  
  #model_scores_C_residue,
  model_scores_C_mutation,
  #model_scores_C_residue_interact_DNA,
  model_scores_C_mutation_interact_DNA,
  #model_scores_C_residue_interact_A,
  model_scores_C_mutation_interact_A,
  #model_scores_C_residue_interact_MG,
  model_scores_C_mutation_interact_MG,
  #model_scores_C_residue_interact_rep,
  model_scores_C_mutation_interact_rep
) #%>%
  #parse_model_group() #%>%
  #add_confidence_metrics(confidence_table_to_bind) %>%
  #add_confidence_metrics(computational_confidence_table)

head(as.data.frame(experimental_enrichment_scores), 0)



# convert to long format
experimental_enrichment_long <- experimental_enrichment_scores %>%
  pivot_longer(
    cols = -group,
    names_to = "feature",
    values_to = "value"
  )

experimental_enrichment_long <- experimental_enrichment_long %>%
  #filter(level == "mut") %>%
  parse_enrichment_feature() %>%
  parse_model_group() %>%
  add_hypothesis_short_plot()

head(as.data.frame(experimental_enrichment_long), 0)
table(experimental_enrichment_long$metric_prefix)
table(experimental_enrichment_long$modifier)
table(experimental_enrichment_long$modifier_plot)
table(experimental_enrichment_long$interaction)





```




## Plottings
### 1) dots plots (Figure S5)
Directly showing enrichment scores, and highlight significant enrichments;
```{r}
# check data
#head(as.data.frame(model_scores_A_interact_DNA), 0)
#table(model_scores_A_interact_DNA$hypothesis_short)

head(as.data.frame(model_scores_A_interact_DNA_FDR), 0)
table(model_scores_A_interact_DNA_FDR$hypothesis_short_plot)

head(as.data.frame(model_scores_A_interact_C_FDR), 0)
table(model_scores_A_interact_C_FDR$hypothesis_short_plot)
#        A-C     A-C-DNA A-C-DNA-Rep 
#         50          50          49 



# plottings 
dot_DMS_enrichment <- plot_model_score_dots(
  #df = model_scores_A_interact_DNA_FDR,
  #df = model_scores_A_interact_DNA_FDR[model_scores_A_interact_DNA_FDR$hypothesis_short %in% c("A-dsDNA", "A-ssDNA"),],
  
  df = model_scores_A_interact_C_FDR,
  
  #x_col = "hypothesis_short",
  x_col = "hypothesis_short_plot",
  
  #or_col = "odds_ratio_A_mut_interact_DNA",
  #y_col = "odds_ratio_A_mut_interact_DNA",
  
  #or_col = "odds_ratio_A_mut_interact_C",
  #y_col = "odds_ratio_A_mut_interact_C",
  
  or_col = "odds_ratio_C_mut_interact_A",
  y_col = "odds_ratio_C_mut_interact_A",
  
  x_lab = "Model",
  y_lab = "Enrichment",
  
  #p_col = "p_value_adj_BH_A_mut_interact_DNA",
  #p_col = "p_value_adj_BH_A_mut_interact_C",
  p_col = "p_value_adj_BH_C_mut_interact_A",
  
  p_cutoff = 0.05,
  
  highlight_significant = T,
  
  #plot_title = "A–DNA interface confidence",
  plot_title = NULL,
  
  y_limits = c(0, 6),
  
  #y_break_by = 0.2,
  
  threshold_value = 1,
  threshold_label = "OR = 1"
)
dot_DMS_enrichment


# saving
ggsave(dot_DMS_enrichment, filename = "./plots/dot_DMS_enrichment_A_DNA.pdf", width = 3, height = 2.5)
ggsave(dot_DMS_enrichment, filename = "./plots/dot_DMS_enrichment_A_C.pdf", width = 3, height = 2.5)
ggsave(dot_DMS_enrichment, filename = "./plots/dot_DMS_enrichment_C_A.pdf", width = 3, height = 2.5)




```




## Predicted Interfaces
Prepare predicted interface look-up tables, to annotate single-amino-acid variant datasets.
```{r}
# Function get contacts from residue table
collapse_model_interacting_residues <- function(
  residue_contacts,
  model_name_keep,
  only_interacting = TRUE,
  add_all_by_gene = TRUE
) {
  
  library(dplyr)
  
  df <- residue_contacts %>%
    filter(model_name %in% model_name_keep)
  
  if (only_interacting) {
    df <- df %>%
      filter(interacts == TRUE)
  }
  
  # ---- 1. Specific interaction summary, e.g. A-DNA, A-MG ----
  interaction_summary <- df %>%
    mutate(
      interaction = paste(gene, partner_gene, sep = "-")
    ) %>%
    group_by(model_name, structure_file, interaction, gene, partner_gene) %>%
    summarise(
      n_residues = n_distinct(residue),
      total_atomic_contacts = sum(n_atomic_contacts, na.rm = TRUE),
      interacting_residues = paste(sort(unique(residue)), collapse = ","),
      .groups = "drop"
    )
  
  if (!add_all_by_gene) {
    return(interaction_summary %>% arrange(model_name, interaction))
  }
  
  # ---- 2. All-partner summary by gene, e.g. A-all ----
  all_gene_summary <- df %>%
    group_by(model_name, structure_file, gene) %>%
    summarise(
      interaction = paste0(first(gene), "-all"),
      partner_gene = "all",
      n_residues = n_distinct(residue),
      total_atomic_contacts = sum(n_atomic_contacts, na.rm = TRUE),
      interacting_residues = paste(sort(unique(residue)), collapse = ","),
      .groups = "drop"
    ) %>%
    select(
      model_name,
      structure_file,
      interaction,
      gene,
      partner_gene,
      n_residues,
      total_atomic_contacts,
      interacting_residues
    )
  
  bind_rows(interaction_summary, all_gene_summary) %>%
    arrange(model_name, gene, partner_gene)
}


# Function to collect interface residues in a look-up table
make_residue_interface_lookup <- function(
  residue_contacts,
  model_name_keep = NULL,
  only_interacting = TRUE
) {
  
  library(dplyr)
  
  df <- residue_contacts
  
  # Optional: keep selected models only
  if (!is.null(model_name_keep)) {
    df <- df %>%
      filter(model_name %in% model_name_keep)
  }
  
  # Optional: keep only interacting residues
  if (only_interacting) {
    df <- df %>%
      filter(interacts == TRUE)
  }
  
  df %>%
    group_by(gene, residue) %>%
    summarise(
      interaction_partner = paste(sort(unique(partner_gene)), collapse = ","),
      n_interaction_partners = n_distinct(partner_gene),
      total_atomic_contacts = sum(n_atomic_contacts, na.rm = TRUE),
      n_models_detected = n_distinct(model_name),
      model_names = paste(sort(unique(model_name)), collapse = ","),
      
      # New: keep AF3 model/hypothesis/binder complex information
      binder_complexes = if ("hypothesis_short" %in% names(df)) {
        paste(sort(unique(hypothesis_short)), collapse = ",")
      } else {
        NA_character_
      },
      
      is_interface_residue = TRUE,
      .groups = "drop"
    ) %>%
    rename(
      aa_mut_gene = gene,
      aa_mut_pos = residue
    ) %>%
    arrange(aa_mut_gene, aa_mut_pos)
}


## Summary of final AF3 models
# 1) get residues summary
A_DNA_model_interface_residue_summary <- collapse_model_interacting_residues(
  residue_contacts,
  #model_name_keep = "gpa_ssDNA_30nt_seed5_model_0"
  model_name_keep = c("gpa_ssDNA_30nt_seed5_model_0", "gpa_dsDNA_30nt_seed2_model_0")
)

head(as.data.frame(A_DNA_model_interface_residue_summary), 2)

# extract complete A-DNA interfaces
A_interface_residue <- A_DNA_model_interface_residue_summary %>%
  filter(interaction == "A-all") %>%
  pull(interacting_residues) %>%
  str_split(",") %>%
  unlist() %>%
  str_trim() %>%
  as.integer() %>%
  unique() %>%
  sort() #%>%
  #paste(collapse = ",")
length(A_interface_residue)
A_interface_residue %>% paste(collapse = ",")



# 2) A-DNA interface lookup table
residue_contacts_expanded <- residue_contacts %>%
  parse_model_group(group_col = "model_name")

head(as.data.frame(residue_contacts_expanded), 2)

A_DNA_interface_lookup_table <- make_residue_interface_lookup(
  #residue_contacts,
  #residue_contacts[residue_contacts$partner_gene != "rep",],
  
  #residue_contacts_expanded,
  residue_contacts_expanded[residue_contacts_expanded$partner_gene != "rep",],
  
  model_name_keep = c(
    "gpa_ssDNA_30nt_seed5_model_0",
    "gpa_dsDNA_30nt_seed2_model_0"
    #"gpa_ssDNA_30nt_2rep_gpc_seed8_model_3"
  ),
  only_interacting = TRUE
)

A_DNA_interface_lookup_table



# 3) A-DNA and A-C interfaces lookup table
af3_interface_lookup_table <- make_residue_interface_lookup(
  #residue_contacts,
  #residue_contacts[residue_contacts$partner_gene != "rep",],
  
  #residue_contacts_expanded,
  #residue_contacts_expanded[residue_contacts_expanded$partner_gene != "rep",],
  residue_contacts_expanded %>%
    dplyr::filter(
      partner_gene != "rep",
      
      # remove DNA contacts from the A-ssDNA-2rep_gpc model
      !(partner_gene == "DNA" & hypothesis_short == "A-ssDNA-2rep_gpc")
    ),
  
  model_name_keep = c(
    "gpa_ssDNA_30nt_seed5_model_0",
    "gpa_dsDNA_30nt_seed2_model_0",
    "gpa_ssDNA_30nt_2rep_gpc_seed8_model_3"
  ),
  only_interacting = TRUE
) 

af3_interface_lookup_table


## Saving
#save(af3_interface_lookup_table, file = "./data/proteins_data/PPI_data/af3_interface_lookup_table.RData")
write_tsv(af3_interface_lookup_table, file = "./data/summary_tables/af3_interface_lookup_table.tsv")



```





## Save script checkpoint
save a snapshot of all objects produced by the script

```{r}
# checkpoint - 06 evaluate structure predictions with fitness measurements in interfaces, select A-ssDNA, A-dsDNA, and A-C models for mechanistic analysis
# save.image(file = "data/checkpoints/06_evaluate_AF3.RData")


```


save a working environment for running all analysis script
```{r}
# save the working environment image
save.image(file = "data/phix_WGM_analysis.RData")


```








